### 2018-08-21 boneyard.tex
Initial ShareLaTeX Import

\item Event boundaries: these shape the distribution because they sharply manipulate similarity as a function of time (whereas within an event similarity varies smoothly with time)
\item These ideas should affect time judgments: biases and variability in time judgments should reflect the shape of the memory distribution

At its core, characterizing episodic memory as mental time travel makes studying and understanding memory about matching up the thoughts (and/or brain activity patterns) participants are experiencing at one moment with the thoughts (and/or brain activity patterns) the experienced at another (e.g.\ prior) moment.

When experience is changing monotonically (i.e.\ in expectation, similarity in experience falls off monotonically with temporal distance, as during random word list experiments) this blurring out over time

\item This view elegantly consolidated years of memory theory (cites) and has shaped much of the modern literature on episodic memory since
\item Context-based memory models (SAM, TCM, derivatives) say that when we retrieve past mental states we reactivate a mixture of prior moments, often centered around the time of the remembered episode.
\item Theoretical advantage: this spreading allows us to connect nearby moments in time
\item Alternative interpretation: what we actually experience when we recollect the past is  not a specific moment, but rather a weighted blend of moments.  This is how we model memory retrieval, so we should discuss memory retrieval as a distribution rather than as jumping back in time.

### 2018-08-21 main.tex
Initial ShareLaTeX Import

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{pxfonts}

\title{Episodic memory: mental time travel or a quantum wave function?}
\author{Jeremy R. Manning\\Dartmouth College, Hanover, NH\\jeremy.r.manning@dartmouth.edu}

\begin{document}
\maketitle

\begin{abstract}
Where do we ``go'' when we recollect our past?  When remembering a past event, it is intuitive to imagine some part of ourselves mentally ``jumping back in time'' to when the event occurred. I propose an alternative view, inspired by recent evidence from my lab and others, as well as by re-examining existing models of episodic recall, that suggests that this notion of mentally revisiting any specific moment of our past is at best incomplete and at worst misleading.  Instead, I suggest that we retrieve information from our past by mentally casting ourselves back simultaneously to \textit{many} time points from our past, much like a quantum wave function spreading its probability mass over many possible states.  This revised conceptual model makes important behavioral and neural predictions about how we retrieve information about our past, and has implications for how we study episodic memory experimentally.
\end{abstract}

\section*{Introduction and overview}
How do our brains organize, retrieve, and act upon the ceaseless flow of incoming information and experiences?  \cite{Tulv83} contends that episodic memory is like a sort of ``mental time travel,'' whereby we send some part of our mental state back along our autobiographical time line to a specific moment in our past.  This view has fundamental implications for how we study and build theories about memory: it suggests that we can match up specific moments of recollection with specific moments from the past.  These implications extend to how we design episodic memory experiments and how we analyze the behavioral and neural data from those experiments in order to understand the neural and/or cognitive mechanisms that underly memory retrieval.  At its core, characterizing episodic memory as mental time travel makes studying and understanding memory about matching up the thoughts (and/or brain activity patterns) participants are experiencing at one moment with the thoughts (and/or brain activity patterns) participants experienced at another previous moment.  The challenge, then, is to understand the neural (structural) and cognitive (functional) mechanisms that characterize when, how, and why our memory systems allow us to relive those prior experiences.  One way of illustrating how this framing misses out on key aspects of memory is to examine how experimentalists and theorists typically conceive of and model a mental construct called \textit{context}.

The notion of context was fundamental to \citeauthor{Tulv83}'s characterization of episodic recall, and continues to be a primary feature of most current theories of episodic memory~\citep[e.g.\ for review see][]{Kaha12}.  Whereas the \textit{content} of a memory provides specific information about the episode itself, contextual components of a memory help to frame the episode within the rememberer's broader experience.  Defining context precisely (e.g.\ as distinct from content) is an ongoing challenge for our field, but theorists generally describe context as reflecting a mental combination of external cues (where you are, who you are with, background sensory information, etc.) and internal cues (thoughts, feelings, emotions, etc.) that uniquely define each moment~\citep[e.g.\ see review by][]{MannEtal15}.  Reactivating the mental representation of the context associated with a given episode is what can give us the feeling of some ``part of ourselves'' re-experiencing the past.

One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, FolkEtal18}.  One intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.  Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.  This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12}.  Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}. This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11}. These ideas have also been formalized in a family of theoretical models by Marc Howard and colleagues~\citep{HowaKaha02, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14}.  In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.  Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.  First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).  Second, the studies suggest that members of this family of representations drifting at different rates become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.  These reactivations explain how our brains reactivate prior contexts when we remember the past.

Although reactivating context is often \textit{described} as a sort of mental time travel back to the one specific moment being remembered, there is some apparent inconsistency between this description and how this process is actually characterized mathematically or measured neurally.  In particular, both the theory and neural measures described in the preceding paragraphs characterize episodic recall mathematically as reflecting the reactivation of a weighted blend of thoughts from a \textit{range} of previously experienced times-- not to any one \textit{specific} time.  This is not simply a matter of imprecision in mental time travel (i.e.\ a reflection of the low effective resolution with which we can revisit specific moments from our past).  Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences. The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each drifting at a different timescale.  The notion of thinking of any one particular moment from the past is incompatible with this multiple timescales view, in that our thoughts and brain activity patterns reflect information from a range of moments, even when we are not specifically engaged in remembering.  If what we consider to be a ``moment'' or ``now'' is guided by our neural representations of our ongoing experiences, then this implies we are continually experiencing \textit{many} ``nows.''  If our thoughts are continually spread across a range of times, the notion of mentally revisiting any particular time loses its meaning.

Where the mental time travel description breaks down most notably is when we consider scenarios in which the particular blend of timepoints reflected in one's current thoughts do not come from temporally contiguous events.  For example, understanding how a new experience fits in with our broader understanding might require integrating information gleaned from many experiences separated in time, each with their own contextual attributes.  That we can integrate information across experiences suggests that the integration of incoming information at different rates need not be a purely mechanical process, but rather may reflect a deeper functional role.  When we simultaneously reactivate thoughts related to separated moments from our past, this provides a means of associating and relating those temporally distinct experiences.  This is how we can integrate information across the \textit{many} discrete experiences relevant to ``now.''

\section*{The temporal dynamics of ongoing experience and how we remember it}
Modern context-based theories of episodic memory posit that the current state of mental context serves to determine which information from our past may be relevant to us and therefore reactivated~\citep[e.g.\ ][]{PolyEtal09}. Although these theories were largely developed and tested in the domain of list-learning paradigms, generally similar principles might be expected to play out in real world memory as well.  For example, prior experiences with shared or overlapping contextual properties (e.g.\ other experiences that occurred in similar spatial or social settings, shared similar goals, etc.) could be leveraged to form situation models that help guide behaviors and expectations according to the current perceived context~\citep{RangRitc12}.

Unlike purely random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g.\ shifting from the quiet peace of writing a paper to the different peace of consoling a distraught child).  A number of studies (using non-random lists and more naturalistic stimuli) have found that the way we remember past experiences can be shaped by these event boundaries.  For example, controlling for elapsed time, memory is impaired for information that occurred prior to a change in the current event or situation~\citep[e.g.\ ][]{RadvCope06, SwalEtal09, SwalEtal11, EzzyDava11, MannEtal16}.  The temporal dynamics of contextual change also underlie how we judge the amount of time elapsed since a prior reference point~\citep{BlocReed78, SahaSmit14}.  These findings suggest that the rapid contextual or situational changes that define these event boundaries shape how we experience and remember, similarly to how spatial boundaries~\citep[e.g.\ environmental barriers;][]{McKeBuzs16, BrunEtal18} shape how we experience and remember spatial environments and layouts, or how conceptual boundaries~\citep[e.g.\ distinctions between semantic categories;][]{BrunEtal18} shape how we experience and remember conceptual information.

Event boundaries are one example of the broader temporal covariance structure that is characteristic of ongoing experience.  Whereas classic approaches to studying memory encourage researchers to treat each moment as separable from the rest of experience, context-based theories of memory (including situation models and event-based models) posit that each moment derives meaning through its deep ties to other experiences.  This rich tapestry of interacting moments and experiences forms the scaffolding for interpreting our new experiences and for retrieving information concerning our past.  According to this view, the reactivation of this rich tapestry of contextual details enfolding the remembered event (i.e.\ how the event relates to the rest of ones experiences across timescales) is even more central to our subjective experience than the specific details of what occurred during that event!  This begs the question: what does it mean to \textit{remember} our past experiences?

\section*{The shape of remembered experience}

\begin{itemize}
\item Not about the number of details-- it's about capturing the overall ``shape'' of past experience
\item Quantifying recall should be about the match between the shape of the original experience (i.e. the covariance structure describing how it unfolds over time), not about the number of details that are recovered
\item When we retrieve memories, we cast ourselves along the full trajectory simultaneously (perhaps covering each component to a different degree), not to a single point (moment)
\item Predictions:
\begin{itemize}
    \item What makes an effective summary?  One that efficiently captures the overall shape of the experience.
    \item How do we communicate our experiences to others?  Describe the path through conceptual space.
    \item How do we communicate our experiences to our ``future selves?''  Describe the path through conceptual space.
    \item Each of these notions is fundamentally about the full trajectory, not about a single point.  Each moment derives meaning only in the context of the rest of experience.
\end{itemize}
\item What can list learning tell us?  We get fine control over  contextual representations at a very limited set of timescales.  We can study the detailed impacts of manipulations on memory.
\item What can't list learning tell us?  We can ask whether people precisely reproduce their prior experiences.  But we can't know whether they recover the ``gist'' of those experiences, since random lists do not have meaningful gists.
\item Other thoughts (from previous brainstorms):
    \begin{itemize}
    \item Reactivated neural representations that seem like noise should actually have predictive power for later memories if they get incorporated into the mental state at the time of remembering.
    \item List learning experiments and other experiments with discrete trials are not well equipped to study time spreading because they effectively impose event boundaries around and between every stimulus presentation.  Rather, we need continuous naturalistic recall experiments to more accurately reflect how our brains operate in the real world.
    \item say something about slow drift vs. reinstatement (cite MannEtal11 and FolkEtal18)
    \item something about schema learning (related to semantic memory?) and how memories are compressed (Chen et al 2017, Baldassano et al 2017, 2018) and transmitted to others
    \end{itemize}
\end{itemize}

Our brains exhibit autocorrelated activity.  In the extreme this is trivially true; our brains are biophysical systems that cannot change infinitely quickly, so neural activity patterns are necessarily similar (to some extent) from one moment to the next.  However, autocorrelated activity patterns \textit{at the time scales we typically examine in laboratory memory studies} are far stronger and persist for far longer than is strictly necessary from a biophysical standpoint.  For example, transient changes in neuronal membrane potentials during action potentials can last just a few milliseconds~\citep{HodgHuxl52} and even local field potentials that reflect the synchronized activities of tens of thousands of neurons can exhibit millisecond timescale changes~\citep{Frie05, Buzs06}.  Nevertheless, our brain activity pattern can remain reliably autocorrelated at timescales of seconds or longer-- and these long timescale autocorrelations predict how our ongoing experiences are remembered later~\citep[e.g.][]{MannEtal07, MannEtal11, HowaEtal12, FolkEtal18}.

\section*{Implications}
\subsection*{Extensions to spatial and semantic memory}
Just as we mentally spread ourselves across a distribution of times, we can perform similar mental operations when we conceptualize and retrieve other types of information.  For example, consider how we use pattern classifiers to connect neural activity patterns to specific semantic~\citep{semantic decoding studies} or spatial~\citep{spatial decoding studies} content being mentally manipulated.  The prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.  Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states.  Traditional approaches then ``clean up'' this distribution over possible mental states by taking its expected or maximum value.  However, following the ``quantum wave function'' logic, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.

\section*{Concluding remarks}

\bibliographystyle{apa}
\bibliography{memlab}
\end{document}

### 2018-10-01 main.tex
added CDL-bibliography as a sub-module
backing up progress

@@
 \usepackage{pxfonts}
+\usepackage{graphicx}
 
@@
 
-The notion of context was fundamental to \citeauthor{Tulv83}'s characterization of episodic recall, and continues to be a primary feature of most current theories of episodic memory~\citep[e.g.\ for review see][]{Kaha12}.
+The notion of context was fundamental to \citeauthor{Tulv83}'s characterization of episodic recall, and continues to be a primary feature of most current theories of episodic memory~\citep[e.g., for review see][]{Kaha12}.
 Whereas the \textit{content} of a memory provides specific information about the episode itself, contextual components of a memory help to frame the episode within the rememberer's broader experience.
-Defining context precisely (e.g.\ as distinct from content) is an ongoing challenge for our field, but theorists generally describe context as reflecting a mental combination of external cues (where you are, who you are with, background sensory information, etc.) and internal cues (thoughts, feelings, emotions, etc.) that uniquely define each moment~\citep[e.g.\ see review by][]{MannEtal15}.
+Defining context precisely (e.g., as distinct from content) is an ongoing challenge for our field, but theorists generally describe context as reflecting a mental combination of external cues (where you are, who you are with, background sensory information, etc.) and internal cues (thoughts, feelings, emotions, etc.) that uniquely define each moment~\citep[e.g., see review by][]{MannEtal15}.
 Reactivating the mental representation of the context associated with a given episode is what can give us the feeling of some ``part of ourselves'' re-experiencing the past.
 
-One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, FolkEtal18}.
+One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
 One intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
 Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.
-This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12}.
+This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12, AlyEtal18}.
 Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.
 This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11}.
-These ideas have also been formalized in a family of theoretical models by Marc Howard and colleagues~\citep{HowaKaha02, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14}.
+Recent work also suggests that the lateral entorhinal cortex plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}.
+These ideas have also been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14}.
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
@@
 In particular, both the theory and neural measures described in the preceding paragraphs characterize episodic recall mathematically as reflecting the reactivation of a weighted blend of thoughts from a \textit{range} of previously experienced times-- not to any one \textit{specific} time.
-This is not simply a matter of imprecision in mental time travel (i.e.\ a reflection of the low effective resolution with which we can revisit specific moments from our past).
+This is not simply a matter of imprecision in mental time travel (i.e., a reflection of the low effective resolution with which we can revisit specific moments from our past).
 Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences.
@@
 \section*{The temporal dynamics of ongoing experience and how we remember it}
-Modern context-based theories of episodic memory posit that the current state of mental context serves to determine which information from our past may be relevant to us and therefore reactivated~\citep[e.g.\ ][]{PolyEtal09}.
-Although these theories were largely developed and tested in the domain of list-learning paradigms, generally similar principles might be expected to play out in real world memory as well.
-For example, prior experiences with shared or overlapping contextual properties (e.g.\ other experiences that occurred in similar spatial or social settings, shared similar goals, etc.) could be leveraged to form situation models that help guide behaviors and expectations according to the current perceived context~\citep{RangRitc12}.
+Modern context-based theories of episodic memory posit that the current state of mental context serves to determine which information from our past may be relevant to us and therefore reactivated~\citep[e.g., ][]{PolyEtal09}.
+Although these theories were largely developed and tested in the domain of list-learning paradigms, conceptually similar principles might also underlie real-world memory.
+For example, prior experiences with shared or overlapping contextual properties (e.g., other experiences that occurred in similar spatial or social settings, shared similar goals, etc.) could be leveraged to form schemas or situation models that help guide behaviors and expectations according to the current perceived context~\citep{RangRitc12, BaldEtal18, AlyEtal18}.
 
-Unlike purely random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g.\ shifting from the quiet peace of writing a paper to the different peace of consoling a distraught child).
+Unlike random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g., shifting from the quiet peace of writing a paper to the different peace of consoling a distraught child).
 A number of studies (using non-random lists and more naturalistic stimuli) have found that the way we remember past experiences can be shaped by these event boundaries.
-For example, controlling for elapsed time, memory is impaired for information that occurred prior to a change in the current event or situation~\citep[e.g.\ ][]{RadvCope06, SwalEtal09, SwalEtal11, EzzyDava11, MannEtal16}.
+For example, controlling for elapsed time, memory is impaired for information that occurred prior to a change in the current event or situation~\citep[e.g., ][]{RadvCope06, SwalEtal09, SwalEtal11, EzzyDava11, MannEtal16}.
 The temporal dynamics of contextual change also underlie how we judge the amount of time elapsed since a prior reference point~\citep{BlocReed78, SahaSmit14}.
-These findings suggest that the rapid contextual or situational changes that define these event boundaries shape how we experience and remember, similarly to how spatial boundaries~\citep[e.g.\ environmental barriers;][]{McKeBuzs16, BrunEtal18} shape how we experience and remember spatial environments and layouts, or how conceptual boundaries~\citep[e.g.\ distinctions between semantic categories;][]{BrunEtal18} shape how we experience and remember conceptual information.
+These findings suggest that the rapid contextual or situational changes that define these event boundaries shape how we experience and remember, similarly to how spatial boundaries~\citep[e.g., environmental barriers;][]{McKeBuzs16, BrunEtal18} shape how we experience and remember spatial environments and layouts, or how conceptual boundaries~\citep[e.g., distinctions between semantic categories;][]{BrunEtal18} shape how we experience and remember conceptual information.
 
@@
 This rich tapestry of interacting moments and experiences forms the scaffolding for interpreting our new experiences and for retrieving information concerning our past.
-According to this view, the reactivation of this rich tapestry of contextual details enfolding the remembered event (i.e.\ how the event relates to the rest of ones experiences across timescales) is even more central to our subjective experience than the specific details of what occurred during that event!
-This begs the question: what does it mean to \textit{remember} our past experiences?
+According to this view, the reactivation of this rich tapestry of contextual details enfolding the remembered event (i.e., how the event relates to the rest of ones experiences across timescales) is even more central to our subjective experience than the specific details of what occurred during that event!
+This begs the question: which thoughts, from which times, do we reactivate when we remember our past experiences?
 
-\section*{The shape of remembered experience}
+\section*{The shapes of remembered experiences}
+In a recent study~\citep{HeusEtal18c} my lab examined functional neuroimaging and behavioral data collected by \cite{ChenEtal17} as participants viewed and then verbally recalled an episode of a television show.
+We used topic models~\citep{BleiEtal03} to characterize the moment-by-moment content of the episode and of participants' recalls.
+Each moment was quantified via a \textit{topic vector} that reflected the mix of topics (themes) the model identified in that moment.
+In this way, the sequence of topic vectors over the course of the episode traces out a \textit{topic trajectory} through a high-dimensional feature space.
+The sequence of topic vectors for a given participant's recalls traces out an analogous trajectory.
+This characterization provides a geometric framework for comparing the \textit{shapes} that reflect how naturalistic experiences unfold over time, to the shapes of how we remember those experiences later (Fig.~\ref{fig:trajectories}).
+One can then use the geometric transformations needed to map an experience's trajectory onto the trajectory of its later remembering to characterize the quality and content of memory.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.8\textwidth]{figs/trajectory}
+\caption{\textbf{Topic trajectories of a video and its memory.}  \textbf{A.} The topic trajectory taken by a 50 minute television episode, projected onto two dimensions.
+Each dot indicates an event identified using a Hidden Markov Model; the dot colors denote the order of the events (early events are in red; later events are in blue), and the connecting lines indicate the transitions between successive events.
+\textbf{B.} The average two-dimensional trajectory captured by 17 participants' recalls of the episode, with the same format and coloring as the trajectory in Panel A.
+The arrows reflect the directions of the average transitions taken by participants whose trajectories crossed that part of topic space; blue denotes reliable agreement across participants.
+Figure adapted from \cite{HeusEtal18c}.}
+\label{fig:trajectories}
+\end{figure}
+
+defines a distinction between the gist
 
@@
 \item Not about the number of details-- it's about capturing the overall ``shape'' of past experience
-\item Quantifying recall should be about the match between the shape of the original experience (i.e.
-the covariance structure describing how it unfolds over time), not about the number of details that are recovered
+\item Quantifying recall should be about the match between the shape of the original experience (i.e., the covariance structure describing how it unfolds over time), not about the number of details that are recovered
 \item When we retrieve memories, we cast ourselves along the full trajectory simultaneously (perhaps covering each component to a different degree), not to a single point (moment)
@@
 \bibliographystyle{apa}
-\bibliography{memlab}
+\bibliography{CDL-bibliography/memlab}
 \end{document}

### 2018-12-07 main.tex
minor code updates, working on text

@@
 
-\title{Episodic memory: mental time travel or a quantum wave function?}
+\title{Episodic memory: mental time travel or a quantum `memory wave' function?}
 \author{Jeremy R. Manning\\Dartmouth College, Hanover, NH\\jeremy.r.manning@dartmouth.edu}
@@
 One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
-One intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
+An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
 Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.
@@
 Recent work also suggests that the lateral entorhinal cortex plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}.
-These ideas have also been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14}.
+These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14}.
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
@@
 The notion of thinking of any one particular moment from the past is incompatible with this multiple timescales view, in that our thoughts and brain activity patterns reflect information from a range of moments, even when we are not specifically engaged in remembering.
-If what we consider to be a ``moment'' or ``now'' is guided by our neural representations of our ongoing experiences, then this implies we are continually experiencing \textit{many} ``nows.''  If our thoughts are continually spread across a range of times, the notion of mentally revisiting any particular time loses its meaning.
+If what we consider to be a ``moment'' or ``now'' is guided by our neural representations of our ongoing experiences, then this implies we are continually experiencing \textit{many} ``nows.''  If our thoughts are continually spread across a range of times, the notion of mentally revisiting any particular moment loses its meaning.
 
-Where the mental time travel description breaks down most notably is when we consider scenarios in which the particular blend of timepoints reflected in one's current thoughts do not come from temporally contiguous events.
+Where the mental time travel framing breaks down most notably is when we consider scenarios in which the particular blend of timepoints reflected in one's current thoughts do not come from temporally contiguous events.
 For example, understanding how a new experience fits in with our broader understanding might require integrating information gleaned from many experiences separated in time, each with their own contextual attributes.
@@
 
-Unlike random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g., shifting from the quiet peace of writing a paper to the different peace of consoling a distraught child).
+Unlike random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g., shifting from the quiet peace of working on a manuscript in the early morning to the different peace of helping a newly awoken child get ready for school).
 A number of studies (using non-random lists and more naturalistic stimuli) have found that the way we remember past experiences can be shaped by these event boundaries.
@@
 Event boundaries are one example of the broader temporal covariance structure that is characteristic of ongoing experience.
-Whereas classic approaches to studying memory encourage researchers to treat each moment as separable from the rest of experience, context-based theories of memory (including situation models and event-based models) posit that each moment derives meaning through its deep ties to other experiences.
+Whereas classic approaches to studying memory (e.g., list-learning paradigms and other trial-based experiments) encourage researchers to treat each moment, trial, or stimulus as separable from the rest of experience, context-based theories of memory (including situation models and event-based models) posit that each moment derives meaning through its deep ties to other experiences.
 This rich tapestry of interacting moments and experiences forms the scaffolding for interpreting our new experiences and for retrieving information concerning our past.
@@
 One can then use the geometric transformations needed to map an experience's trajectory onto the trajectory of its later remembering to characterize the quality and content of memory.
+
+\textbf{JRM STOPPED HERE}
 

### 2018-12-18 main.tex
checking in changes
combined movie and fr reinstatement figs

@@
 
+\section*{Shared contexts of temporally proximal and remote events}
+
+\begin{figure}[tp] \centering \includegraphics[width=0.75\textwidth]{figs/pres_rec_corrmats} \caption{\textbf{Temporal correlation matrices.}  Each panel displays the timepoint-by-timepoint correlation matrix for presented stimuli (original experience, left panels) and participants' memories of those stimuli (recalled experience, right panels).
+}
+\label{fig:corrmats}
+\end{figure}
+
+\textbf{JRM NOTE: add section on the temporal structure of list-learning vs. naturalistic stimulus, and on the temporal structure of recalls.}
+
 \section*{The shapes of remembered experiences}

### 2018-12-19 main.tex
added new figures to paper
working on results section

@@
 
-\section*{Shared contexts of temporally proximal and remote events}
+\section*{How do our experiences and memories unfold over time?}
+Although real-world experiences can exhibit rich temporal structures, it can be illustrative to consider the transformation of how we experience much simpler and less naturalistic events (e.g.
+studying random word lists) versus how we later remember those experiences.
+Figure~\ref{fig:corrmats}A displays an average temporal correlation matrix describing the similarity structure of 1072 random word lists, each comprising 16 words.
+I applied a topic model (trained on a curated corpus of Wikipedia articles) to yield a 100-dimensional embedding of each presented word~\citep{BleiEtal03}.
+The left panel displays the average correlations between the topic vectors of each pair of presented words as a function of their presentation positions.
+As shown, nearly all of the correlation ``mass'' is along the diagonal, and the off-diagonal entries are very close to zero.
+Because the word orders were randomized, the temporal correlation matrix reflects that there are no consistent similarity patterns between any given pair of timepoints.
+The right panel of Figure~\ref{fig:corrmats}A displays the average temporal correlation matrices describing participants' \textit{recalls} of those same lists.
+Although the recall matrix exhibits the strongest correlations along its diagonal, some correlation mass bleeds into nearby recalls, reflecting (in part) participants' tendancies to recall semantically related words successively.\footnote{Note that participants recalled different numbers of words from each studied list; I used linear interpolation to resample each recall correlation matrix that preserved the size of the presentation correlation matrix.
+This process resulted in some additional blurring along the diagonal.
+In other words, the blurring along the diagonal reflects two phenomena: participants' tendancies to semanticly cluster their recalls, and an artifact of the temporal resampling.}  Differences in the temporal correlation structure between the original lists of words and the orders of participants' recalls illustrate one way that our memory systems can reorganize our experiences such that we recall them in a different way than we originally experienced them.
 
 \begin{figure}[tp] \centering \includegraphics[width=0.75\textwidth]{figs/pres_rec_corrmats} \caption{\textbf{Temporal correlation matrices.}  Each panel displays the timepoint-by-timepoint correlation matrix for presented stimuli (original experience, left panels) and participants' memories of those stimuli (recalled experience, right panels).
-}
+\textbf{A.
+Presentation and recall of random word lists.}  The panels reflect temporal correlation matrices between the topic vectors of presented words (left) and recalled words (right) over 1072 randomly ordered 16-word lists.
+\textbf{B.
+Presentation and recall of categorized word lists.}  Same format as Panel A, but here participants have studied and recalled 480 categorized word lists, where the word presentations were ordered by category.
+\textbf{C.
+Presentation and recall of a television episode.} Each moment of a television episode (left) and participants' retellings of what happened in the episode (right) are characterized by topic vectors that reflect the moments' semantic contents~\citep{HeusEtal18c}.}
 \label{fig:corrmats}
+\end{figure}
+
+Whereas random word lists exhibit no stable off-diagonal correlations, one can induce temporal correlations in random word lists by manipulating the presentation orders of words on each list.
+The left panel of Figure~\ref{fig:corrmats}B displays the average temporal correlation matrix exhibited by 240 16-word lists, each of whose items are drawn from 4 randomly selected categories~\citep{ZimaEtal18}.
+The words were then presented to participants in categorized order, creating a block-diagonal correlation structure.
+The right panel displays the temporal correlation matrix of the participants' recalls.
+Despite
+
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Mental time travel in random word lists and video.}  }
+  % colors: individual participants
+  % black: average across participants
+  % gray: autocorrelation with topic vector at timepoint
+  % circles: local maxima in the across-participants average (i.e., movie events that share structure with the given timepoint that is distinct from other temporally proximal events)
+\label{fig:reinstatement}
 \end{figure}

### 2018-12-19 main.tex
removed categorized word list results

@@
 studying random word lists) versus how we later remember those experiences.
-Figure~\ref{fig:corrmats}A displays an average temporal correlation matrix describing the similarity structure of 1072 random word lists, each comprising 16 words.
+Figure~\ref{fig:corrmats}A displays an average temporal correlation matrix describing the similarity structure of 1072 random word lists, each comprising 16 words~\citep{ZimaEtal18}.
 I applied a topic model (trained on a curated corpus of Wikipedia articles) to yield a 100-dimensional embedding of each presented word~\citep{BleiEtal03}.
@@
 \textbf{A.
-Presentation and recall of random word lists.}  The panels reflect temporal correlation matrices between the topic vectors of presented words (left) and recalled words (right) over 1072 randomly ordered 16-word lists.
+Presentation and recall of random word lists.}  The panels reflect temporal correlation matrices between the topic vectors of presented words (left) and recalled words (right) over 1072 randomly ordered 16-word lists~\citep{ZimaEtal18}.
 \textbf{B.
-Presentation and recall of categorized word lists.}  Same format as Panel A, but here participants have studied and recalled 480 categorized word lists, where the word presentations were ordered by category.
-\textbf{C.
 Presentation and recall of a television episode.} Each moment of a television episode (left) and participants' retellings of what happened in the episode (right) are characterized by topic vectors that reflect the moments' semantic contents~\citep{HeusEtal18c}.}
@@
 \end{figure}
-
-Whereas random word lists exhibit no stable off-diagonal correlations, one can induce temporal correlations in random word lists by manipulating the presentation orders of words on each list.
-The left panel of Figure~\ref{fig:corrmats}B displays the average temporal correlation matrix exhibited by 240 16-word lists, each of whose items are drawn from 4 randomly selected categories~\citep{ZimaEtal18}.
-The words were then presented to participants in categorized order, creating a block-diagonal correlation structure.
-The right panel displays the temporal correlation matrix of the participants' recalls.
-Despite
 
@@
 Just as we mentally spread ourselves across a distribution of times, we can perform similar mental operations when we conceptualize and retrieve other types of information.
-For example, consider how we use pattern classifiers to connect neural activity patterns to specific semantic~\citep{semantic decoding studies} or spatial~\citep{spatial decoding studies} content being mentally manipulated.
+For example, consider how we use pattern classifiers to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12} or spatial~\citep{MillEtal13} content being mentally manipulated.
 The prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.

### 2018-12-31 code/notebooks/reinstatement.py
explorations in VS Code

@@ -1,0 +1,833 @@
+#%%
+
+import numpy as np
+import seaborn as sns
+import pandas as pd
+import matplotlib.pyplot as plt
+import matplotlib.patches as patches
+import matplotlib as mpl
+import scipy.signal as sgn
+import pickle
+import os
+import warnings
+import quail as q
+import hypertools as hyp
+import future
+from pathlib import Path
+import urllib
+from glob import glob
+from scipy.spatial.distance import cdist, pdist, squareform
+from scipy.interpolate import pchip
+from scipy.signal import find_peaks as localmax
+from scipy.stats import ttest_ind as ttest
+
+#%% 
+
+# ## Set path, load sherlock video and recall trajectories, load free recall data
+
+# In[3]:
+x = os.getcwd()
+
+rootdir = str(Path(notebook_path()).parent.parent.parent)
+datadir = os.path.join(rootdir, 'data')
+figdir = os.path.join(rootdir, 'data', 'figs')
+if not os.path.exists(figdir):
+    os.makedirs(figdir)
+video_model, recall_models = np.load(os.path.join(datadir, 'processed', 'models_t100_v50_r10.npy'))
+fr_random_data = q.load_egg(os.path.join(datadir, 'free recall', 'exp1.egg'))
+fr_cat_data = q.load_egg(os.path.join(datadir, 'free recall', 'exp2.egg'))
+
+
+
+
+# ## Plot figure
+
+# In[4]:
+
+
+h = sns.heatmap(pd.DataFrame(recall_models[0]).T.corr())
+
+
+# In[5]:
+
+
+def parse_diag(M):
+    n = M.shape[1]
+    x = np.nan * np.zeros([M.shape[0], 2 * n + 1])
+    
+    for i in np.arange(n):
+        x[i, (n - i):n] = M[i, 0:i]
+        x[i, n] = M[i, i]
+        x[i, (n+1):(n + n - i)] = M[i, (i+1):]
+    
+    return pd.DataFrame(x, columns=np.arange(-n, n+1))
+
+
+# In[6]:
+
+
+M = pd.DataFrame(recall_models[0]).T.corr().values
+np.fill_diagonal(M, np.nan)
+x = parse_diag(M)
+
+
+# In[7]:
+
+
+h = sns.heatmap(x)
+
+
+# In[295]:
+
+
+def plot_ribbon(x, color='k', ribbon_alpha=0.2):
+    mean = x.mean()
[…]

### 2018-12-31 main.tex
added a bunch of thoughts on reorganizing, and questions to address

@@
 This begs the question: which thoughts, from which times, do we reactivate when we remember our past experiences?
+
+\section*{BRAINSTORMING OUTLINE OF THE REST OF THE PAPER}
+\begin{enumerate}
+  \item Show video correlation matrix + recall correlation matrix.
+Key idea: off-diagonal mass is introduced (primarily) by how people \textit{remember} the movie, not due to the contents of the movie itself (evidence: correlation matrices).
+Further, this is not simply a matter of imprecision in how people recall, and/or our approach's ability to characterize recalled events (evidence: reinstatement figure showing which scenes are recalled similarly).
+This sort of integration is how we ascribe meaning to events that unfold over disjointed extended timescales.
+  \item In theory this could also happen during \textit{any} sort of experience.
+For example, when people study random word lists, they often re-order their recalls so that semantically related words are closer in the recall sequence than they were in the presented sequence.
+This can be captured to an extent by semantic clustering measures.
+However, it's difficult to examine this phenomenon in more depth because word lists necessarily lack the richness of more naturalistic stimuli.
+There are no ``deeper ties'' between study events (i.e., individual words), so the sorts of integration operations that allow us to make sense of real-world experiences aren't given the same freedom of expression in list-learning experiments.
+Evidence: presentation/recall correlation matrices during random free recall don't show the same sort of off-diagonal increase.
+Claim: mental time travel and quantum wave reinstatement look the same for random word list remembering, but very different for real-world remembering.
+  \item The specific ties between unfolding events forms a scaffolding for understanding how those events are related.
+This can be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces (although see Flatland Fallacy paper, with HyperTools paper as a potential counterpoint).
+This reveals another key property of how we remember: although there are substantial individual differences in the specific words people use to describe a shared experience, and how much detail they get into, the overall ``shapes'' of those recalled experience trajectories are highly similar across people.
+This has potentially exciting implications for how we communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.
+It also has implications for how we learn by analogy: perhaps events that share schemas (cite Baldassano et al) unfold in similar ways (e.g.
+have similar shapes).
+Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).
+    \item These ideas also have implications for other forms of learning (e.g.
+spatial learning) and neural decoding.
+When we attempt to characterize thoughts (e.g.
+training a classifier to distinguish between states), we tend to train decoders to estimate the ``one'' state that the brain is in.
+However, perhaps we should instead consider that our brains are contemplating, or holding in our thoughts, many possible states simultaneously.
+Therefore we should consider the full distribution of states (rather than only considering its maximum) as a primary metric of characterizing mental function.
+(Probably people have proposed this-- e.g.
+representing the posterior and using it to compute...need citations.)
+
+    \end{enumerate}
+
+\textbf{KEY QUESTIONS}:
+\begin{enumerate}
+  \item Is this more of a ``review'' paper, or should I highlight the new analyses more?
+  \item Who is this paper alienating and how can the narrative be reframed in a more inclusive way?
+  \item Which ideas are ``new'' vs. restatements of old ideas?
+  \item What are the key testable predictions to highlight, and how can they be communicated most clearly and effectively
+  \item What's an appropriate venue for this?
+\end{enumerate}
 

### 2019-01-02 boneyard.tex
working on refactor of results section and logical flow

@@
+Further, this finding would appear to be at odds with classic context-based accounts of memory, to the extent that these models proport to explain memory associations between thematically similar experiences.
+Specifically, if thematic overlap between temporally distant moments of the episode is rare, what
+
 \item Event boundaries: these shape the distribution because they sharply manipulate similarity as a function of time (whereas within an event similarity varies smoothly with time)

### 2019-01-02 main.tex
working on sherlock reinstatement and corrmat figs
working on refactor of results section and logical flow

@@
 
+\subsection*{How is our past replayed when we remember?}
+Our ongoing experiences can cue us to dredge up information about our past.
+This can happen explicitly (e.g., during a conversation about a prior event) or implicitly (e.g., when something happening now reminds you of what happened before).
+Context-based theories of memory posit that the particular memories that our ongoing experiences cue will be related~\citep[contextually, semantically, functionally, etc.; e.g., ][]{PolyEtal09}.
+One potential challenge to studying how we remember the past relates to detangling which links between our experiences are specifically driven by our memory systems versus which are imposed ``externally'' through the similarity structure of those experiences.
+For example, consider your daily commute into work.
+If you were to verbally recount a specific day's commute later, one might expect some aspects of your description to apply to other days' commutes as well (e.g., other commutes where you took a similar route or travelled via similar modes of transportation, listened to similar music along the way, travelled at a similar time, etc.).
+But to what extent could one hope to separate out descriptive matches driven solely by similarities in those commuting experiences, versus true reactivations (in memory) of specific memories of those other prior commutes?
+
+Once insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \citep{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
+The authors used topic models~\citep{BleiEtal03} applied to human annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.
+When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change.
+This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \citep{HeusEtal18c} perform detailed analyses of these putative events and how they are remembered.
+Notably, the off-diagonal entries of the temporal correlation matrix (which reflects content similarity between temporally distant events in the episode) are nearly all close to 0.
+This is surprising, because the episode itself comprises a murder mystery with recurring motifes that must be associated in the viewer's mind in order to make sense of the episode.
+Nonetheless, when considering solely the semantic content of each moment of the episode, those long timescale associations are nearly entirely absent.
+To the extent that this finding extends to real-world experiences, it comports with Heraclitus' notion that one cannot step in the same river twice~\citep{Hera}.
+
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/pres_rec_corrmats_sherlock} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal18c}, where full details may be found.
+\textbf{A.
+Temporal correlations exhibited by the content of the television episode.}  Each moment of an episode of the BBC television show \textit{Sherlock} has been characterized using a topic model~\citep{BleiEtal03} fit to detailed human-generated annotations of each scene cut.
+The correlations between the topic vectors of a given pair of timepoints reflects the similarity in content between those timepoints.
+\textbf{B.
+Temporal correlations exhibited by the (average) content of participants' recalls of the episode.} The topic model used to characterize the episode in Panel A has been applied to each sentence, from each participants' recall transcript.
+The temporal correlation matrices describing each participants' recalls were re-sampled using cubic interpolation to have 100 rows and columns prior to averaging the resampled matrices across participants.
+\textbf{C.
+Temporal correlations in the content of individual participants' recalls of the episode.} Each sub-panel displays the temporal correlation matrix for one experimental particpant.}
+\label{fig:corrmats}
+\end{figure}
+
+Although the specific content of each moment of the episode is largely unique, participants' subjective experiences of the episode entail weaving together those temporally separated moments into a cohesive narrative.
+This is reflected in how people verbally recount the episode later.
+If we examine the temporal autocorrelation matrices of participants' recalls of the television episode (characterized using the same topic model used to capture the content of the episode; Figs.~\ref{fig:corrmats}B, C), we see a different structure than that of the episode itself.
+Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off diagonal correlations between temporally distant events are substantially stronger in participants' recalls than in the original episode.
+This indicates that the way participants \textit{recount} temporally separated events imposes additional similarity on those events.
+One might then ask \textbf{JRM STOPPED HERE}
+%where i was going with this: is that structure meaningful?
+one can ask which specific events are described similarly (reinstatement fig).
+argument: yes, the structure is meaningful-- example: the way people describe the particular event (sherlock surmising a new murder has happened between bantering with Watson) is similar to how participants describe the other murders and other sherlock/watson banterings
+
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Mental time travel in random word lists and video.}  }
+  % colors: individual participants
+  % black: average across participants
+  % gray: autocorrelation with topic vector at timepoint
+  % circles: local maxima in the across-participants average (i.e., movie events that share structure with the given timepoint that is distinct from other temporally proximal events)
+\label{fig:reinstatement}
+\end{figure}
+
 \section*{BRAINSTORMING OUTLINE OF THE REST OF THE PAPER}
@@
 representing the posterior and using it to compute...need citations.)
-
     \end{enumerate}
@@
 In other words, the blurring along the diagonal reflects two phenomena: participants' tendancies to semanticly cluster their recalls, and an artifact of the temporal resampling.}  Differences in the temporal correlation structure between the original lists of words and the orders of participants' recalls illustrate one way that our memory systems can reorganize our experiences such that we recall them in a different way than we originally experienced them.
-
-\begin{figure}[tp] \centering \includegraphics[width=0.75\textwidth]{figs/pres_rec_corrmats} \caption{\textbf{Temporal correlation matrices.}  Each panel displays the timepoint-by-timepoint correlation matrix for presented stimuli (original experience, left panels) and participants' memories of those stimuli (recalled experience, right panels).
-\textbf{A.
-Presentation and recall of random word lists.}  The panels reflect temporal correlation matrices between the topic vectors of presented words (left) and recalled words (right) over 1072 randomly ordered 16-word lists~\citep{ZimaEtal18}.
-\textbf{B.
-Presentation and recall of a television episode.} Each moment of a television episode (left) and participants' retellings of what happened in the episode (right) are characterized by topic vectors that reflect the moments' semantic contents~\citep{HeusEtal18c}.}
-\label{fig:corrmats}
-\end{figure}
-
-\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Mental time travel in random word lists and video.}  }
-  % colors: individual participants
-  % black: average across participants
-  % gray: autocorrelation with topic vector at timepoint
-  % circles: local maxima in the across-participants average (i.e., movie events that share structure with the given timepoint that is distinct from other temporally proximal events)
-\label{fig:reinstatement}
-\end{figure}
 

### 2019-01-03 main.tex
working on reinstatement description-- Fig 2

@@
 Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off diagonal correlations between temporally distant events are substantially stronger in participants' recalls than in the original episode.
-This indicates that the way participants \textit{recount} temporally separated events imposes additional similarity on those events.
-One might then ask \textbf{JRM STOPPED HERE}
+This indicates that the way participants \textit{recount} temporally separated events imposes additional similarity on those events, beyond what was present in the original events themselves.
+Is this simply a matter of imprecision in recall, either with respect to participants' recalls themselves, or with respect to the quality with which they are characterized via semantic feature vectors?
+Or might there be a functional role to that additional imposed similarity structure?
+
+To examine which specific events are described (during recall) as more similar than what their original experiences (as characterized by semantic feature vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.
+\textbf{JRM STOPPED HERE...}
+
 %where i was going with this: is that structure meaningful?
@@
 
-\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Mental time travel in random word lists and video.}  }
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Content drift in a television episode and participants' recalls of that episode.
+A.
+Autocorrelations in the episode's content and in the content of participants' recalls.}  The black line denotes the average correlations between the topic vectors from pairs of moments in the episode as a function of their temporal separation (times are relative to the full length of the episode).
+Each colored line denotes an analogous autocorrelation function, but for each individual participant's recalls of the episode (times are relative to the full length of each participant's recalls).
+The error ribbons denote 95\% confidence intervals, taken across all moments in the episode (or in participants' recalls of the episode).
+\textbf{B.
+Recall density functions for one recalled moment.} Each curve reflects the correlations between the topic vector at each (relative) moment of the episode (or participants' recalls of the episode), and the topic vector of an event beginning approximately $\frac{1}{3}$ through the episode (relative time: 0.328).
+The gray curve denotes these correlations for the topic vectors derived from annotations of the episode itself.
+\textbf{JRM STOPPED HERE...}}
   % colors: individual participants

### 2019-01-04 main.tex
working through revised outline

@@
 Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off diagonal correlations between temporally distant events are substantially stronger in participants' recalls than in the original episode.
-This indicates that the way participants \textit{recount} temporally separated events imposes additional similarity on those events, beyond what was present in the original events themselves.
-Is this simply a matter of imprecision in recall, either with respect to participants' recalls themselves, or with respect to the quality with which they are characterized via semantic feature vectors?
+Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode and participants' recalls (Fig.~\ref{fig:reinstatement}A).
+Whereas the autocorrelations for the original episode rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around 0.2.
+These analyses indicate that the way participants \textit{recount} temporally separated events imposes additional similarity on those events, beyond what was present in the original events themselves.
+One can then ask: is this simply a matter of imprecision in recall, either with respect to participants' recalls themselves, or with respect to the quality with which they are characterized via semantic feature vectors?
 Or might there be a functional role to that additional imposed similarity structure?
@@
 To examine which specific events are described (during recall) as more similar than what their original experiences (as characterized by semantic feature vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.
-\textbf{JRM STOPPED HERE...}
-
-%where i was going with this: is that structure meaningful?
-one can ask which specific events are described similarly (reinstatement fig).
-argument: yes, the structure is meaningful-- example: the way people describe the particular event (sherlock surmising a new murder has happened between bantering with Watson) is similar to how participants describe the other murders and other sherlock/watson banterings
+Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were described in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
+The example event explored in the figure comprises a scene where Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another victim has been murdered.
+The language used by human annotators describing the moment-by-moment contents of the scene used a similar mix of themes only for moments in the episode that were temporally proximal to the event in question (gray line).
+However, the way participants \textit{described} what happened when they recounted the episode later revealed a much different pattern.
+Specifically, their language mirrored that used when they were recalling other murders, and other moments when Holmes and Watson were bantering (Fig.~\ref{fig:reinstatment}C).
+This suggests that these events were associated in participants' memories, perhaps contributing to the understanding that those events were linked in the episode's narrative.
 
@@
 The gray curve denotes these correlations for the topic vectors derived from annotations of the episode itself.
-\textbf{JRM STOPPED HERE...}}
+The black line denotes the average correlations across all participants' recalls.
+Local maxima are marked with colored dots.
+Local maxima closer than 2.5\% of the total recall duration to a higher local maximum were excluded, as were maxima with peaks with correlation coefficients lower than 0.1 (arbitrarily chosen; these ).
+The colored lines denote the correlations for each individual participant's recalls.
+\textbf{C.
+Local maxima in the average recall density function.}  Representative screen captures from the scenes at each local maxima in Panel B are displayed (dot colors match those in Panel B).
+The text in each screen capture displays a brief summary of what happened in each scene.}
   % colors: individual participants
@@
 
+\textbf{JRM STOPPED HERE}
+
 \section*{BRAINSTORMING OUTLINE OF THE REST OF THE PAPER}
 \begin{enumerate}
-  \item Show video correlation matrix + recall correlation matrix.
+  \item \textit{Show video correlation matrix + recall correlation matrix.
 Key idea: off-diagonal mass is introduced (primarily) by how people \textit{remember} the movie, not due to the contents of the movie itself (evidence: correlation matrices).
 Further, this is not simply a matter of imprecision in how people recall, and/or our approach's ability to characterize recalled events (evidence: reinstatement figure showing which scenes are recalled similarly).
-This sort of integration is how we ascribe meaning to events that unfold over disjointed extended timescales.
+This sort of integration is how we ascribe meaning to events that unfold over disjointed extended timescales.}
   \item In theory this could also happen during \textit{any} sort of experience.

### 2019-01-07 main.tex
backing up progress on results section

@@
 However, the way participants \textit{described} what happened when they recounted the episode later revealed a much different pattern.
-Specifically, their language mirrored that used when they were recalling other murders, and other moments when Holmes and Watson were bantering (Fig.~\ref{fig:reinstatment}C).
-This suggests that these events were associated in participants' memories, perhaps contributing to the understanding that those events were linked in the episode's narrative.
+Specifically, their language mirrored that used when they were recalling other murders, and other moments when Sherlock and John were bantering (Fig.~\ref{fig:reinstatement}C).
+Further, the way they described this example event did \textit{not} mirror the way they described other semantically (but not conceptually) related events, such as events involving unrelated violence, or Sherlock and John interacting with other characters.
+Taken together, this suggests that these specific events were associated in participants' memories, perhaps contributing to (or reflecting) their understanding that those events were linked in the episode's narrative.
 
@@
 \end{figure}
+
+A rich recall density function analogous to the one displayed in Figure~\ref{fig:reinstatement}B could, in principle, be characteristic of memory for \textit{any} past event or experience.
+For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline, were functionally associated in memory.
+Nonetheless, a potential challenge to testing this hypothesis in the laboratory is that in most memory studies there is no particular reason for participants to functionally associate the stimuli-- there is no ``deeper narrative'' to make sense of.
+However, random word list learning experiments might happen to contain (for a given list) semantically related words.
+And indeed, a number of studies indicate that participants spontaneously form associations between semantically related words on random word lists, even when those related words are separated in time~\citep[e.g.,][]{WixtRohr94, MannKaha12, MannEtal12}.
+This suggests that our memory systems pick up on statistical regularities, even in ostensibly ``random'' stimulus sequences, and that we leverage those regularities when we recall our past~\citep{PolyEtal09}.
+Other work suggests that our memory systems may also leverage these regularities to parse our ongoing experiences into discrete events~\citep{SchaEtal13}.
+Given this well-developed literature on random list learning studies indicates that our memory systems leverage ``accidental'' structure in random sequences, \textbf{JRM STOPPED HERE}.
+%is it "sufficient"?
+pros and cons of lists vs.
+naturalistic...
+
+key question is whether the additional
 

### 2019-01-07 main.tex
updated bibliography again, working through results section [WIP]

@@
 Other work suggests that our memory systems may also leverage these regularities to parse our ongoing experiences into discrete events~\citep{SchaEtal13}.
-Given this well-developed literature on random list learning studies indicates that our memory systems leverage ``accidental'' structure in random sequences, \textbf{JRM STOPPED HERE}.
-%is it "sufficient"?
-pros and cons of lists vs.
-naturalistic...
+Are these examples in the literature on random list learning studies, indicating that our memory systems leverage ``accidental'' structure in random sequences, a reflection of the same fundamental principle driving the sorts of memory reactivation function displayed in Figure~\ref{fig:reinstatement}B?
 
-key question is whether the additional
+It is difficult to explicitly test whether (apparent) semantic reinstatement in word list learning, as reflected in people's tendencies to semantically cluster recalls of random word lists, reflects integration of information across the experiences of studying those words.
+However, remembering word lists does not \textit{require} integrating information across words in order to specifically gain a deep understanding of the list.
+By design, there \textit{is} no deeper meaning of random word lists.
 
@@
 This sort of integration is how we ascribe meaning to events that unfold over disjointed extended timescales.}
-  \item In theory this could also happen during \textit{any} sort of experience.
+  \item \textit{In theory this could also happen during \textit{any} sort of experience.
 For example, when people study random word lists, they often re-order their recalls so that semantically related words are closer in the recall sequence than they were in the presented sequence.
@@
 There are no ``deeper ties'' between study events (i.e., individual words), so the sorts of integration operations that allow us to make sense of real-world experiences aren't given the same freedom of expression in list-learning experiments.
-Evidence: presentation/recall correlation matrices during random free recall don't show the same sort of off-diagonal increase.
-Claim: mental time travel and quantum wave reinstatement look the same for random word list remembering, but very different for real-world remembering.
+Evidence: presentation/recall correlation matrices during random free recall don't show the same sort of off-diagonal increase.}  Claim: mental time travel and quantum wave reinstatement look the same for random word list remembering, but very different for real-world remembering.
   \item The specific ties between unfolding events forms a scaffolding for understanding how those events are related.

### 2019-01-10 boneyard.tex
backing up progress

@@
+\textbf{KEY QUESTIONS}:
+\begin{enumerate}
+  \item Is this more of a ``review'' paper, or should I highlight the new analyses more?
+  \item Who is this paper alienating and how can the narrative be reframed in a more inclusive way?
+  \item Which ideas are ``new'' vs. restatements of old ideas?
+  \item What are the key testable predictions to highlight, and how can they be communicated most clearly and effectively
+  \item What's an appropriate venue for this?
+\end{enumerate}
+
+% Next to expand on:
+% - Option 1: go into more depth about what the reinstatement functions look like for mental time travel vs.
+"quantum wave functions," under naturalistic vs.
+list-learning experiments.
+Claim: different for naturalistic experiments but similar for list-learning experiments.
+% - Option 2: get rid of the list learning stuff and transition directly to the "shape" of how experiences unfold.
+
+\section*{How do our experiences and memories unfold over time?}
+Although real-world experiences can exhibit rich temporal structures, it can be illustrative to consider the transformation of how we experience much simpler and less naturalistic events (e.g.
+studying random word lists) versus how we later remember those experiences.
+Figure~\ref{fig:corrmats}A displays an average temporal correlation matrix describing the similarity structure of 1072 random word lists, each comprising 16 words~\citep{ZimaEtal18}.
+I applied a topic model (trained on a curated corpus of Wikipedia articles) to yield a 100-dimensional embedding of each presented word~\citep{BleiEtal03}.
+The left panel displays the average correlations between the topic vectors of each pair of presented words as a function of their presentation positions.
+As shown, nearly all of the correlation ``mass'' is along the diagonal, and the off-diagonal entries are very close to zero.
+Because the word orders were randomized, the temporal correlation matrix reflects that there are no consistent similarity patterns between any given pair of timepoints.
+The right panel of Figure~\ref{fig:corrmats}A displays the average temporal correlation matrices describing participants' \textit{recalls} of those same lists.
+Although the recall matrix exhibits the strongest correlations along its diagonal, some correlation mass bleeds into nearby recalls, reflecting (in part) participants' tendancies to recall semantically related words successively.\footnote{Note that participants recalled different numbers of words from each studied list; I used linear interpolation to resample each recall correlation matrix that preserved the size of the presentation correlation matrix.
+This process resulted in some additional blurring along the diagonal.
+In other words, the blurring along the diagonal reflects two phenomena: participants' tendancies to semanticly cluster their recalls, and an artifact of the temporal resampling.}  Differences in the temporal correlation structure between the original lists of words and the orders of participants' recalls illustrate one way that our memory systems can reorganize our experiences such that we recall them in a different way than we originally experienced them.
+
 Further, this finding would appear to be at odds with classic context-based accounts of memory, to the extent that these models proport to explain memory associations between thematically similar experiences.

### 2019-01-10 main.tex
continuing work on paper, added some notes
backing up progress

@@
 By design, there \textit{is} no deeper meaning of random word lists.
+One way to characterize the meaning reflected by how an experience unfolds in time is through models that describe the geometric path that the experience takes through an appropriate representational space.
+
+\section*{The shapes of remembered experiences}
+In a recent study~\citep{HeusEtal18c} my lab examined functional neuroimaging and behavioral data collected by \cite{ChenEtal17} as participants viewed and then verbally recalled an episode of a television show.
+We used topic models~\citep{BleiEtal03} to characterize the moment-by-moment content of the episode and of participants' recalls.
+Each moment was quantified via a \textit{topic vector} that reflected the mix of topics (themes) the model identified in that moment.
+In this way, the sequence of topic vectors over the course of the episode traces out a \textit{topic trajectory} through a high-dimensional feature space.
+The sequence of topic vectors for a given participant's recalls traces out an analogous trajectory.
+This characterization provides a geometric framework for comparing the \textit{shapes} that reflect how naturalistic experiences unfold over time, to the shapes of how we remember those experiences later (Fig.~\ref{fig:trajectories}).
+One can then use the geometric transformations needed to map an experience's trajectory onto the trajectory of its later remembering to characterize the quality and content of memory.
+
+The specific ties between unfolding events forms a scaffolding for understanding how those events are related.
+In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\citep{JollChan18}.
+\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats}, \ref{fig:reinstatement}, and~\ref{fig:trajectories}, reveals another key property of how we remember.
+Although there are substantial individual differences in the specific words people use to describe a shared experience, and how much detail they get into, the overall ``shapes'' of those recalled experience trajectories are highly similar across people.
+This has potentially exciting implications for how we communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.
+It also has implications for how we learn by analogy: perhaps events that share schemas~\citep[e.g., ][]{BaldEtal18}  unfold in similar ways (e.g., have similar shapes).
+Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).
+
+% analysis idea: matrix of reinstatement functions for each timepoint-- e.g.
+considering each timepoint in turn (rows), how much are participants (on average) using language that mirrors how they talk about each other timepoint (columns).
+there should be a strong diagonal, with some blockiness, and possibly some disjoint groups showing how events cluster.
 
@@
 There are no ``deeper ties'' between study events (i.e., individual words), so the sorts of integration operations that allow us to make sense of real-world experiences aren't given the same freedom of expression in list-learning experiments.
-Evidence: presentation/recall correlation matrices during random free recall don't show the same sort of off-diagonal increase.}  Claim: mental time travel and quantum wave reinstatement look the same for random word list remembering, but very different for real-world remembering.
-  \item The specific ties between unfolding events forms a scaffolding for understanding how those events are related.
+Evidence: presentation/recall correlation matrices during random free recall don't show the same sort of off-diagonal increase.
+Claim: mental time travel and quantum wave reinstatement look the same for random word list remembering, but very different for real-world remembering.}
+  \item \textit{The specific ties between unfolding events forms a scaffolding for understanding how those events are related.
 This can be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces (although see Flatland Fallacy paper, with HyperTools paper as a potential counterpoint).
@@
 have similar shapes).
-Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).
+Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).}
     \item These ideas also have implications for other forms of learning (e.g.
@@
 
-\textbf{KEY QUESTIONS}:
-\begin{enumerate}
-  \item Is this more of a ``review'' paper, or should I highlight the new analyses more?
-  \item Who is this paper alienating and how can the narrative be reframed in a more inclusive way?
-  \item Which ideas are ``new'' vs. restatements of old ideas?
-  \item What are the key testable predictions to highlight, and how can they be communicated most clearly and effectively
-  \item What's an appropriate venue for this?
-\end{enumerate}
-
-\section*{How do our experiences and memories unfold over time?}
-Although real-world experiences can exhibit rich temporal structures, it can be illustrative to consider the transformation of how we experience much simpler and less naturalistic events (e.g.
-studying random word lists) versus how we later remember those experiences.
-Figure~\ref{fig:corrmats}A displays an average temporal correlation matrix describing the similarity structure of 1072 random word lists, each comprising 16 words~\citep{ZimaEtal18}.
-I applied a topic model (trained on a curated corpus of Wikipedia articles) to yield a 100-dimensional embedding of each presented word~\citep{BleiEtal03}.
-The left panel displays the average correlations between the topic vectors of each pair of presented words as a function of their presentation positions.
-As shown, nearly all of the correlation ``mass'' is along the diagonal, and the off-diagonal entries are very close to zero.
-Because the word orders were randomized, the temporal correlation matrix reflects that there are no consistent similarity patterns between any given pair of timepoints.
-The right panel of Figure~\ref{fig:corrmats}A displays the average temporal correlation matrices describing participants' \textit{recalls} of those same lists.
-Although the recall matrix exhibits the strongest correlations along its diagonal, some correlation mass bleeds into nearby recalls, reflecting (in part) participants' tendancies to recall semantically related words successively.\footnote{Note that participants recalled different numbers of words from each studied list; I used linear interpolation to resample each recall correlation matrix that preserved the size of the presentation correlation matrix.
-This process resulted in some additional blurring along the diagonal.
-In other words, the blurring along the diagonal reflects two phenomena: participants' tendancies to semanticly cluster their recalls, and an artifact of the temporal resampling.}  Differences in the temporal correlation structure between the original lists of words and the orders of participants' recalls illustrate one way that our memory systems can reorganize our experiences such that we recall them in a different way than we originally experienced them.
-
 \textbf{JRM NOTE: add section on the temporal structure of list-learning vs. naturalistic stimulus, and on the temporal structure of recalls.}
-
-\section*{The shapes of remembered experiences}
-In a recent study~\citep{HeusEtal18c} my lab examined functional neuroimaging and behavioral data collected by \cite{ChenEtal17} as participants viewed and then verbally recalled an episode of a television show.
-We used topic models~\citep{BleiEtal03} to characterize the moment-by-moment content of the episode and of participants' recalls.
-Each moment was quantified via a \textit{topic vector} that reflected the mix of topics (themes) the model identified in that moment.
-In this way, the sequence of topic vectors over the course of the episode traces out a \textit{topic trajectory} through a high-dimensional feature space.
-The sequence of topic vectors for a given participant's recalls traces out an analogous trajectory.
-This characterization provides a geometric framework for comparing the \textit{shapes} that reflect how naturalistic experiences unfold over time, to the shapes of how we remember those experiences later (Fig.~\ref{fig:trajectories}).
-One can then use the geometric transformations needed to map an experience's trajectory onto the trajectory of its later remembering to characterize the quality and content of memory.
-
-\textbf{JRM STOPPED HERE}
 

### 2019-01-10 boneyard.tex
working on section describing how ideas extend to other domains

@@
 
+\begin{itemize}
+\item Not about the number of details-- it's about capturing the overall ``shape'' of past experience
+\item Quantifying recall should be about the match between the shape of the original experience (i.e., the covariance structure describing how it unfolds over time), not about the number of details that are recovered
+\item When we retrieve memories, we cast ourselves along the full trajectory simultaneously (perhaps covering each component to a different degree), not to a single point (moment)
+\item Predictions:
+\begin{itemize}
+    \item What makes an effective summary?  One that efficiently captures the overall shape of the experience.
+    \item How do we communicate our experiences to others?  Describe the path through conceptual space.
+    \item How do we communicate our experiences to our ``future selves?''  Describe the path through conceptual space.
+    \item Each of these notions is fundamentally about the full trajectory, not about a single point.  Each moment derives meaning only in the context of the rest of experience.
+\end{itemize}
+\item What can list learning tell us?  We get fine control over  contextual representations at a very limited set of timescales.  We can study the detailed impacts of manipulations on memory.
+\item What can't list learning tell us?
+We can ask whether people precisely reproduce their prior experiences.
+But we can't know whether they recover the ``gist'' of those experiences, since random lists do not have meaningful gists.
+\item Other thoughts (from previous brainstorms):
+    \begin{itemize}
+    \item Reactivated neural representations that seem like noise should actually have predictive power for later memories if they get incorporated into the mental state at the time of remembering.
+    \item List learning experiments and other experiments with discrete trials are not well equipped to study time spreading because they effectively impose event boundaries around and between every stimulus presentation.
+Rather, we need continuous naturalistic recall experiments to more accurately reflect how our brains operate in the real world.
+    \item say something about slow drift vs. reinstatement (cite MannEtal11 and FolkEtal18)
+    \item something about schema learning (related to semantic memory?) and how memories are compressed (Chen et al 2017, Baldassano et al 2017, 2018) and transmitted to others
+    \end{itemize}
+  \end{itemize}
+
+  \section*{BRAINSTORMING OUTLINE OF THE REST OF THE PAPER}
+\begin{enumerate}
+  \item \textit{Show video correlation matrix + recall correlation matrix.
+Key idea: off-diagonal mass is introduced (primarily) by how people \textit{remember} the movie, not due to the contents of the movie itself (evidence: correlation matrices).
+Further, this is not simply a matter of imprecision in how people recall, and/or our approach's ability to characterize recalled events (evidence: reinstatement figure showing which scenes are recalled similarly).
+This sort of integration is how we ascribe meaning to events that unfold over disjointed extended timescales.}
+  \item \textit{In theory this could also happen during \textit{any} sort of experience.
+For example, when people study random word lists, they often re-order their recalls so that semantically related words are closer in the recall sequence than they were in the presented sequence.
+This can be captured to an extent by semantic clustering measures.
+However, it's difficult to examine this phenomenon in more depth because word lists necessarily lack the richness of more naturalistic stimuli.
+There are no ``deeper ties'' between study events (i.e., individual words), so the sorts of integration operations that allow us to make sense of real-world experiences aren't given the same freedom of expression in list-learning experiments.
+Evidence: presentation/recall correlation matrices during random free recall don't show the same sort of off-diagonal increase.
+Claim: mental time travel and quantum wave reinstatement look the same for random word list remembering, but very different for real-world remembering.}
+  \item \textit{The specific ties between unfolding events forms a scaffolding for understanding how those events are related.
+This can be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces (although see Flatland Fallacy paper, with HyperTools paper as a potential counterpoint).
+This reveals another key property of how we remember: although there are substantial individual differences in the specific words people use to describe a shared experience, and how much detail they get into, the overall ``shapes'' of those recalled experience trajectories are highly similar across people.
+This has potentially exciting implications for how we communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.
+It also has implications for how we learn by analogy: perhaps events that share schemas (cite Baldassano et al) unfold in similar ways (e.g.
+have similar shapes).
+Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).}
+    \item These ideas also have implications for other forms of learning (e.g.
+spatial learning) and neural decoding.
+When we attempt to characterize thoughts (e.g.
+training a classifier to distinguish between states), we tend to train decoders to estimate the ``one'' state that the brain is in.
+However, perhaps we should instead consider that our brains are contemplating, or holding in our thoughts, many possible states simultaneously.
+Therefore we should consider the full distribution of states (rather than only considering its maximum) as a primary metric of characterizing mental function.
+(Probably people have proposed this-- e.g.
+representing the posterior and using it to compute...need citations.)
+    \end{enumerate}
+
 % Next to expand on:
@@
 % - Option 2: get rid of the list learning stuff and transition directly to the "shape" of how experiences unfold.
+
+    \textbf{JRM NOTE: add section on the temporal structure of list-learning vs. naturalistic stimulus, and on the temporal structure of recalls.}
+
+    Our brains exhibit autocorrelated activity.
+In the extreme this is trivially true; our brains are biophysical systems that cannot change infinitely quickly, so neural activity patterns are necessarily similar (to some extent) from one moment to the next.
+However, autocorrelated activity patterns \textit{at the time scales we typically examine in laboratory memory studies} are far stronger and persist for far longer than is strictly necessary from a biophysical standpoint.
+For example, transient changes in neuronal membrane potentials during action potentials can last just a few milliseconds~\citep{HodgHuxl52} and even local field potentials that reflect the synchronized activities of tens of thousands of neurons can exhibit millisecond timescale changes~\citep{Frie05, Buzs06}.
+Nevertheless, our brain activity pattern can remain reliably autocorrelated at timescales of seconds or longer-- and these long timescale autocorrelations predict how our ongoing experiences are remembered later~\citep[e.g.][]{MannEtal07, MannEtal11, HowaEtal12, FolkEtal18}.
 

### 2019-01-10 main.tex
working on section describing how ideas extend to other domains

@@
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.8\textwidth]{figs/trajectory}
+\caption{\textbf{The shape of an unfolding experience.}  \textbf{A.
+The trajectory through topic space taken by a television episode.} The topic trajectory taken by a 50 minute television episode, projected onto two dimensions.
+Each dot indicates an event identified using a Hidden Markov Model; the dot colors denote the order of the events (early events are in red; later events are in blue), and the connecting lines indicate the transitions between successive events.
+\textbf{B.
+The trajectory through topic space taken by participants' recalls of the television episode.} The average two-dimensional trajectory captured by 17 participants' recalls of the episode, with the same format and coloring as the trajectory in Panel A.
+The arrows reflect the directions of the average transitions taken by participants whose trajectories crossed that part of topic space; blue denotes reliable agreement across participants.
+Figure adapted from \cite{HeusEtal18c}.}
+\label{fig:trajectories}
+\end{figure}
+
 The specific ties between unfolding events forms a scaffolding for understanding how those events are related.
@@
 
-\textbf{JRM STOPPED HERE}
-
-\section*{BRAINSTORMING OUTLINE OF THE REST OF THE PAPER}
-\begin{enumerate}
-  \item \textit{Show video correlation matrix + recall correlation matrix.
-Key idea: off-diagonal mass is introduced (primarily) by how people \textit{remember} the movie, not due to the contents of the movie itself (evidence: correlation matrices).
-Further, this is not simply a matter of imprecision in how people recall, and/or our approach's ability to characterize recalled events (evidence: reinstatement figure showing which scenes are recalled similarly).
-This sort of integration is how we ascribe meaning to events that unfold over disjointed extended timescales.}
-  \item \textit{In theory this could also happen during \textit{any} sort of experience.
-For example, when people study random word lists, they often re-order their recalls so that semantically related words are closer in the recall sequence than they were in the presented sequence.
-This can be captured to an extent by semantic clustering measures.
-However, it's difficult to examine this phenomenon in more depth because word lists necessarily lack the richness of more naturalistic stimuli.
-There are no ``deeper ties'' between study events (i.e., individual words), so the sorts of integration operations that allow us to make sense of real-world experiences aren't given the same freedom of expression in list-learning experiments.
-Evidence: presentation/recall correlation matrices during random free recall don't show the same sort of off-diagonal increase.
-Claim: mental time travel and quantum wave reinstatement look the same for random word list remembering, but very different for real-world remembering.}
-  \item \textit{The specific ties between unfolding events forms a scaffolding for understanding how those events are related.
-This can be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces (although see Flatland Fallacy paper, with HyperTools paper as a potential counterpoint).
-This reveals another key property of how we remember: although there are substantial individual differences in the specific words people use to describe a shared experience, and how much detail they get into, the overall ``shapes'' of those recalled experience trajectories are highly similar across people.
-This has potentially exciting implications for how we communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.
-It also has implications for how we learn by analogy: perhaps events that share schemas (cite Baldassano et al) unfold in similar ways (e.g.
-have similar shapes).
-Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).}
-    \item These ideas also have implications for other forms of learning (e.g.
-spatial learning) and neural decoding.
-When we attempt to characterize thoughts (e.g.
-training a classifier to distinguish between states), we tend to train decoders to estimate the ``one'' state that the brain is in.
-However, perhaps we should instead consider that our brains are contemplating, or holding in our thoughts, many possible states simultaneously.
-Therefore we should consider the full distribution of states (rather than only considering its maximum) as a primary metric of characterizing mental function.
-(Probably people have proposed this-- e.g.
-representing the posterior and using it to compute...need citations.)
-    \end{enumerate}
-
-\textbf{JRM NOTE: add section on the temporal structure of list-learning vs. naturalistic stimulus, and on the temporal structure of recalls.}
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.8\textwidth]{figs/trajectory}
-\caption{\textbf{Topic trajectories of a video and its memory.}  \textbf{A.} The topic trajectory taken by a 50 minute television episode, projected onto two dimensions.
-Each dot indicates an event identified using a Hidden Markov Model; the dot colors denote the order of the events (early events are in red; later events are in blue), and the connecting lines indicate the transitions between successive events.
-\textbf{B.} The average two-dimensional trajectory captured by 17 participants' recalls of the episode, with the same format and coloring as the trajectory in Panel A.
-The arrows reflect the directions of the average transitions taken by participants whose trajectories crossed that part of topic space; blue denotes reliable agreement across participants.
-Figure adapted from \cite{HeusEtal18c}.}
-\label{fig:trajectories}
-\end{figure}
-
-defines a distinction between the gist
-
-\begin{itemize}
-\item Not about the number of details-- it's about capturing the overall ``shape'' of past experience
-\item Quantifying recall should be about the match between the shape of the original experience (i.e., the covariance structure describing how it unfolds over time), not about the number of details that are recovered
-\item When we retrieve memories, we cast ourselves along the full trajectory simultaneously (perhaps covering each component to a different degree), not to a single point (moment)
-\item Predictions:
-\begin{itemize}
-    \item What makes an effective summary?  One that efficiently captures the overall shape of the experience.
-    \item How do we communicate our experiences to others?  Describe the path through conceptual space.
-    \item How do we communicate our experiences to our ``future selves?''  Describe the path through conceptual space.
-    \item Each of these notions is fundamentally about the full trajectory, not about a single point.  Each moment derives meaning only in the context of the rest of experience.
-\end{itemize}
-\item What can list learning tell us?  We get fine control over  contextual representations at a very limited set of timescales.  We can study the detailed impacts of manipulations on memory.
-\item What can't list learning tell us?
-We can ask whether people precisely reproduce their prior experiences.
-But we can't know whether they recover the ``gist'' of those experiences, since random lists do not have meaningful gists.
-\item Other thoughts (from previous brainstorms):
-    \begin{itemize}
-    \item Reactivated neural representations that seem like noise should actually have predictive power for later memories if they get incorporated into the mental state at the time of remembering.
-    \item List learning experiments and other experiments with discrete trials are not well equipped to study time spreading because they effectively impose event boundaries around and between every stimulus presentation.
-Rather, we need continuous naturalistic recall experiments to more accurately reflect how our brains operate in the real world.
-    \item say something about slow drift vs. reinstatement (cite MannEtal11 and FolkEtal18)
-    \item something about schema learning (related to semantic memory?) and how memories are compressed (Chen et al 2017, Baldassano et al 2017, 2018) and transmitted to others
-    \end{itemize}
-\end{itemize}
-
-Our brains exhibit autocorrelated activity.
-In the extreme this is trivially true; our brains are biophysical systems that cannot change infinitely quickly, so neural activity patterns are necessarily similar (to some extent) from one moment to the next.
-However, autocorrelated activity patterns \textit{at the time scales we typically examine in laboratory memory studies} are far stronger and persist for far longer than is strictly necessary from a biophysical standpoint.
-For example, transient changes in neuronal membrane potentials during action potentials can last just a few milliseconds~\citep{HodgHuxl52} and even local field potentials that reflect the synchronized activities of tens of thousands of neurons can exhibit millisecond timescale changes~\citep{Frie05, Buzs06}.
-Nevertheless, our brain activity pattern can remain reliably autocorrelated at timescales of seconds or longer-- and these long timescale autocorrelations predict how our ongoing experiences are remembered later~\citep[e.g.][]{MannEtal07, MannEtal11, HowaEtal12, FolkEtal18}.
-
-\section*{Implications}
 \subsection*{Extensions to spatial and semantic memory}
@@
 However, following the ``quantum wave function'' logic, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.
+What are the implications of this view?
 

### 2019-01-11 main.tex
fleshing out a bunch of stuff in the paper; one more section to go, on extensions to non-temporal domains

@@
 Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.
-This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11}.
-Recent work also suggests that the lateral entorhinal cortex plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}.
-These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14}.
+This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
+Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}.
+These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
@@
 Although reactivating context is often \textit{described} as a sort of mental time travel back to the one specific moment being remembered, there is some apparent inconsistency between this description and how this process is actually characterized mathematically or measured neurally.
-In particular, both the theory and neural measures described in the preceding paragraphs characterize episodic recall mathematically as reflecting the reactivation of a weighted blend of thoughts from a \textit{range} of previously experienced times-- not to any one \textit{specific} time.
+In particular, both the theory and neural measures described in the preceding paragraphs characterize episodic recall (mathematically) as reflecting the reactivation of a weighted blend of thoughts from a \textit{range} of previously experienced times-- not to any one \textit{specific} time.
 This is not simply a matter of imprecision in mental time travel (i.e., a reflection of the low effective resolution with which we can revisit specific moments from our past).
@@
 For example, understanding how a new experience fits in with our broader understanding might require integrating information gleaned from many experiences separated in time, each with their own contextual attributes.
-That we can integrate information across experiences suggests that the integration of incoming information at different rates need not be a purely mechanical process, but rather may reflect a deeper functional role.
+That we can integrate information across experiences suggests that the integration of incoming information at different rates need not be a purely mechanical process, but rather may reflect a deeper functional role~\citep[also see][]{BrigEtal18}.
 When we simultaneously reactivate thoughts related to separated moments from our past, this provides a means of associating and relating those temporally distinct experiences.
@@
 Event boundaries are one example of the broader temporal covariance structure that is characteristic of ongoing experience.
-Whereas classic approaches to studying memory (e.g., list-learning paradigms and other trial-based experiments) encourage researchers to treat each moment, trial, or stimulus as separable from the rest of experience, context-based theories of memory (including situation models and event-based models) posit that each moment derives meaning through its deep ties to other experiences.
-This rich tapestry of interacting moments and experiences forms the scaffolding for interpreting our new experiences and for retrieving information concerning our past.
+Whereas classic approaches to studying memory (e.g., list-learning paradigms and other trial-based experiments) encourage researchers to treat each moment, trial, or stimulus as separable from the rest of experience, context-based theories of memory (including situation models and event-based models) posit that each moment derives meaning through its deep ties to other related (but potentially temporally separated) experiences.
+This rich tapestry of interacting moments and experiences forms a scaffolding for interpreting our new experiences and for retrieving information concerning our past.
 According to this view, the reactivation of this rich tapestry of contextual details enfolding the remembered event (i.e., how the event relates to the rest of ones experiences across timescales) is even more central to our subjective experience than the specific details of what occurred during that event!
@@
 Our ongoing experiences can cue us to dredge up information about our past.
-This can happen explicitly (e.g., during a conversation about a prior event) or implicitly (e.g., when something happening now reminds you of what happened before).
+This can happen explicitly (e.g., during a conversation about a specific prior event) or implicitly (e.g., when something happening now reminds you of what happened before).
 Context-based theories of memory posit that the particular memories that our ongoing experiences cue will be related~\citep[contextually, semantically, functionally, etc.; e.g., ][]{PolyEtal09}.
@@
 
-Once insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \citep{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
+Once insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \cite{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
 The authors used topic models~\citep{BleiEtal03} applied to human annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.
+These \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode.
 When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change.
-This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \citep{HeusEtal18c} perform detailed analyses of these putative events and how they are remembered.
-Notably, the off-diagonal entries of the temporal correlation matrix (which reflects content similarity between temporally distant events in the episode) are nearly all close to 0.
+This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \cite{HeusEtal18c} performed detailed analyses of these putative events and how they are remembered.
+Notably, the off-diagonal entries of the temporal correlation matrix (which reflect content similarity between temporally distant events in the episode) are nearly all close to 0.
 This is surprising, because the episode itself comprises a murder mystery with recurring motifes that must be associated in the viewer's mind in order to make sense of the episode.
@@
 If we examine the temporal autocorrelation matrices of participants' recalls of the television episode (characterized using the same topic model used to capture the content of the episode; Figs.~\ref{fig:corrmats}B, C), we see a different structure than that of the episode itself.
-Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off diagonal correlations between temporally distant events are substantially stronger in participants' recalls than in the original episode.
+Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off-diagonal correlations between the thematic content of temporally distant events are substantially stronger in participants' recalls than in the original episode.
 Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode and participants' recalls (Fig.~\ref{fig:reinstatement}A).
-Whereas the autocorrelations for the original episode rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around 0.2.
+Whereas the autocorrelations in the original episode's content rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around 0.2.
 These analyses indicate that the way participants \textit{recount} temporally separated events imposes additional similarity on those events, beyond what was present in the original events themselves.
 One can then ask: is this simply a matter of imprecision in recall, either with respect to participants' recalls themselves, or with respect to the quality with which they are characterized via semantic feature vectors?
-Or might there be a functional role to that additional imposed similarity structure?
+Or might there be a functional role of that additional imposed similarity structure?
 
-To examine which specific events are described (during recall) as more similar than what their original experiences (as characterized by semantic feature vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.
-Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were described in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
+To examine which specific events are described (during recall) as more similar than what their original experiences (as characterized by their topic vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.
+Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were characterized in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
 The example event explored in the figure comprises a scene where Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another victim has been murdered.
@@
 A rich recall density function analogous to the one displayed in Figure~\ref{fig:reinstatement}B could, in principle, be characteristic of memory for \textit{any} past event or experience.
-For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline, were functionally associated in memory.
-Nonetheless, a potential challenge to testing this hypothesis in the laboratory is that in most memory studies there is no particular reason for participants to functionally associate the stimuli-- there is no ``deeper narrative'' to make sense of.
+For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline, are functionally associated in memory.
+Nonetheless, a potential challenge to testing this hypothesis in the laboratory is that in many memory studies there is no particular reason for participants to functionally associate the stimuli.
+For example, there is often no ``deeper narrative'' to make sense of in list-learning studies or in most standard trial-based memory experiments.
 However, random word list learning experiments might happen to contain (for a given list) semantically related words.
@@
 \section*{The shapes of remembered experiences}
-In a recent study~\citep{HeusEtal18c} my lab examined functional neuroimaging and behavioral data collected by \cite{ChenEtal17} as participants viewed and then verbally recalled an episode of a television show.
-We used topic models~\citep{BleiEtal03} to characterize the moment-by-moment content of the episode and of participants' recalls.
-Each moment was quantified via a \textit{topic vector} that reflected the mix of topics (themes) the model identified in that moment.
-In this way, the sequence of topic vectors over the course of the episode traces out a \textit{topic trajectory} through a high-dimensional feature space.
+The study by \cite{HeusEtal18c} described above suggests one way of characterizing the dynamic content of complex naturalistic stimuli like television episodes.
+The sequence of topic vectors over the course of the episode traces out a \textit{topic trajectory} through a high-dimensional feature space.
 The sequence of topic vectors for a given participant's recalls traces out an analogous trajectory.
@@
 
-The specific ties between unfolding events forms a scaffolding for understanding how those events are related.
-In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\citep{JollChan18}.
-\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats}, \ref{fig:reinstatement}, and~\ref{fig:trajectories}, reveals another key property of how we remember.
+In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
+\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats}, \ref{fig:reinstatement}, and~\ref{fig:trajectories}, revealing another key property of how we remember.
 Although there are substantial individual differences in the specific words people use to describe a shared experience, and how much detail they get into, the overall ``shapes'' of those recalled experience trajectories are highly similar across people.
@@
 
-\subsection*{Extensions to spatial and semantic memory}
+\subsection*{Extensions to spatial, semantic, and social memory}
 Just as we mentally spread ourselves across a distribution of times, we can perform similar mental operations when we conceptualize and retrieve other types of information.
-For example, consider how we use pattern classifiers to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12} or spatial~\citep{MillEtal13} content being mentally manipulated.
+For example, consider how we use pattern classifiers to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, and social~\citep{MeyeEtal18} content being mentally manipulated.
 The prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.
@@
 Traditional approaches then ``clean up'' this distribution over possible mental states by taking its expected or maximum value.
-However, following the ``quantum wave function'' logic, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.
+However, following the ``quantum wave function'' logic described in this manuscript, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.
 What are the implications of this view?
 
+\begin{itemize}
+\item event boundaries and goal boundaries-- environment barriers, non-sensical concepts \item not every nearby stimulus is equally likely or equally reachable.
+Stimuli live on a manifold; could use the shape to understand the geometry of thought, nature of problem space, improve decoding, etc.
+  \item links to GANs (pix2pix, etc.)?
+\end{itemize}
+
 \section*{Concluding remarks}
+When we revisit an experience from our past, we bring back thoughts related to a range of other semantically and functionally related experiences.
+This can serve an important functional role: it allows us to integrate information across temporally separated experiences.
+Although this view is captured mathematically by context-based theories of episodic memory, it is \textit{not} specifically captured by the conceptual notion of mental time travel that is often used to describe episodic recall.
+The alternative conceptualization presented here, that when we remember our past we distribute our thoughts over those associated with a weighted blend of our prior experiences, also has links with non-temporal or episodic domains such as semantic, spatial, and social representations.
+Characterizing how our thoughts reflect \textit{distributions} over representational spaces is central to understanding how we think about ourselves, our experiences, other people, and the world around us.
 

### 2019-01-11 main.tex
finished complete rough draft

@@
 Notably, the off-diagonal entries of the temporal correlation matrix (which reflect content similarity between temporally distant events in the episode) are nearly all close to 0.
-This is surprising, because the episode itself comprises a murder mystery with recurring motifes that must be associated in the viewer's mind in order to make sense of the episode.
+This is surprising, because the episode itself comprises a murder mystery with recurring motifs that must be associated in the viewer's mind in order to make sense of the episode.
 Nonetheless, when considering solely the semantic content of each moment of the episode, those long timescale associations are nearly entirely absent.
@@
 \textbf{C.
-Temporal correlations in the content of individual participants' recalls of the episode.} Each sub-panel displays the temporal correlation matrix for one experimental particpant.}
+Temporal correlations in the content of individual participants' recalls of the episode.} Each sub-panel displays the temporal correlation matrix for one experimental participant.
+Figure adapted from \cite{HeusEtal18c}.}
 \label{fig:corrmats}
@@
 Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states.
-Traditional approaches then ``clean up'' this distribution over possible mental states by taking its expected or maximum value.
+Traditional approaches then ``clean up'' this distribution over possible mental states by taking its average (expected), maximum, or modal value.
 However, following the ``quantum wave function'' logic described in this manuscript, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.
@@
 
-\begin{itemize}
-\item event boundaries and goal boundaries-- environment barriers, non-sensical concepts \item not every nearby stimulus is equally likely or equally reachable.
-Stimuli live on a manifold; could use the shape to understand the geometry of thought, nature of problem space, improve decoding, etc.
-  \item links to GANs (pix2pix, etc.)?
-\end{itemize}
+One example of how considering the full distribution over possible mental states can provide insights into cognition may be seen in how the brain encodes spatial locations.
+Theoretical~\citep[e.g., ][]{GersAbbo97} and empirical~\citep[e.g., ][]{PfeiFost13} work has demonstrated how place cell receptive fields may be modulated according to environmental features such as obstacles and the locations of rewarded goals.
+In other words, although place cells encode an animal's location within an environment, there is some variability with respect to where precisely within the environment the cell modulates its firing rate.
+These studies suggest that variability in the set of locations to which a place cell responds is not merely a reflection of noise or imprecision.
+Rather, this variability might have some functional relevance in terms of helping to learn paths around obstacles and towards goals.
+
+The notion that variability can be informative also has implications for how our brains might efficiently represent information.
+Recent work suggests that adversarial networks trained to learn a generative model of natural images~\citep[e.g., ][]{GatyEtal16, IsolEtal17} may also be used to refine estimates of decoded visual stimuli~\citep{StYvNase18}.
+Related manifold learning approaches that ``snap'' decoded estimates to the nearest surface location on a learned manifold may be used in a similar way to decode sequences of movie frames~\citep{HeusEtal18a}.
+Taken together, these findings suggest that not all \textit{possible} images (and by extension, all possible stimuli) are equally likely.
+Rather, the set of stimuli likely to be encountered live within subspaces or on manifolds in the relevant representational space.
+The geometry of these manifolds provides insights into the statistical structure of the stimuli.
+It also provides potential insights into how the brain represents those stimuli, for example to the extent that the geometries of these manifolds may be leveraged to improve brain decoding.
 

### 2019-01-12 main.tex
polishing text

@@
 
-Once insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \cite{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
-The authors used topic models~\citep{BleiEtal03} applied to human annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.
+One insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \cite{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
+The authors used topic models~\citep{BleiEtal03} applied to human-generated annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.
 These \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode.
 When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change.
-This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \cite{HeusEtal18c} performed detailed analyses of these putative events and how they are remembered.
+This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \cite{HeusEtal18c} performed detailed analyses of these putative \textit{events} and how they are remembered.
 Notably, the off-diagonal entries of the temporal correlation matrix (which reflect content similarity between temporally distant events in the episode) are nearly all close to 0.
@@
 
-\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/pres_rec_corrmats_sherlock} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal18c}, where full details may be found.
+\begin{figure}[tp] \centering \includegraphics[width=0.8\textwidth]{figs/pres_rec_corrmats_sherlock} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal18c}, where full details may be found.
 \textbf{A.
-Temporal correlations exhibited by the content of the television episode.}  Each moment of an episode of the BBC television show \textit{Sherlock} has been characterized using a topic model~\citep{BleiEtal03} fit to detailed human-generated annotations of each scene cut.
-The correlations between the topic vectors of a given pair of timepoints reflects the similarity in content between those timepoints.
+Temporal correlations exhibited by the content of the television episode.}  Each moment of an episode of the BBC television show \textit{Sherlock} has been characterized using a topic model~\citep{BleiEtal03} fit to detailed human-generated annotations between each scene cut.
+The correlations between the topic vectors of a given pair of timepoints reflects the similarity in thematic content at those timepoints.
 \textbf{B.
-Temporal correlations exhibited by the (average) content of participants' recalls of the episode.} The topic model used to characterize the episode in Panel A has been applied to each sentence, from each participants' recall transcript.
+Temporal correlations exhibited by the (average) content of participants' recalls of the episode.} The topic model used to characterize the episode in Panel A has been applied to each sentence from each participants' recall transcript.
 The temporal correlation matrices describing each participants' recalls were re-sampled using cubic interpolation to have 100 rows and columns prior to averaging the resampled matrices across participants.
 \textbf{C.
-Temporal correlations in the content of individual participants' recalls of the episode.} Each sub-panel displays the temporal correlation matrix for one experimental participant.
-Figure adapted from \cite{HeusEtal18c}.}
+Temporal correlations in the content of individual participants' recalls of the episode.} Each sub-panel displays the temporal correlation matrix for one experimental participant.}
 \label{fig:corrmats}
@@
 
-Although the specific content of each moment of the episode is largely unique, participants' subjective experiences of the episode entail weaving together those temporally separated moments into a cohesive narrative.
+Although the specific mix of contents in each moment of the episode is nearly unique, participants' subjective experiences of the episode entail weaving together those temporally separated moments into a cohesive narrative.
 This is reflected in how people verbally recount the episode later.
@@
 Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode and participants' recalls (Fig.~\ref{fig:reinstatement}A).
-Whereas the autocorrelations in the original episode's content rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around 0.2.
+Whereas the autocorrelations in the original episode's content rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around an average of 0.2.
 These analyses indicate that the way participants \textit{recount} temporally separated events imposes additional similarity on those events, beyond what was present in the original events themselves.
@@
 
-To examine which specific events are described (during recall) as more similar than what their original experiences (as characterized by their topic vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.
+To understand which specific events are described (during recall) as more similar than what their original experiences (as characterized by their topic vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.
 Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were characterized in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
@@
 The language used by human annotators describing the moment-by-moment contents of the scene used a similar mix of themes only for moments in the episode that were temporally proximal to the event in question (gray line).
-However, the way participants \textit{described} what happened when they recounted the episode later revealed a much different pattern.
-Specifically, their language mirrored that used when they were recalling other murders, and other moments when Sherlock and John were bantering (Fig.~\ref{fig:reinstatement}C).
-Further, the way they described this example event did \textit{not} mirror the way they described other semantically (but not conceptually) related events, such as events involving unrelated violence, or Sherlock and John interacting with other characters.
+However, the way participants described what happened when they \textit{recounted} the episode later revealed a much different pattern.
+Specifically, their language mirrored that used when they were recalling other murders, and other moments when Sherlock and John were bantering with or about each other (Fig.~\ref{fig:reinstatement}C).
+The way they described the example event did \textit{not} mirror the way they described other semantically (but not conceptually) related events, such as events involving unrelated violence, or Sherlock and John interacting with other characters about other subjects.
 Taken together, this suggests that these specific events were associated in participants' memories, perhaps contributing to (or reflecting) their understanding that those events were linked in the episode's narrative.
@@
 Local maxima in the average recall density function.}  Representative screen captures from the scenes at each local maxima in Panel B are displayed (dot colors match those in Panel B).
-The text in each screen capture displays a brief summary of what happened in each scene.}
+The text in each screen capture displays a brief summary of what happened in each scene.
+The colored outline denotes the reference scene being recalled.}
   % colors: individual participants
@@
 However, random word list learning experiments might happen to contain (for a given list) semantically related words.
-And indeed, a number of studies indicate that participants spontaneously form associations between semantically related words on random word lists, even when those related words are separated in time~\citep[e.g.,][]{WixtRohr94, MannKaha12, MannEtal12}.
+Indeed, a number of studies indicate that participants spontaneously form associations between semantically related words on random word lists, even when those related words are separated in time~\citep[e.g.,][]{WixtRohr94, MannKaha12, MannEtal12}.
 This suggests that our memory systems pick up on statistical regularities, even in ostensibly ``random'' stimulus sequences, and that we leverage those regularities when we recall our past~\citep{PolyEtal09}.
@@
 The arrows reflect the directions of the average transitions taken by participants whose trajectories crossed that part of topic space; blue denotes reliable agreement across participants.
-Figure adapted from \cite{HeusEtal18c}.}
+Figure adapted from \cite{HeusEtal18c}, where additional details may be found.}
 \label{fig:trajectories}
@@
 \cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats}, \ref{fig:reinstatement}, and~\ref{fig:trajectories}, revealing another key property of how we remember.
-Although there are substantial individual differences in the specific words people use to describe a shared experience, and how much detail they get into, the overall ``shapes'' of those recalled experience trajectories are highly similar across people.
-This has potentially exciting implications for how we communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.
+Although there are substantial individual differences in the specific words people use to describe their shared experience of watching the same television episode, and how much detail they recollected, the overall shapes of those recalled experience trajectories are highly similar across people.
+This has potentially exciting implications for what it means to communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.
 It also has implications for how we learn by analogy: perhaps events that share schemas~\citep[e.g., ][]{BaldEtal18}  unfold in similar ways (e.g., have similar shapes).
@@
 
-% analysis idea: matrix of reinstatement functions for each timepoint-- e.g.
-considering each timepoint in turn (rows), how much are participants (on average) using language that mirrors how they talk about each other timepoint (columns).
-there should be a strong diagonal, with some blockiness, and possibly some disjoint groups showing how events cluster.
-
 \subsection*{Extensions to spatial, semantic, and social memory}
-Just as we mentally spread ourselves across a distribution of times, we can perform similar mental operations when we conceptualize and retrieve other types of information.
-For example, consider how we use pattern classifiers to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, and social~\citep{MeyeEtal18} content being mentally manipulated.
-The prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.
+Just as we mentally spread ourselves across a distribution of times when we remember, we can perform similar mental operations when we conceptualize and retrieve other types of information.
+Further these distributions may be estimated using existing methods.
+For example, consider how we use pattern classifiers~\citep{NormEtal06b} to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, and social~\citep{MeyeEtal18} content being mentally manipulated.
+A prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.
 Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states.
@@
 However, following the ``quantum wave function'' logic described in this manuscript, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.
-What are the implications of this view?
+What are some implications of this view for non-memory domains?
 
@@
 Theoretical~\citep[e.g., ][]{GersAbbo97} and empirical~\citep[e.g., ][]{PfeiFost13} work has demonstrated how place cell receptive fields may be modulated according to environmental features such as obstacles and the locations of rewarded goals.
-In other words, although place cells encode an animal's location within an environment, there is some variability with respect to where precisely within the environment the cell modulates its firing rate.
+In other words, although place cells encode an animal's location within an environment, even for a fixed cell and environment there is some variability with respect to where precisely within that environment the cell modulates its firing rate.
 These studies suggest that variability in the set of locations to which a place cell responds is not merely a reflection of noise or imprecision.
-Rather, this variability might have some functional relevance in terms of helping to learn paths around obstacles and towards goals.
+Rather, this variability might have some functional relevance in terms of facilitating the learning of paths around obstacles and towards goals.
 
-The notion that variability can be informative also has implications for how our brains might efficiently represent information.
-Recent work suggests that adversarial networks trained to learn a generative model of natural images~\citep[e.g., ][]{GatyEtal16, IsolEtal17} may also be used to refine estimates of decoded visual stimuli~\citep{StYvNase18}.
+The notion that variability and ``uncertainty'' can be informative also has implications for how our brains might efficiently represent information.
+Recent work suggests that adversarial networks trained to learn a generative model of natural images~\citep[e.g., ][]{GatyEtal16, IsolEtal17} may be used to refine estimates of decoded visual stimuli~\citep{StYvNase18}.
 Related manifold learning approaches that ``snap'' decoded estimates to the nearest surface location on a learned manifold may be used in a similar way to decode sequences of movie frames~\citep{HeusEtal18a}.
@@
 It also provides potential insights into how the brain represents those stimuli, for example to the extent that the geometries of these manifolds may be leveraged to improve brain decoding.
+The shapes fo these manifolds can potentially provide insights into how our thoughts are distributed over representational spaces.
 
 \section*{Concluding remarks}
-When we revisit an experience from our past, we bring back thoughts related to a range of other semantically and functionally related experiences.
+When we revisit an experience from our past, we bring back thoughts related to a range of other semantically, conceptually, and functionally related experiences.
 This can serve an important functional role: it allows us to integrate information across temporally separated experiences.
 Although this view is captured mathematically by context-based theories of episodic memory, it is \textit{not} specifically captured by the conceptual notion of mental time travel that is often used to describe episodic recall.
-The alternative conceptualization presented here, that when we remember our past we distribute our thoughts over those associated with a weighted blend of our prior experiences, also has links with non-temporal or episodic domains such as semantic, spatial, and social representations.
+The alternative conceptualization presented here, that when we remember our past we reinstate thoughts associated with a weighted blend of our prior experiences, also has links with non-temporal or episodic domains such as semantic, spatial, and social representations.
 Characterizing how our thoughts reflect \textit{distributions} over representational spaces is central to understanding how we think about ourselves, our experiences, other people, and the world around us.

### 2019-01-14 main.tex
incorporated TSM feedback, updated figs to PNGs to reduce file size, added presubmission inquiry letter

@@
 How do our brains organize, retrieve, and act upon the ceaseless flow of incoming information and experiences?
-\cite{Tulv83} contends that episodic memory is like a sort of ``mental time travel,'' whereby we send some part of our mental state back along our autobiographical time line to a specific moment in our past.
+\cite{Tulv83} contends that episodic memory is like a sort of ``mental time travel,'' whereby we send some part of our mental state back along our autobiographical timeline to a specific moment in our past.
 This view has fundamental implications for how we study and build theories about memory: it suggests that we can match up specific moments of recollection with specific moments from the past.
 These implications extend to how we design episodic memory experiments and how we analyze the behavioral and neural data from those experiments in order to understand the neural and/or cognitive mechanisms that underly memory retrieval.
-At its core, characterizing episodic memory as mental time travel makes studying and understanding memory about matching up the thoughts (and/or brain activity patterns) participants are experiencing at one moment with the thoughts (and/or brain activity patterns) participants experienced at another previous moment.
+At its core, characterizing episodic memory as mental time travel makes studying and understanding memory about matching.
+Specifically, the primary implied questions are about how and when we match up the thoughts (and/or brain activity patterns) we are experiencing at one moment with the thoughts (and/or brain activity patterns) we experienced at another previous moment.
 The challenge, then, is to understand the neural (structural) and cognitive (functional) mechanisms that characterize when, how, and why our memory systems allow us to relive those prior experiences.
@@
 Where the mental time travel framing breaks down most notably is when we consider scenarios in which the particular blend of timepoints reflected in one's current thoughts do not come from temporally contiguous events.
-For example, understanding how a new experience fits in with our broader understanding might require integrating information gleaned from many experiences separated in time, each with their own contextual attributes.
+Understanding how a new experience fits in with our broader understanding might require integrating information gleaned from many experiences separated in time, each with their own contextual attributes.
+For example, having an important conversation with a friend might require bringing in information acquired during other interactions with that friend, other conversations about related or similarly important topics, general knowledge about communication styles, etc.
+\citep[This notion of bringing a range of prior experiences to bear on guiding our ongoing behaviors is also related to a literature on \textit{situtation models}; for review see][]{RangRitc12}.
 That we can integrate information across experiences suggests that the integration of incoming information at different rates need not be a purely mechanical process, but rather may reflect a deeper functional role~\citep[also see][]{BrigEtal18}.
@@
 
-Unlike random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g., shifting from the quiet peace of working on a manuscript in the early morning to the different peace of helping a newly awoken child get ready for school).
+Unlike random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g., shifting from the quiet peace of working on a manuscript in the early morning to the frenzied chaos of helping a newly awoken child get ready for school).
 A number of studies (using non-random lists and more naturalistic stimuli) have found that the way we remember past experiences can be shaped by these event boundaries.
@@
 This rich tapestry of interacting moments and experiences forms a scaffolding for interpreting our new experiences and for retrieving information concerning our past.
-According to this view, the reactivation of this rich tapestry of contextual details enfolding the remembered event (i.e., how the event relates to the rest of ones experiences across timescales) is even more central to our subjective experience than the specific details of what occurred during that event!
+According to this view, the reactivation of this rich tapestry of contextual details enfolding the remembered event (i.e., how the event relates to the rest of one's experiences across timescales) is even more central to our subjective experience than the specific details of what occurred during that event!
 This begs the question: which thoughts, from which times, do we reactivate when we remember our past experiences?
@@
 
-\begin{figure}[tp] \centering \includegraphics[width=0.8\textwidth]{figs/pres_rec_corrmats_sherlock} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal18c}, where full details may be found.
+\begin{figure}[tp] \centering \includegraphics[width=0.8\textwidth]{figs/pres_rec_corrmats_sherlock.png} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal18c}, where full details may be found.
 \textbf{A.
@@
 
-\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Content drift in a television episode and participants' recalls of that episode.
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement.png} \caption{\textbf{Content drift in a television episode and participants' recalls of that episode.
 A.
@@
 Local maxima are marked with colored dots.
-Local maxima closer than 2.5\% of the total recall duration to a higher local maximum were excluded, as were maxima with peaks with correlation coefficients lower than 0.1 (arbitrarily chosen; these ).
+Local maxima closer than 2.5\% of the total recall duration to a higher local maximum were excluded (this constraint served to eliminate very nearby peaks that corresponded to duplicate events), as were maxima with peaks with correlation coefficients lower than 0.1 (this threshold was chosen arbitrarily; this constraint served to eliminate events that were only very weakly associated with the example event).
 The colored lines denote the correlations for each individual participant's recalls.
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figs/trajectory}
+\includegraphics[width=0.8\textwidth]{figs/trajectory.png}
 \caption{\textbf{The shape of an unfolding experience.}  \textbf{A.
@@
 In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
-\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats}, \ref{fig:reinstatement}, and~\ref{fig:trajectories}, revealing another key property of how we remember.
+\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figure~\ref{fig:trajectories}, revealing another key property of how we remember.
 Although there are substantial individual differences in the specific words people use to describe their shared experience of watching the same television episode, and how much detail they recollected, the overall shapes of those recalled experience trajectories are highly similar across people.
@@
 
-\subsection*{Extensions to spatial, semantic, and social memory}
+\subsection*{Extensions to spatial, semantic, affective, and social memory}
 Just as we mentally spread ourselves across a distribution of times when we remember, we can perform similar mental operations when we conceptualize and retrieve other types of information.
-Further these distributions may be estimated using existing methods.
-For example, consider how we use pattern classifiers~\citep{NormEtal06b} to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, and social~\citep{MeyeEtal18} content being mentally manipulated.
+Further, these distributions may be estimated using existing methods.
+For example, consider how we use pattern classifiers~\citep{NormEtal06b} to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, affective~\citep{ChanEtal18}, and social~\citep{MeyeEtal18} content being mentally manipulated.
 A prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.
@@
 
-One example of how considering the full distribution over possible mental states can provide insights into cognition may be seen in how the brain encodes spatial locations.
+One example of how the full distribution over possible mental states can provide insights into cognition may be seen in how the brain encodes spatial locations.
 Theoretical~\citep[e.g., ][]{GersAbbo97} and empirical~\citep[e.g., ][]{PfeiFost13} work has demonstrated how place cell receptive fields may be modulated according to environmental features such as obstacles and the locations of rewarded goals.
@@
 It also provides potential insights into how the brain represents those stimuli, for example to the extent that the geometries of these manifolds may be leveraged to improve brain decoding.
-The shapes fo these manifolds can potentially provide insights into how our thoughts are distributed over representational spaces.
+The shapes of these manifolds can potentially provide insights into how our thoughts are distributed over representational spaces.
 

### 2019-01-14 tics_presubmission_inquiry.tex
incorporated TSM feedback, updated figs to PNGs to reduce file size, added presubmission inquiry letter

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}

\title{Episodic memory: mental time travel or a quantum `memory wave' function?}
\author{Jeremy R. Manning\\Dartmouth College, Hanover, NH\\jeremy.r.manning@dartmouth.edu}

\begin{document}
\section*{Proposed manuscript information}
\begin{itemize}
  \item \textbf{Title:} Episodic memory: mental time travel or a quantum `memory wave' function?
  \item \textbf{Author:} Jeremy R. Manning
  \item \textbf{Affiliation:} Department of Psychological and Brain Sciences, Dartmouth College
  \item \textbf{Email address:} jeremy.r.manning@dartmouth.edu
  \item \textbf{Phone number:} 603-646-2777
\end{itemize}

\section*{Proposed summary}
Where do we ``go'' when we recollect our past?  When remembering a past event, it is intuitive to imagine some part of ourselves mentally ``jumping back in time'' to when the event occurred. I propose an alternative view, inspired by recent evidence from my lab and others, as well as by re-examining existing models of episodic recall, that suggests that this notion of mentally revisiting any specific moment of our past is at best incomplete and at worst misleading.  Instead, I suggest that we retrieve information from our past by mentally casting ourselves back simultaneously to \textit{many} time points from our past, much like a quantum wave function spreading its probability mass over many possible states.  This revised conceptual model makes important behavioral and neural predictions about how we retrieve information about our past, and has implications for how we study episodic memory experimentally.

\section*{Timeliness of this work}
Over the past several years, the episodic memory literature has grown along several key fronts, ranging from better characterizations of the neural basis of the formation and retrieval of episodic memories, to advances in characterizations of episodic memory behaviors.  Alongside these advances has been a renewed focus on experiments that incorporate the use of complex ``naturalistic'' stimuli such as videos, natural images, and recorded speech.  Further, several recent studies have eschewed traditional trial-based designs in favor of less tightly controlled ``free form'' experiments.  Finally, advances in machine learning (especially deep learning) and related neuropsychological fields (especially spatial navigation) have lent some additional perspective to how information might be represented from a functional standpoint, and retreived during memory search.  Taken together, I believe it is time to reconsider how we characterize and interpret a number of old findings in the ``classic'' episodic memory literature.  Specifically, I propose that Tulving's notion of \textit{mental time travel} that has driven progress over the past several decades is ready to be refreshed in light of these new findings and approaches.

\nocite{AlyEtal18, BaldEtal17, BaldEtal18, BrunEtal18, ChenEtal17, FolkEtal18, LohnEtal18, MauEtal18, McKeBuzs16, Rang18, TsaoEtal18, BrigEtal18, GatyEtal16, IsolEtal17, JollChan18, LongKaha18, StYvNase18}

\renewcommand\refname{Key recent papers}
\bibliographystyle{apa}
\bibliography{CDL-bibliography/memlab}
\end{document}

### 2019-01-14 main.tex
added data and code availability statement

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 
+\section*{Acknowledgements}
+I thank Luke Chang, Janice Chen, Paxton Fitzpatrick, Andrew Heusser, Chris Honey, Talia Manning, Lucy Owen, and Kirsten Ziman for feedback and scientific discussions.
+I also thank Janice Chen, Yuan Chang Leong, Kenneth Norman, and Uri Hasson for sharing the data re-analyzed here, and Paxton Fitzpatrick and Andrew Heusser for contributing analysis code that I adapted from a prior publication.
+This work was supported in part by NSF EPSCoR Award Number 1632738.
+The content is solely my responsibility and does not necessarily represent the official views of my supporting organizations.
+
+\section*{Data and code availability}
+All analysis code and data used in the present manuscript may be found \href{https://github.com/ContextLab/mental-time-travel-paper}{\underline{here}}.
+
 \bibliographystyle{apa}

### 2019-01-14 main.tex
minor tweaks

@@
 For example, consider how we use pattern classifiers~\citep{NormEtal06b} to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, affective~\citep{ChanEtal18}, and social~\citep{MeyeEtal18} content being mentally manipulated.
-A prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.
+A prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} concept or mental state.
 Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states.
@@
 Although this view is captured mathematically by context-based theories of episodic memory, it is \textit{not} specifically captured by the conceptual notion of mental time travel that is often used to describe episodic recall.
-The alternative conceptualization presented here, that when we remember our past we reinstate thoughts associated with a weighted blend of our prior experiences, also has links with non-temporal or episodic domains such as semantic, spatial, and social representations.
+The alternative conceptualization presented here, that when we remember our past we reinstate thoughts associated with a weighted blend of our prior experiences, also has links with non-temporal or episodic domains such as semantic, spatial, affective, and social representations.
 Characterizing how our thoughts reflect \textit{distributions} over representational spaces is central to understanding how we think about ourselves, our experiences, other people, and the world around us.
@@
 \section*{Acknowledgements}
-I thank Luke Chang, Janice Chen, Paxton Fitzpatrick, Andrew Heusser, Chris Honey, Talia Manning, Lucy Owen, and Kirsten Ziman for feedback and scientific discussions.
+I thank Luke Chang, Janice Chen, Paxton Fitzpatrick, Andrew Heusser, Christopher Honey, Talia Manning, Lucy Owen, Matthijs van der Meer, and Kirsten Ziman for feedback and scientific discussions.
 I also thank Janice Chen, Yuan Chang Leong, Kenneth Norman, and Uri Hasson for sharing the data re-analyzed here, and Paxton Fitzpatrick and Andrew Heusser for contributing analysis code that I adapted from a prior publication.

### 2019-01-15 main.tex
minor tweaks

@@
 
-\subsection*{Extensions to spatial, semantic, affective, and social memory}
+\subsection*{Extensions to spatial, semantic, affective, and social representations}
 Just as we mentally spread ourselves across a distribution of times when we remember, we can perform similar mental operations when we conceptualize and retrieve other types of information.

### 2019-01-15 main.tex
added DuBrDava16 and MontEtal19 references

@@
 This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
-Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}.
+Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
 These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
@@
 The temporal dynamics of contextual change also underlie how we judge the amount of time elapsed since a prior reference point~\citep{BlocReed78, SahaSmit14}.
-These findings suggest that the rapid contextual or situational changes that define these event boundaries shape how we experience and remember, similarly to how spatial boundaries~\citep[e.g., environmental barriers;][]{McKeBuzs16, BrunEtal18} shape how we experience and remember spatial environments and layouts, or how conceptual boundaries~\citep[e.g., distinctions between semantic categories;][]{BrunEtal18} shape how we experience and remember conceptual information.
+These findings suggest that the rapid contextual or situational changes that define these event boundaries shape how we experience and remember~\citep{DuBrDava16}, similarly to how spatial boundaries~\citep[e.g., environmental barriers;][]{McKeBuzs16, BrunEtal18} shape how we experience and remember spatial environments and layouts, or how conceptual boundaries~\citep[e.g., distinctions between semantic categories;][]{BrunEtal18} shape how we experience and remember conceptual information.
 

### 2019-01-16 main.tex
working on cartoon figure
incorporated suggestions from KZ

@@
 Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.
-This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
+This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
 Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
@@
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
+Other recent work
+indicates that different aspects of the reinstatement process itself might also occur at multiple timescales.
+For example, \cite{LindEtal19} found that high-level semantic information was reinstated prior to low-level (e.g.
+perceptual) details in an associative memory task.
 Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.
@@
 \subsection*{How is our past replayed when we remember?}
-Our ongoing experiences can cue us to dredge up information about our past.
+Our ongoing experiences can cue us to remember information about our past.
 This can happen explicitly (e.g., during a conversation about a specific prior event) or implicitly (e.g., when something happening now reminds you of what happened before).
@@
 This suggests that our memory systems pick up on statistical regularities, even in ostensibly ``random'' stimulus sequences, and that we leverage those regularities when we recall our past~\citep{PolyEtal09}.
-Other work suggests that our memory systems may also leverage these regularities to parse our ongoing experiences into discrete events~\citep{SchaEtal13}.
+Other work suggests that our memory systems may also leverage these regularities to parse our ongoing experiences into discrete events~\citep{SchaEtal13, Shap19}.
 Are these examples in the literature on random list learning studies, indicating that our memory systems leverage ``accidental'' structure in random sequences, a reflection of the same fundamental principle driving the sorts of memory reactivation function displayed in Figure~\ref{fig:reinstatement}B?
@@
 \section*{Concluding remarks}
-When we revisit an experience from our past, we bring back thoughts related to a range of other semantically, conceptually, and functionally related experiences.
+When we remember an experience, we bring back thoughts related to a range of other semantically, conceptually, and functionally related experiences.
 This can serve an important functional role: it allows us to integrate information across temporally separated experiences.

### 2019-01-16 code/notebooks/reinstatement.py
file reorganization

@@ -1,833 +1,0 @@
-#%%
-
-import numpy as np
-import seaborn as sns
-import pandas as pd
-import matplotlib.pyplot as plt
-import matplotlib.patches as patches
-import matplotlib as mpl
-import scipy.signal as sgn
-import pickle
-import os
-import warnings
-import quail as q
-import hypertools as hyp
-import future
-from pathlib import Path
-import urllib
-from glob import glob
-from scipy.spatial.distance import cdist, pdist, squareform
-from scipy.interpolate import pchip
-from scipy.signal import find_peaks as localmax
-from scipy.stats import ttest_ind as ttest
-
-#%% 
-
-# ## Set path, load sherlock video and recall trajectories, load free recall data
-
-# In[3]:
-x = os.getcwd()
-
-rootdir = str(Path(notebook_path()).parent.parent.parent)
-datadir = os.path.join(rootdir, 'data')
-figdir = os.path.join(rootdir, 'data', 'figs')
-if not os.path.exists(figdir):
-    os.makedirs(figdir)
-video_model, recall_models = np.load(os.path.join(datadir, 'processed', 'models_t100_v50_r10.npy'))
-fr_random_data = q.load_egg(os.path.join(datadir, 'free recall', 'exp1.egg'))
-fr_cat_data = q.load_egg(os.path.join(datadir, 'free recall', 'exp2.egg'))
-
-
-
-
-# ## Plot figure
-
-# In[4]:
-
-
-h = sns.heatmap(pd.DataFrame(recall_models[0]).T.corr())
-
-
-# In[5]:
-
-
-def parse_diag(M):
-    n = M.shape[1]
-    x = np.nan * np.zeros([M.shape[0], 2 * n + 1])
-    
-    for i in np.arange(n):
-        x[i, (n - i):n] = M[i, 0:i]
-        x[i, n] = M[i, i]
-        x[i, (n+1):(n + n - i)] = M[i, (i+1):]
-    
-    return pd.DataFrame(x, columns=np.arange(-n, n+1))
-
-
-# In[6]:
-
-
-M = pd.DataFrame(recall_models[0]).T.corr().values
-np.fill_diagonal(M, np.nan)
-x = parse_diag(M)
-
-
-# In[7]:
-
-
-h = sns.heatmap(x)
-
-
-# In[295]:
-
-
-def plot_ribbon(x, color='k', ribbon_alpha=0.2):
-    mean = x.mean()
[…]

### 2019-01-16 paper/misc/psych_review_cover_letter.tex
added cover letter

\title{mental time travel letter}
%
% See http://texblog.org/2013/11/11/latexs-alternative-letter-class-newlfm/
% and http://www.ctan.org/tex-archive/macros/latex/contrib/newlfm
% for more information.
%
\documentclass[11pt,stdletter,orderfromtodate,sigleft]{newlfm}
\usepackage{blindtext, xfrac, animate, hyperref, pxfonts, geometry}

  \setlength{\voffset}{0in}

\newlfmP{dateskipbefore=0pt}
\newlfmP{sigsize=20pt}
\newlfmP{sigskipbefore=10pt}

\newlfmP{Headlinewd=0pt,Footlinewd=0pt}

\namefrom{\vspace{-0.3in}Jeremy R. Manning}
\addrfrom{
	Dartmouth College\\
    %Department of Psychological \& Brain Sciences\\
    %HB 6207 Moore Hall\\
	Hanover, NH  03755}

\addrto{}
\dateset{\today}

\greetto{To the editors of \textit{Psychological Review}:}

\closeline{Sincerely,}
%\usepackage{setspace}
%\linespread{0.85}
% How will your work make others in the field think differently and move the field forward?
% How does your work relate to the current literature on the topic?
% Who do you consider to be the most relevant audience for this work?
% Have you made clear in the letter what the work has and has not achieved?

\begin{document}
\begin{newlfm}
I have enclosed my manuscript entitled \textit{Episodic memory: mental time travel or a quantum `memory wave' function?} to be considered for publication as an \textit{Article}.  The manuscript provides a new perspective on what it means to remember.

Tulving (1983) described episodic memory as a sort of ``mental time travel,'' whereby we send some part of our mental state back along our autobiographical timeline to a specific moment from our past.  This view (and other similar perspectives) have shaped how memory research is carried out in the laboratory, and how results are interpreted.  In my manuscript I argue that this framing misses fundamental aspects of where we ``go'' when we remember.  Rather than revisiting any particular moment from our past recent neural, behavioral, and theoretical evidence suggests that remembering entails casting ourselves back to a \textit{distribution} of moments from our past.  This provides one means of integrating information acquired during experiences that are not temporally contiguous.  In addition to re-framing episodic memory in the context of the recent literature, I outline the implications of these ideas with respect to studying the neural and behavioral basis of episodic memory.  I also use several analyses of previously published data to provide concrete examples that permeate the manuscript.

I expect that this article will be of interest to memory researchers and theorists, as well as to scientists interested in how we distribute thoughts over representational spaces when we carry out a variety of mental operations.

Thank you for considering this manuscript, and I hope you will find it suitable for publication in \textit{Psychological Reveiew}.

\end{newlfm}
\end{document}

### 2020-09-09 paper/boneyard.tex
updated bibliography, working on text, adding new notebook with QWF vs MTT figure

--- 2019-01-10 boneyard.tex
@@
+For example, the \textit{remember-know paradigm}~\citep{Tulv85b} asks participants to distinguish a set of probe items according to whether they \textit{remember} specific (episodic) details about the experience of having studied each item previously in the experiment, or whether they feel that they \textit{know} generally that they had studied the item but are unable to recollect specific details.
+L
+
 \textbf{KEY QUESTIONS}:
@@
 
+% analysis idea: matrix of reinstatement functions for each timepoint-- e.g.
+considering each timepoint in turn (rows), how much are participants (on average) using language that mirrors how they talk about each other timepoint (columns).
+there should be a strong diagonal, with some blockiness, and possibly some disjoint groups showing how events cluster.
+
     Our brains exhibit autocorrelated activity.

### 2020-09-09 paper/main.tex
updated bibliography, working on text, adding new notebook with QWF vs MTT figure

--- 2019-01-16 main.tex
@@
 \usepackage{graphicx}
+\usepackage{setspace}
 \usepackage{hyperref}
+\usepackage{lineno}
+
+\doublespacing
+\linenumbers
 
@@
 How do our brains organize, retrieve, and act upon the ceaseless flow of incoming information and experiences?
-\cite{Tulv83} contends that episodic memory is like a sort of ``mental time travel,'' whereby we send some part of our mental state back along our autobiographical timeline to a specific moment in our past.
+\cite{Tulv83} contends that episodic memory is like a sort of ``mental time travel,'' whereby we send some part of our mental state back along our autobiographical timeline to a specific moment in our past~\citep[also see][]{Tulv72, HassEtal07b, SchaEtal07, MannEtal11}.
+This process has also been described as mentally ``jumping back in time''~\citep[e.g.,][]{HowaEtal12, FolkEtal18} to when we experienced what we are now remembering.
 This view has fundamental implications for how we study and build theories about memory: it suggests that we can match up specific moments of recollection with specific moments from the past.
@@
 Specifically, the primary implied questions are about how and when we match up the thoughts (and/or brain activity patterns) we are experiencing at one moment with the thoughts (and/or brain activity patterns) we experienced at another previous moment.
-The challenge, then, is to understand the neural (structural) and cognitive (functional) mechanisms that characterize when, how, and why our memory systems allow us to relive those prior experiences.
+The implied challenge, then, is to understand the neural (structural) and cognitive (functional) mechanisms that characterize when, how, and why our memory systems allow us to relive those prior experiences.
 One way of illustrating how this framing misses out on key aspects of memory is to examine how experimentalists and theorists typically conceive of and model a mental construct called \textit{context}.
@@
 Defining context precisely (e.g., as distinct from content) is an ongoing challenge for our field, but theorists generally describe context as reflecting a mental combination of external cues (where you are, who you are with, background sensory information, etc.) and internal cues (thoughts, feelings, emotions, etc.) that uniquely define each moment~\citep[e.g., see review by][]{MannEtal15}.
-Reactivating the mental representation of the context associated with a given episode is what can give us the feeling of some ``part of ourselves'' re-experiencing the past.
+Reactivating the mental representation of the context associated with a given episode is what can give us the feeling of some ``part of ourselves'' re-experiencing the past~\citep[Tulving refers to this feeling as \textit{autonoetic awareness};][]{Tulv02}.
 
@@
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
-Other recent work
-indicates that different aspects of the reinstatement process itself might also occur at multiple timescales.
+Other recent work indicates that different aspects of the reinstatement process itself might also occur at multiple timescales.
 For example, \cite{LindEtal19} found that high-level semantic information was reinstated prior to low-level (e.g.
@@
 This is not simply a matter of imprecision in mental time travel (i.e., a reflection of the low effective resolution with which we can revisit specific moments from our past).
-Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences.
+Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences, analogous to a quantum wave function spreading its probability mass over space.
 The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each drifting at a different timescale.
@@
 This is how we can integrate information across the \textit{many} discrete experiences relevant to ``now.''
+
+\section*{Defining the ``quantum memory wave function'' metaphor and distinguishing it from the ``mental time travel'' metaphor}
+Unlike in classical physics, where matter dutifully obeys Newton's laws, quantum physics postulates a different (though potentially complementary) set of rules governing how matter behaves and interacts.
+Perhaps the most famous equation in quantum physics is \cite{Schr26}'s equation describing the quantum \textit{wave function}~\citep[this finding was also inspired by][]{deBr24}.
+Schr\"{o}dinger's equation describes, in probabilistic terms, the location of a particle in space, at a given time.
+In other words, it defines a function, $\Psi(x, t)$, such that the probability that a particle is located at position $x$ at time $t$ is equal to $|\Psi(x, t)|^2$.
+The central idea is that, although it may not be possible to know or predict the particle's precise position at a given moment, the wave function tells us how it's location's probabilty mass extends over space.
+In turn, this may allow one to use the wave function to predict behaviors or constraints of the system(s) in which the particle is behaving and interacting.
+Within the domain of episodic memory, for the purposes of describing the \textit{quantum memory wave function metaphor}, we can conceptualize the moments along our \textit{autobiographical timeline} as the spatial coordinates ($x$) in Schr\"{o}dinger's equation.
+Then our mental state at time $t$ may be conceptualized as spreading its probability mass over our autobiographical timeline.
+In other words, like Schr\"{o}dinger's equation, we could in principle define a function describing how our mental state at time $t$ reflects the blends of thoughts we experienced at other moments of our lives.
+Under this metaphor, the effect of mental operations we perform with respect to our autobiographical timeline, such as remembering past experiences or predicting what will happen in the future, is to shift the probability mass of our quantum memory wave function (Fig.~\ref{fig:qwave}).
 

### 2020-09-11 paper/main.tex
added some notes and figure template to QMWF section
working on qmwf vs mtt section

@@
 Defining context precisely (e.g., as distinct from content) is an ongoing challenge for our field, but theorists generally describe context as reflecting a mental combination of external cues (where you are, who you are with, background sensory information, etc.) and internal cues (thoughts, feelings, emotions, etc.) that uniquely define each moment~\citep[e.g., see review by][]{MannEtal15}.
-Reactivating the mental representation of the context associated with a given episode is what can give us the feeling of some ``part of ourselves'' re-experiencing the past~\citep[Tulving refers to this feeling as \textit{autonoetic awareness};][]{Tulv02}.
+Reactivating the mental representation of the context associated with a given episode is what can give us the feeling of some ``part of ourselves'' re-experiencing the past~\citep[Tulving refers to this feeling as \textit{autonoetic awareness};][]{Tulv02a}.
 
-One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
+One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
 An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
 Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.
-This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12, AlyEtal18}.
+This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12a, AlyEtal18}.
 Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.
@@
 Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
-These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
+These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02a, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
@@
 
-\section*{Defining the ``quantum memory wave function'' metaphor and distinguishing it from the ``mental time travel'' metaphor}
+\section*{Defining the ``quantum memory wave function'' metaphor and comparing it to the ``mental time travel'' metaphor}
 Unlike in classical physics, where matter dutifully obeys Newton's laws, quantum physics postulates a different (though potentially complementary) set of rules governing how matter behaves and interacts.
@@
 In turn, this may allow one to use the wave function to predict behaviors or constraints of the system(s) in which the particle is behaving and interacting.
-Within the domain of episodic memory, for the purposes of describing the \textit{quantum memory wave function metaphor}, we can conceptualize the moments along our \textit{autobiographical timeline} as the spatial coordinates ($x$) in Schr\"{o}dinger's equation.
+Within the domain of episodic memory, for the purposes of describing the \textit{quantum memory wave function} metaphor, we can conceptualize the moments along our \textit{autobiographical timeline} as the spatial coordinates ($x$) in Schr\"{o}dinger's equation.
 Then our mental state at time $t$ may be conceptualized as spreading its probability mass over our autobiographical timeline.
@@
 Under this metaphor, the effect of mental operations we perform with respect to our autobiographical timeline, such as remembering past experiences or predicting what will happen in the future, is to shift the probability mass of our quantum memory wave function (Fig.~\ref{fig:qwave}).
+
+\begin{figure}[tp] \centering \caption{\textbf{Memory wave functions.
+All panels.} Each panel displays a quantum memory wave function for a hypothetical individual at a particular moment (time $t = 0$).
+The $x$-axes correspond to the position ($t$) along an autobiographical timeline.
+The $y$-axes display how much the individual's thoughts at time $t = 0$ include information about the individual's past (or future) perceived or predicted experiences at other times.
+\textbf{A.
+Contextual drift (no explicit reinstatement).}  The blend of thoughts leading up to $t = 0$ reflects a recency-weighted average of information from the past.
+\textbf{B.
+Mental time travel.} This constitutes a special case of the quantum memory wave function framework, whereby the individual's thoughts reflect a recency-weighted blend of prior thoughts, along with thoughts specific to one remembered prior experience (occuring at the timepoint indicated by the $\uparrow$ symbol).
+Each curve's shading varies with the amount of temporal blurring.
+In this example, greater amounts of blurring denotes that the individual is bringing to mind information that is farther from the remembered moment or from $t = 0$.
+\textbf{C.
+Quantum memory waves.}  In this example, the individual remembering the moment indicated by the $\uparrow$ symbol brings back a blend of thoughts similar to those shown in Panel B.
+In addition, they bring to mind information pertaining to other conceptually related (but temporally non-contiguous) moments, indicated by $\star$ symbols.
+As in Panel B, different curves are shaded according to their degree of temporal blurring.
+\textbf{D.
+Generating predictions about the future.} To predict what will happen during a future event ($\diamond$) the individual spreads their thoughts over their prior experiences according to the expected relevance of those experiences to the future event.
+Local peaks in the curve denote prior experiences with high expected relevance.
+The black part of the curve illustrates how the individual's thoughts are spread out over their past, and the gray part of the curve illustrates how the individual's thoughts are spread out over their (predicted) future.}
+\label{fig:qwave}
+\end{figure}
+
+The \textit{mental time travel} metaphor is defined by Tulving as ``the type of awareness experienced when one thinks back to a \textit{specific moment} in one's personal past and consciusly recollects some prior episode or state as it was previously experienced''~\citep[Fig.~\ref{fig:qwave}B][emphasis added]{WheeEtal97}.
+Whereas mental time travel refers to mentally revisiting to one specific moment in addition to now (potentially with some temporal blurring), the quantum memory wave framework implies that episodic memory entails spreading our awareness over \textit{many} moments (Fig.~\ref{fig:qwave}C).
+In this way, the key distinction between these framings is whether remembering entails simultaneously revisiting one single moment versus many moments.
+The two frameworks converge when we engage in \textit{chronesthesia}~\citep{Tulv02b}-- i.e., when we are consciously aware of subjective time.
+Analogous to how measuring the physical location of a particle collapses its quantum wave function at the moment of measurement, engaging in chronesthesia can collapse our quantum memory wave function such that our thoughts temporarily converge on the moment we are revisiting.
+This wave function expands again when our mental state changes in the next moment.
+
+The quantum memory wave and mental time travel metaphors also appear to diverge in explaining how we might predict what will happen in the future.
+The quantum memory wave framing suggests that our predictions about the future draw from information pertaining to a blend of our prior experiences, weighted by their expected relevance to the future event in question (Fig.~\ref{fig:qwave}D.
+To the extent that the mental time travel metaphor assumes that we can mentally visit just one other time (or time range) in addition to now~\citep[e.g.,][]{SzpuTulv11}, it is not clear how this integration over temporally separated prior experiences might occur.
+Specifically, integrating over those experiences would seem to require some mechanism for associating and simultaneously revisiting them while also casting some part of our thoughts forward to the future.
 
@@
 Further, these distributions may be estimated using existing methods.
-For example, consider how we use pattern classifiers~\citep{NormEtal06b} to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, affective~\citep{ChanEtal18}, and social~\citep{MeyeEtal18} content being mentally manipulated.
+For example, consider how we use pattern classifiers~\citep{NormEtal06b} to connect neural activity patterns to specific semantic~\citep{PolyEtal05a, MitcEtal08a, MannEtal12}, spatial~\citep{MillEtal13}, affective~\citep{ChanEtal18}, and social~\citep{MeyeEtal18} content being mentally manipulated.
 A prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} concept or mental state.

### 2020-09-14 paper/main.tex
working on qwf vs mtt notebook

@@
 \section*{Acknowledgements}
-I thank Luke Chang, Janice Chen, Paxton Fitzpatrick, Andrew Heusser, Christopher Honey, Talia Manning, Lucy Owen, Matthijs van der Meer, and Kirsten Ziman for feedback and scientific discussions.
-I also thank Janice Chen, Yuan Chang Leong, Kenneth Norman, and Uri Hasson for sharing the data re-analyzed here, and Paxton Fitzpatrick and Andrew Heusser for contributing analysis code that I adapted from a prior publication.
+I thank Luke Chang, Janice Chen, Paxton Fitzpatrick, Andrew Heusser, Christopher Honey, Caroline Lee, Talia Manning, Lucy Owen, Matthijs van der Meer, and Kirsten Ziman for feedback and scientific discussions.
+I also thank Janice Chen, Yuan Chang Leong, Christopher Honey, Chung Yong, Kenneth Norman, and Uri Hasson for sharing the data re-analyzed here~\citep{ChenEtal17}, and Paxton Fitzpatrick and Andrew Heusser for contributing analysis code that I adapted from a prior publication~\citep{HeusEtal18c}.
 This work was supported in part by NSF EPSCoR Award Number 1632738.

### 2020-09-16 paper/main.tex
added section on matching problem, working on word embedding section

@@
 
-\begin{figure}[tp] \centering \caption{\textbf{Memory wave functions.
-All panels.} Each panel displays a quantum memory wave function for a hypothetical individual at a particular moment (time $t = 0$).
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/wave_functions}
+  \caption{\textbf{Memory wave functions.
+All panels.} Each panel displays a quantum memory wave function for a hypothetical individual at a particular moment (time $t=0$ indicated by $\blacktriangle$)).
 The $x$-axes correspond to the position ($t$) along an autobiographical timeline.
 The $y$-axes display how much the individual's thoughts at time $t = 0$ include information about the individual's past (or future) perceived or predicted experiences at other times.
+All curves have been normalized to have a maximum value of 1 and a minimum value of 0.
 \textbf{A.
@@
 \textbf{B.
-Mental time travel.} This constitutes a special case of the quantum memory wave function framework, whereby the individual's thoughts reflect a recency-weighted blend of prior thoughts, along with thoughts specific to one remembered prior experience (occuring at the timepoint indicated by the $\uparrow$ symbol).
-Each curve's shading varies with the amount of temporal blurring.
+Mental time travel.} This constitutes a special case of the quantum memory wave function framework, whereby the individual's thoughts reflect a recency-weighted blend of prior thoughts, along with thoughts specific to one remembered prior experience (occuring at the timepoint indicated by the $\bigstar$ symbol).
+Each curve's coloring denotes the amount of temporal blurring.
 In this example, greater amounts of blurring denotes that the individual is bringing to mind information that is farther from the remembered moment or from $t = 0$.
 \textbf{C.
-Quantum memory waves.}  In this example, the individual remembering the moment indicated by the $\uparrow$ symbol brings back a blend of thoughts similar to those shown in Panel B.
+Quantum memory waves.}  In this example, the individual remembering the moment indicated by the $\bigstar$ symbol brings back a blend of thoughts similar to those shown in Panel B.
 In addition, they bring to mind information pertaining to other conceptually related (but temporally non-contiguous) moments, indicated by $\star$ symbols.
-As in Panel B, different curves are shaded according to their degree of temporal blurring.
+As in Panel B, different curves are colored according to their degree of temporal blurring.
 \textbf{D.
-Generating predictions about the future.} To predict what will happen during a future event ($\diamond$) the individual spreads their thoughts over their prior experiences according to the expected relevance of those experiences to the future event.
+Generating predictions about the future.} To predict what will happen during a series of future events ($\diamond$ symbols) the individual spreads their thoughts over their prior experiences according to the expected relevance of those experiences to the future events.
 Local peaks in the curve denote prior experiences with high expected relevance.
-The black part of the curve illustrates how the individual's thoughts are spread out over their past, and the gray part of the curve illustrates how the individual's thoughts are spread out over their (predicted) future.}
+The left part of the panel (prior to $\blacktriangle$) illustrates how the individual's thoughts are spread out over their past, and the right part of the panel illustrates how the individual's thoughts are spread out over their (predicted) future.}
 \label{fig:qwave}
@@
 The authors used topic models~\citep{BleiEtal03} applied to human-generated annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.
-These \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode.
+Each shot's annotation contained the following information: a brief narrative description of what was happening, the location where the scene took place, whether that location was indoors or outdoors, the names of all on-screen characters, the names of the character(s) who were in focus in the camera shot, the name(s) of any character(s) who spoke during the scene, the camera angle of the shot, a transcription of any on-screen text, and whether or not there was any music present in the auditory background~\citep[full details may be found in][]{ChenEtal17}.
+The resulting \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode.
 When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change.
@@
 
-\begin{figure}[tp] \centering \includegraphics[width=0.8\textwidth]{figs/pres_rec_corrmats_sherlock.png} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal18c}, where full details may be found.
+\begin{figure}[tp] \centering \includegraphics[width=0.8\textwidth]{figs/pres_rec_corrmats_sherlock} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal18c}, where full details may be found.
 \textbf{A.
@@
 Specifically, their language mirrored that used when they were recalling other murders, and other moments when Sherlock and John were bantering with or about each other (Fig.~\ref{fig:reinstatement}C).
+
 The way they described the example event did \textit{not} mirror the way they described other semantically (but not conceptually) related events, such as events involving unrelated violence, or Sherlock and John interacting with other characters about other subjects.
@@
 
-\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement.png} \caption{\textbf{Content drift in a television episode and participants' recalls of that episode.
+One concrete example of this phenomenon is that, early in the episode, Sherlock and John appear to be investigating a series of suicides.
+Later in the episode, it becomes clear that the apparent suicides are actually murders.
+The annotations refer to early deaths as suicides, whereas participants often refer to those early deaths as ``murders'' even though they could not have known that they were murders early on in the episode.
+In this way, participants' experiences \textit{subsequent to} a target remembered event may distort its memory by incorporating information learned during other experiences.
+
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Content drift in a television episode and participants' recalls of that episode.
 A.
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figs/trajectory.png}
+\includegraphics[width=0.8\textwidth]{figs/trajectory}
 \caption{\textbf{The shape of an unfolding experience.}  \textbf{A.
@@
 Although there are substantial individual differences in the specific words people use to describe their shared experience of watching the same television episode, and how much detail they recollected, the overall shapes of those recalled experience trajectories are highly similar across people.
-This has potentially exciting implications for what it means to communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.
+This has potentially exciting implications for what it means to communicate memories to other people: we need to convey the ``essence'' (shape) of our experience.
 It also has implications for how we learn by analogy: perhaps events that share schemas~\citep[e.g., ][]{BaldEtal18}  unfold in similar ways (e.g., have similar shapes).
 Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).
+
+\subsection*{The \textit{matching problem}}
+The trajectories through word embedding space displayed in Figure~\ref{fig:trajectories} demonstrate that topic models are sufficiently expressive and robust to provide one solution to the \textit{matching problem} in memory analysis.
+The matching problem refers to discovering an alignment in the labels between experienced and remembered events.
+Traditional memory paradigms most commonly employed over the past century~\citep{Kaha12} ask participants to study lists of \textit{items} (often randomly selected words) and then engage in a memory task, such as verbally recalling any words the participant remembers studying, in any order.
+The matching problem in such experiments is trivial: if we treat each studied word on a list as an ``event,'' then we can assume that when the participant verbalizes that word during their memory test, they are remembering that word's event.
+In other words, matching entails finding the exact match between a word the participant is verbalizing and that same word on the list that they studied.
+
+For real-world memory phenomena, and in many so-called \textit{naturalistic} memory paradigms, solving the matching problem is more complex.
+People often experience real-world events (and scenes from movies or stories) as rich sensory experiences that derive deep meaning from surrounding moments, other prior experiences, emotions, goals, high-order understandings of the current situation, and so on.
+Many of these aspects of our real-world experiences are not immediately verbalizable-- and even when we are able to describe some specific aspects of what we are experiencing, there are typically many possibly ways we could have done so (e.g., varying in level of detail, specific wording, objectivity versus subjectivity, contextual details, etc.).
+Further, because the way we experience each moment depends in part on our unique prior experiences, even ostensibly shared experiences can seem different to each individual.
+Similarly, the way we recount a given experience is often a reflection of our own idiosyncratic experience of the original event, colored and distorted in an idiosyncratic way by our memory systems.
+This distorted memory must then be verbalized or communicated, which is itself an idiosyncratic (and potentially lossy in the information theoretic sense) process.
+Solving the matching problem under these constraints requires building explicit models of the dynamic conceptual content of our experiences and memories, and then estimating when and how that experienced and remembered content aligns.
+The approach featured in Figure~\ref{fig:trajectories} uses a word embedding space that is defined by a topic model~\citep{BleiEtal03}.
+Each coordinate in this space reflects the conceptual content underlying a set of words, such that two different text passages that describe the same conceptual content (but using different words, levels of detail, etc.) are mapped on to nearby points in the word embedding space.
+When these models are sufficiently expressive and robust to model our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal18c}.
+
+\subsection*{Defining the geometries of ongoing thoughts and memories}
+%JRM STOPPED HERE
 

### 2020-09-18 paper/main.tex
working on word embedding text

@@
 \subsection*{Defining the geometries of ongoing thoughts and memories}
-%JRM STOPPED HERE
+Defining geometric spaces that can appropriately capture the conceptual contents of our thoughts and experiences has become a major focus of computational science.
+An class of models that have shown particular promise over the past several decades is broadly referred to as \textit{word embedding models} (Fig.~\ref{fig:embedding-models}).
+The fundamental goal of these models is to take as input a sequence of words (text) and produce as output one or more feature vectors whose coordinates reflect their conceptual (semantic) content.
+For example, when two similar documents are passed through these models, the goal is to produce two feature vectors that are closer (e.g., in Euclidean distance) than the feature vectors for two dissimilar documents.
+
+Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{LandDuma97} attempted to compute these embeddings using word co-occurance statistics.
+The logic driving these approaches posited that documents that contained similar mixtures of words would be likely to be conceptually similar, and words that appeared in similar documents would also likely be conceptually similar.
+To quantify these intuitions, \cite{LandDuma97} first constructed \textit{word count matrices} (Fig.~\ref{fig:embedding-models}A) for several text corpora (each comprising many thousands of documents).
+Each entry in these matrices denotes the number of time a given document (row) contained a given word (column).
+They then used the Singular Value Decomposition (SVD) to decompose the number-of-documents by number-of-words word count matrix into the product of a number-of-documents $\times K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values} matrix, and a $K \times $ number-of-unique-words \textit{word embeddings matrix}.
+The rows of the document embeddings matrix reflected the similarities between documents, and the columns of the word embeddings matrix reflected similarities between words.
+
+A related approach, termed \textit{Word Association Spaces}~\citep[WAS;][]{NelsEtal04, SteyEtal04} replaces the word count matrix with data from a large free-association experiment.
+In the experiment, each participant was asked to free-associate (i.e., say the first word that came to mind) in response to a series of cue words.
+The authors then constructed a number-of-cue-words by number-of-unique-response-words matrix whose entries reflected the proportion of times each cue word (row) produced a given response word (column) as an associate.
+After using SVD to decompose this \textit{word frequency matrix}, they obtained low-dimensional embeddings of each cue word and response word.
+In one sense, WAS provides a closer reflection of how people think about the meanings of words than LSA, since nearby WAS vectors reflect either cue words that produced similar responses, or response words that were evoked by similar cues.
+In other words, WAS derives embeddings directly from people's semantic judgement behaviors.
+However, because WAS is based on a large amount of experimental data, updating the embeddings requires collecting new data, which can be slow and costly.
+As the target application or population diverges from the experimental population (e.g., over time, across cultures, demographics, etc.) the WAS embeddings may become less accurate or relevant.
+By contrast, LSA may be re-fit relatively quickly to a new corpus, enabling the resulting embeddings to stay current.
+Further, because LSA may be applied to any corpus, word embeddings may be easily computed and/or compared across languages, times, subject areas, etc.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.7\textwidth]{figs/word_embedding_models}
+\caption{\textbf{Three approaches to text embedding.}  \textbf{A.
+Matrix factorization models.} Per-document (or per-participant) word counts and/or frequencies are collected into a word count matrix (left).
+To compute $K$-dimensional embeddings of the documents and words, the word count matrix is factorized (typically by computing its Singular Value Decomposition) into the product of a number-of-documents $\times K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values} matrix, and a $K \times $ number-of-unique-words \textit{word embeddings matrix}.
+The rows of the document embeddings matrix reflect the similarities between documents, and the columns of the word embeddings matrix reflect similarities between words.
+\textbf{B.
+Topic models.}  Fitting a topic model to a word count matrix (e.g., Panel A) reveals a set of latent \textit{topics}, or themes that pervade each document.
+The topics are specified in a $K \times$ number-of-words \textit{topics matrix}, and the per-document themes are specified in the number-of-documents $\times K$ \textit{topic proportions matrix}.
+A given document's topic proportions reflect the specific mix of themes that are reflected in that document (bar graph on the left).
+These proportions are discovered automatically by estimating the topic from which each word in the document is most probably drawn (highlighted words in \textit{Document text}).
+\textbf{C.
+Deep learning models.}  A wide variety of deep learning approaches to text embedding have been proposed.
+Broadly, these approaches often combine a common set of elements.
+\textit{Transformer} models take a sequence as input and output a new modified sequence~\citep[diagram adapted from][]{ViswEtal17}.
+The transformed outputs distort the original inputs in a way that often enables subseqeuent components of a deep text embedding model to compute more stable and meaningful embeddings.
+\textit{Deep averaging networks} pool information from different observations in the original data in order to magnify small but meaningful differences.
+\textit{Recurrent neural networks} (and variants) are often used to model temporal relations in sequence data~\citep[the deep averaging network and recurrent neural network diagrams are loosely inspired by figures from][]{IyyeEtal15}.
+\textit{Deep autoencoders} attempt to reconstruct inputted patterns after passing them through a series of increasingly smaller layers in an encoding network.
+The resulting activation patterns in the discriminator layer, which contains the fewest number of neurons, may be used as a low-dimensional embedding of the input.
+The discriminator's activations are then passed through a decoder network which contains successively larger layers.
+Finally, a reconstructed output is produced.
+After fitting this model, outputs may be estimated from arbitrary activations of the discriminator layer~\citep[diagram loosly inspired by][]{YousHame17}.}
+\label{fig:embedding-models}
+\end{figure}
+
+% topic models-- deeper themes, stress interpretability.
+example: bat (animal, sports equipment, verb).
+"As the mutant vampire bat swooped low to attack, the explorer took a mighty swing of her trusty baseball bat.
+The swing connected, stunning the creature before it scrambled about on the ground and flew away.
+She didn't even bat an eye-- as the dark forces grew in power, such things had become increasingly common, nearly to the point of mundanity."
+
+%deep learning approaches
 

### 2020-09-19 paper/main.tex
working on word embedding section

@@
 
-% topic models-- deeper themes, stress interpretability.
-example: bat (animal, sports equipment, verb).
-"As the mutant vampire bat swooped low to attack, the explorer took a mighty swing of her trusty baseball bat.
+\textit{Topic models}, such as \textit{Latent Dirichlet Allocation}~\citep[LDA;][]{BleiEtal03} represent an attempt to characterize the latent \textit{topics} or themes that documents comprise (Fig.~\ref{fig:embedding-models}B).
+Whereas matrix factorization approaches represent semantic meanings using low-dimensional approximations of word co-occurances or free association frequencies, topic models leverage these properties of documents (or behaviors) to explain \textit{why} words co-occur.
+LDA defines a topic as a distribution of weights over words in the vocabulary.
+For example, if one takes the \textit{vocabulary} to be the top 10,000 most commonly used words in English (excluding stop words that are devoid of explicit semantic meaning such as ``and,'' ``or,'' ``the,'' ``but,'' etc.), then each topic would comprise a vector of 10,000 numbers (weights) between zero and one, inclusive, where the full set of weights summed to one.
+As shown in the cartoon on the left of the Panel (where the four top-weighted words from each of five topics are displayed), a fruit-related topic (shown in red) might weight words like ``apple,'' ``banana,'' ``pear,'' and ``orange'' heavily, whereas words like ``cat,'' ``dog,'' ``dolphin,'' and ``lobster'' might receive relatively small weights.
+Given a set of $K$ topic vectors, topic models compute the blend of topics most probably reflected in a given document-- i.e., its \textit{topic proportions}.
+These topic proportion estimates are made by examining how different topics weight each of the words in the document, and by assuming that documents tend to reflect few topics (rather than many topics).
+Given only a word count matrix as input (just as in LSA), topic models automatically discover a $K$ by number-of-words \textit{topics matrix}, which describes how each topic (row) weights each word in the vocabulary (column), and a number-of-documents by $K$ \textit{topic proportions matrix}, which describes each document's topic proportions.
+The rows of the topic proportions matrix may be treated as low-dimensional embeddings of the set of documents.
+
+A key advantage that topic models have over matrix factorization models is the ability to explain potentially ambiguous word usage.
+For example, consider the use of the word ``bat'' in the following passage:
+\begin{quote}
+  \textit{``...As the mutant vampire bat swooped low to attack, the explorer took a mighty swing of her trusty baseball bat.
 The swing connected, stunning the creature before it scrambled about on the ground and flew away.
-She didn't even bat an eye-- as the dark forces grew in power, such things had become increasingly common, nearly to the point of mundanity."
+She didn't even bat an eye-- as the dark forces grew in power, such things had become increasingly common, nearly to the point of mundanity...''}
+  \end{quote}
+At different ponts in the passage, ``bat'' refers to an animal, a piece of sports equipment, and a verb.
+Because each of these uses of the word appear in the same document, models driven solely by word co-occurance statistics will blend these different meanings such that the embedding of ``bat'' is somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
+By contrast, treating each document as reflecting a blend of topics allows these uses to be disambiguated.
+If other documents that mention the word ``bat'' describe vampires, swooping, or other mammals, but \textit{not} sports equipment or eye-related verbs, then a topic model could explain this observation by inferring that multiple topics place substantial weight on the word ``bat.''  Another strength of topic models is their interpretability.
+Because the models directly output each document's topic proportions along with each topic's weights on all words in the vocabulary, the general themes that each topic reflects are often intuitive and the per-document topic proportions may be directly related to those themes.
 
-%deep learning approaches
+% deep learning approaches
+% bag of words vs. sequence models, context and grammar to convey meaning, etc.
 

### 2020-09-21 paper/boneyard.tex
working on qmwf vs mtt section
finished draft of section on word embedding models

@@
 L
+
+A nuance of the quantum memory wave function framework is that,
+
+However, because the thoughts we have at a given moment may themselves draw on other moments of our past,
+
+The central limitation of the mental time travel metaphor that the quantum memory wave function addresses is that
+
+  .
+The transformed outputs distort the original inputs in a way that often enables subseqeuent components of a deep text embedding model to compute more stable and meaningful embeddings.
+\textit{Deep averaging networks} pool information from different observations in the original data in order to magnify small but meaningful differences.
+\textit{Recurrent neural networks} (and variants) are often used to model temporal relations in sequence data~\citep[the deep averaging network and recurrent neural network diagrams are loosely inspired by figures from][]{IyyeEtal15}.
+\textit{Deep autoencoders} attempt to reconstruct inputted patterns after passing them through a series of increasingly smaller layers in an encoding network.
+The resulting activation patterns in the discriminator layer, which contains the fewest number of neurons, may be used as a low-dimensional embedding of the input.
+The discriminator's activations are then passed through a decoder network which contains successively larger layers.
+Finally, a reconstructed output is produced.
+After fitting this model, outputs may be estimated from arbitrary activations of the discriminator layer~\citep[diagram loosly inspired by][]{YousHame17}.
 

### 2020-09-21 paper/main.tex
finished draft of section on word embedding models

@@
 Defining geometric spaces that can appropriately capture the conceptual contents of our thoughts and experiences has become a major focus of computational science.
-An class of models that have shown particular promise over the past several decades is broadly referred to as \textit{word embedding models} (Fig.~\ref{fig:embedding-models}).
+One family of models that has shown particular promise over the past several decades is broadly referred to as \textit{word embedding models} (Fig.~\ref{fig:embedding-models}).
 The fundamental goal of these models is to take as input a sequence of words (text) and produce as output one or more feature vectors whose coordinates reflect their conceptual (semantic) content.
@@
 
-Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{LandDuma97} attempted to compute these embeddings using word co-occurance statistics.
+Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{DeerEtal90, LandDuma97} attempted to compute these embeddings using word co-occurance statistics.
 The logic driving these approaches posited that documents that contained similar mixtures of words would be likely to be conceptually similar, and words that appeared in similar documents would also likely be conceptually similar.
@@
 \centering
-\includegraphics[width=0.7\textwidth]{figs/word_embedding_models}
-\caption{\textbf{Three approaches to text embedding.}  \textbf{A.
-Matrix factorization models.} Per-document (or per-participant) word counts and/or frequencies are collected into a word count matrix (left).
+\includegraphics[width=0.75\textwidth]{figs/word_embedding_models}
+\caption{\textbf{Three approaches to text embedding.}  \textbf{A.~Matrix factorization models.} Per-document (or per-participant) word counts and/or frequencies are collected into a word count matrix (left).
 To compute $K$-dimensional embeddings of the documents and words, the word count matrix is factorized (typically by computing its Singular Value Decomposition) into the product of a number-of-documents $\times K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values} matrix, and a $K \times $ number-of-unique-words \textit{word embeddings matrix}.
-The rows of the document embeddings matrix reflect the similarities between documents, and the columns of the word embeddings matrix reflect similarities between words.
-\textbf{B.
-Topic models.}  Fitting a topic model to a word count matrix (e.g., Panel A) reveals a set of latent \textit{topics}, or themes that pervade each document.
+\textbf{B.~Topic models.}  Fitting a topic model to a word count matrix (e.g., Panel A) reveals a set of latent \textit{topics}, or themes that pervade each document.
 The topics are specified in a $K \times$ number-of-words \textit{topics matrix}, and the per-document themes are specified in the number-of-documents $\times K$ \textit{topic proportions matrix}.
@@
 These proportions are discovered automatically by estimating the topic from which each word in the document is most probably drawn (highlighted words in \textit{Document text}).
-\textbf{C.
-Deep learning models.}  A wide variety of deep learning approaches to text embedding have been proposed.
-Broadly, these approaches often combine a common set of elements.
-\textit{Transformer} models take a sequence as input and output a new modified sequence~\citep[diagram adapted from][]{ViswEtal17}.
-The transformed outputs distort the original inputs in a way that often enables subseqeuent components of a deep text embedding model to compute more stable and meaningful embeddings.
-\textit{Deep averaging networks} pool information from different observations in the original data in order to magnify small but meaningful differences.
-\textit{Recurrent neural networks} (and variants) are often used to model temporal relations in sequence data~\citep[the deep averaging network and recurrent neural network diagrams are loosely inspired by figures from][]{IyyeEtal15}.
-\textit{Deep autoencoders} attempt to reconstruct inputted patterns after passing them through a series of increasingly smaller layers in an encoding network.
-The resulting activation patterns in the discriminator layer, which contains the fewest number of neurons, may be used as a low-dimensional embedding of the input.
-The discriminator's activations are then passed through a decoder network which contains successively larger layers.
-Finally, a reconstructed output is produced.
-After fitting this model, outputs may be estimated from arbitrary activations of the discriminator layer~\citep[diagram loosly inspired by][]{YousHame17}.}
+\textbf{C.~Deep learning models.}  A wide variety of deep learning approaches to text embedding have been proposed.
+Broadly, these approaches often combine a common set of elements, including \textit{transformers}~\citep[diagram adapted from][]{ViswEtal17}, \textit{deep averaging networks}~\citep[diagram adapted from][]{IyyeEtal15}, \textit{recurrent neural networks}~\citep[diagram adapted from][]{IyyeEtal15}, and \textit{deep autoencoders}~\citep[diagram adapted from][]{YousHame17}.}
 \label{fig:embedding-models}
@@
 
-% deep learning approaches
-% bag of words vs. sequence models, context and grammar to convey meaning, etc.
+Both the matrix factorization models and topic model described above are each examples of \textit{bag of words} models.
+Given a document, these models will generate the same embeddings for any random re-ordering of the same words.
+This over-simplifies the intended conceptual content of documents, where word order often conveys important meaning.
+An alternative framing of the text embedding problem is to use \textit{sequence} models that explicitly account for word order effects.
+Most modern sequence-based models are implemented using deep neural networks (Fig.~\ref{fig:embedding-models}C).
+
+The Neural Probabilistic Language Model~\citep[NPLM;][]{BengEtal03} was one of the earliest deep learning-based sequence models.
+NLPM derives word embeddings for each word in the vocabulary (such that nearby embeddings reflect semantically related words), and then models word sequences according to their meanings.
+This model was inspired in part by $n$-gram models~\citep{BrowEtal92}, which estimate the probability of the next word in a sequence given the $n$ words that came before it.
+Effectively, NLPM attempts to learn arbitrarily long sequences.
+In a related approach, \cite{MnihHint09} proposed the Scalable Hierarchical Distributed Language Model, which learns sequences of $n$-grams to achieve greater explanatory power.
+
+The next major advance in word embeddings came from attempts to more precisely capture the geometric relations between words.
+For example, if each word's semantic meaning is captured by a feature vector, then in principle one could perform vector algebra on different words' embeddings to produce new concepts that combine their meanings.
+For example, subtracting the vector for ``man'' from that of ``king'' and then adding the vector for ``woman'' would result in a new feature vector similar to that of ``queen''~\citep{MikoEtal13b}.
+\cite{MikoEtal13a} developed the deep learning model \textit{word2vec} to explicitly optimize the accuracy of these vector operations.
+Whereas word2vec is a bag of words model, Global Vectors for Word Representation~\citep[GloVe;][]{PennEtal14} extends this approach to model each word's local (temporal) context~\citep[also see][]{PeteEtal18}.
+
+Several recent models have begun to move beyond single word-level embeddings.
+For example, SkipThought~\citep{KiroEtal15}, Universal Sentence Encoder~\citep{CerEtal18}, InferSent~\citep{ConnEtal18}, and RandSent~\citep{WietKiel19} each compute sentence-level embeddings.
+The central insight of these approaches is that capturing the nuanced meaning of a particular concept often requires \textit{several} words that have been organized into a complete sentence or sentence-like grammatical construct.
+
+Cutting-edge approaches to word embeddings that have achieved the best-to-date performance on modern natural language processing benchmarks are built using enormous deep neural networks often with many billions of parameters.
+Because these models are very computationally expensive to fit, an important discovery has been that many layers in these models may be pre-trained without substantially affecting performance, even on new tasks.
+One of the first of this new class of model, Bidirectional Encoder Representations from Transformers~\citep[BERT;][]{DevlEtal18}, uses a pre-trained deep neural network, whose weights in a single output layer may be fine-tuned to suit a wide range of new tasks.
+The latest additions to this class of models are called Generative Pre-Trained (GPT) transformers~\citep{RadfEtal18, RadfEtal19, BrowEtal20}.
+
+% deep learning approaches:
+% - bag of words vs. sequence models vs. sentence-level embeddings
+% - accounting for use of context and grammar to convey meaning
+% - pre-training vs. customized training vs. training-free
+
+% Neural Probabilistilc Language Model (NPLM): BengEtal03 --> explicitly model word sequences
+% Scalable Hierarchical Distributed Language Model (SHDLM): MnihHint09 --> sequences of word trees
+
+%word2vec: MikoEtal13 --> semantic relations reflected in geometric operations
+%GloVe: PennEtal14 --> combines matrix factorization and local context window methods
+%SkipThought: KiroEtal15 --> reconstructs surrounding sentences of an encoded passage
+%BERT: DevlEtal18 --> pretraining language representations, context dependence of word embeddings
+%ELMO: PeteEtal18 --> how is it different from BERT/GPT?
+%InferSent: ConnEtal18 --> embeddings of sentences
+%Universal Sentence Encoder: CerEtal18 --> embeddings of sentences
+%Open-GPT: RadfEtal18 --> pretrained transformer (how is it different from BERT?)
+%RandSent: WietKiel19 --> training-free sentence-level embeddings
+%GPT-2: RadfEtal19
+%GPT-3: BrowEtal20
 

### 2020-09-22 paper/main.tex
minor updates to QWF section
finished draft addressing r1 comments, working on r2 comments [WIP]

@@
 Under this metaphor, the effect of mental operations we perform with respect to our autobiographical timeline, such as remembering past experiences or predicting what will happen in the future, is to shift the probability mass of our quantum memory wave function (Fig.~\ref{fig:qwave}).
+This metaphor is also related to work on quantum probability decision models~\citep[e.g.,][]{KhreEtal18} in that it extends similar concepts to the domain of episodic memory.
 
@@
 
-The way they described the example event did \textit{not} mirror the way they described other semantically (but not conceptually) related events, such as events involving unrelated violence, or Sherlock and John interacting with other characters about other subjects.
+The way they described the example target event did \textit{not} mirror the way they described other events that were not directly narratively linked with the target event, even when those other events overlapped semantically.
+For example, thematic elements of events involving unrelated violence, or Sherlock and John interacting with other characters about other subjects, tended not to feature in participants' recountings of the target event.
 Taken together, this suggests that these specific events were associated in participants' memories, perhaps contributing to (or reflecting) their understanding that those events were linked in the episode's narrative.
@@
 
+In their paper, \cite{HeusEtal18c} use the geometric shapes of experience trajectories (through word embedding space) to functionally distinguish high-level essential details of the experience from low-level (non-essential) details.
+When the trajectory exhibits a sharp change in direction, this indicates a rapid shift in the thematic content of the episode.
+\cite{HeusEtal18c} used hidden Markov models to detect these rapid shifts and segment the trajectories into discrete \textit{events}.
+The high-level (essential) details of the episode are reflected in the shape defined by the sequence of coordinates of its events.
+This may be characterized by examining the low spatial frequency details of the trajectory's shape.
+Low-level details are reflected by (typically smaller-scale) within-event geometric patterns, which may be characterized by examining the high spatial frequency details of the trajectory's shape.
+A key finding of that work was that people who watched the episode recalled high-level details reliably, and in a similar way across people, whereas people varied substantially in how they remembered low-level details.
+This suggests that our memory systems may prioritize certain types of information (e.g., high-level details) about our experience relative to others (e.g., low-level details) as we encode our experiences into memories or when we recount our prior experiences.
+
 \subsection*{The \textit{matching problem}}
@@
 Each entry in these matrices denotes the number of time a given document (row) contained a given word (column).
-They then used the Singular Value Decomposition (SVD) to decompose the number-of-documents by number-of-words word count matrix into the product of a number-of-documents $\times K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values} matrix, and a $K \times $ number-of-unique-words \textit{word embeddings matrix}.
+They then used the Singular Value Decomposition (SVD) to decompose the number-of-documents by number-of-words word count matrix into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K~\times~K$ \textit{singular values} matrix, and a $K~\times$ number-of-unique-words \textit{word embeddings matrix}.
 The rows of the document embeddings matrix reflected the similarities between documents, and the columns of the word embeddings matrix reflected similarities between words.
@@
 \centering
-\includegraphics[width=0.75\textwidth]{figs/word_embedding_models}
+\includegraphics[width=0.8\textwidth]{figs/word_embedding_models}
 \caption{\textbf{Three approaches to text embedding.}  \textbf{A.~Matrix factorization models.} Per-document (or per-participant) word counts and/or frequencies are collected into a word count matrix (left).
-To compute $K$-dimensional embeddings of the documents and words, the word count matrix is factorized (typically by computing its Singular Value Decomposition) into the product of a number-of-documents $\times K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values} matrix, and a $K \times $ number-of-unique-words \textit{word embeddings matrix}.
+To compute $K$-dimensional embeddings of the documents and words, the word count matrix is factorized (typically by computing its Singular Value Decomposition) into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values} matrix, and a $K~\times $ number-of-unique-words \textit{word embeddings matrix}.
 \textbf{B.~Topic models.}  Fitting a topic model to a word count matrix (e.g., Panel A) reveals a set of latent \textit{topics}, or themes that pervade each document.
-The topics are specified in a $K \times$ number-of-words \textit{topics matrix}, and the per-document themes are specified in the number-of-documents $\times K$ \textit{topic proportions matrix}.
+The topics are specified in a $K~\times$ number-of-words \textit{topics matrix}, and the per-document themes are specified in the number-of-documents $\times~K$ \textit{topic proportions matrix}.
 A given document's topic proportions reflect the specific mix of themes that are reflected in that document (bar graph on the left).
@@
 A key advantage that topic models have over matrix factorization models is the ability to explain potentially ambiguous word usage.
-For example, consider the use of the word ``bat'' in the following passage:
+For example, consider the use of the word ``bat'' in the following illustrative passage:
 \begin{quote}
-  \textit{``...As the mutant vampire bat swooped low to attack, the explorer took a mighty swing of her trusty baseball bat.
-The swing connected, stunning the creature before it scrambled about on the ground and flew away.
-She didn't even bat an eye-- as the dark forces grew in power, such things had become increasingly common, nearly to the point of mundanity...''}
+  \textit{``As the slavering mutant vampire bat swooped low to attack, its eyes ablaze, the explorer grunted and took a mighty swing of her trusty aluminum baseball bat.
+The powerful swing connected with a echoing metallic \textit{clink}, stunning the creature before it scrambled about on the cave floor and flew away.
+She didn't even bat an eye-- as the Dark Wizard Chiroptera continued to draw strength from the mystic crystals, such things had become increasingly common of late, nearly to the point of mundanity...''}
   \end{quote}
@@
 
-% deep learning approaches:
-% - bag of words vs. sequence models vs. sentence-level embeddings
-% - accounting for use of context and grammar to convey meaning
-% - pre-training vs. customized training vs. training-free
+\subsection*{Extensions to spatial, semantic, affective, and social representations}
+The way we navigate through \textit{thought space} (i.e., word embedding spaces; Figs.~\ref{fig:trajectories}, \ref{fig:embedding-models}) differs from how we navigate in physical spatial environments.
+First, we cannot teleport in real space, whereas our thoughts can exhibit rapid jumps between different non-contiguous regions of word embedding space.
+Second, when two people visit the same fixed physical landmark, they necessarily visit the same physical location.
+When two people describe a shared experience, their idiosyncratic thoughts, goals, prior experiences, etc.
+can lead them to process and remember the experience differently.
+Third, when we revisit a fixed physical landmark, we (by definition) revisit the same spatial location.
+In other words, it is possible to visit the same spatial location twice.
+However, we cannot revisit our prior experiences in this way.
+We only get to experience each moment once; subsequent ``visits'' to a prior experience must occur through a different medium (e.g., our memory, another person’s memory, a physical recording, a written account, etc.).
+Fourth, we can revisit different aspects of the same prior experience on different ``visits''-- e.g., when we remember a specific event we can focus on the physical occurrence in one remembering, the emotional occurence in another remembering, the social implications in another remembering, etc.
+We can also revisit the same experience in different levels of detail or depth.
+There are no analogs of these phenomena in spatial navigation.
+Finally, \textit{physically} (at a macro scale), we are always located at a single moment and in a single location.
+The main argument of the quantum memory wave function framework I propose here is that we may be \textit{mentally} spread over many times or locations simultaneously.
+For example, this provides a way for us to integrate information from multiple prior experiences into a single decision or conceptualization, even if those experiences were not temporally contiguous.
 
-% Neural Probabilistilc Language Model (NPLM): BengEtal03 --> explicitly model word sequences
-% Scalable Hierarchical Distributed Language Model (SHDLM): MnihHint09 --> sequences of word trees
-
-%word2vec: MikoEtal13 --> semantic relations reflected in geometric operations
-%GloVe: PennEtal14 --> combines matrix factorization and local context window methods
-%SkipThought: KiroEtal15 --> reconstructs surrounding sentences of an encoded passage
-%BERT: DevlEtal18 --> pretraining language representations, context dependence of word embeddings
-%ELMO: PeteEtal18 --> how is it different from BERT/GPT?
-%InferSent: ConnEtal18 --> embeddings of sentences
-%Universal Sentence Encoder: CerEtal18 --> embeddings of sentences
-%Open-GPT: RadfEtal18 --> pretrained transformer (how is it different from BERT?)
-%RandSent: WietKiel19 --> training-free sentence-level embeddings
-%GPT-2: RadfEtal19
-%GPT-3: BrowEtal20
-
-\subsection*{Extensions to spatial, semantic, affective, and social representations}
+Many of these differences between physical and mental navigation also apply to other mental domains, including how we \textit{conceptualize} space (e.g., when we trace out a potential route in our mind).
 Just as we mentally spread ourselves across a distribution of times when we remember, we can perform similar mental operations when we conceptualize and retrieve other types of information.
@@
 A prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} concept or mental state.
-Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states.
+Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states, including non-contiguous mental states.
 Traditional approaches then ``clean up'' this distribution over possible mental states by taking its average (expected), maximum, or modal value.
-However, following the ``quantum wave function'' logic described in this manuscript, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.
+However, following the ``quantum wave function'' logic described in this manuscript, these multi-peaked receptive fields may provide a mechanism for associating non-contiguous representations.
 What are some implications of this view for non-memory domains?
@@
 In other words, although place cells encode an animal's location within an environment, even for a fixed cell and environment there is some variability with respect to where precisely within that environment the cell modulates its firing rate.
-These studies suggest that variability in the set of locations to which a place cell responds is not merely a reflection of noise or imprecision.
+Further, a number of studies have suggested that some neurons respond to distinct (non-contiguous) places~\citep{FentEtal08, RichEtal14b, LeeEtal19, DerdEtal09, GrieEtal20} and times~\citep{PastEtal08}.
+Taken together, these studies suggest that variability in the set of locations and/or times to which a neuron responds is not merely a reflection of noise or imprecision.
 Rather, this variability might have some functional relevance in terms of facilitating the learning of paths around obstacles and towards goals.

### 2020-09-23 paper/main.tex
working on R2 comments

@@
 Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.
-This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
+This body of work suggests that different brain regions reflect \textit{temporal receptive windows} that characterize the timescale(s) to which that region is maximally sensitive.
+This research complements work on the \textit{temporal window of integration} in the speech and perception literatures~\citep[e.g.,][]{ViemWake91, GautEtal12, vanWEtal07}.
+The temporal window of integration refers to the duration over which physically distinct stimuli are treated (by some brain system or network) as a single percept or an atomic unit.
+This systems-level work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
 Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
@@
 First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).
-Second, the studies suggest that members of this family of representations drifting at different rates become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.
+Second, the studies suggest that representations that reflect different timescales become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.
 These reactivations explain how our brains reactivate prior contexts when we remember the past.

### 2020-09-23 paper/main.tex
finished first pass of full revision addressing all R1 and R2 comments

@@
 Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
-These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02a, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
+In addition to this work suggesting that individual system or neurons maintain representations at their preferred timescales, a number of studies also report neurons and small populations exhibiting \textit{temporal multiplexing}, whereby information at different timescales is reflected in different features of an individual neuron's (or population's) activity patterns, such as spike timing versus firing rate and local field potential oscillations at different frequencies~\citep{CaruEtal18, KnigEich13, WatrEtal13, PanzEtal10}.
+
+The above ideas about parallel neural representations that reflect information at multiple timescales have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02a, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
@@
 In other words, it defines a function, $\Psi(x, t)$, such that the probability that a particle is located at position $x$ at time $t$ is equal to $|\Psi(x, t)|^2$.
-The central idea is that, although it may not be possible to know or predict the particle's precise position at a given moment, the wave function tells us how it's location's probabilty mass extends over space.
+The central idea is that, although it may not be possible to know or predict the particle's precise position at a given moment, the wave function tells us how it's location's probability mass extends over space.
 In turn, this may allow one to use the wave function to predict behaviors or constraints of the system(s) in which the particle is behaving and interacting.
@@
 \textbf{B.
-Mental time travel.} This constitutes a special case of the quantum memory wave function framework, whereby the individual's thoughts reflect a recency-weighted blend of prior thoughts, along with thoughts specific to one remembered prior experience (occuring at the timepoint indicated by the $\bigstar$ symbol).
+Mental time travel.} This constitutes a special case of the quantum memory wave function framework, whereby the individual's thoughts reflect a recency-weighted blend of prior thoughts, along with thoughts specific to one remembered prior experience (occurring at the timepoint indicated by the $\bigstar$ symbol).
 Each curve's coloring denotes the amount of temporal blurring.
@@
 
-The \textit{mental time travel} metaphor is defined by Tulving as ``the type of awareness experienced when one thinks back to a \textit{specific moment} in one's personal past and consciusly recollects some prior episode or state as it was previously experienced''~\citep[Fig.~\ref{fig:qwave}B][emphasis added]{WheeEtal97}.
+The \textit{mental time travel} metaphor is defined by Tulving as ``the type of awareness experienced when one thinks back to a \textit{specific moment} in one's personal past and consciously recollects some prior episode or state as it was previously experienced''~\citep[Fig.~\ref{fig:qwave}B][emphasis added]{WheeEtal97}.
 Whereas mental time travel refers to mentally revisiting to one specific moment in addition to now (potentially with some temporal blurring), the quantum memory wave framework implies that episodic memory entails spreading our awareness over \textit{many} moments (Fig.~\ref{fig:qwave}C).
@@
 Specifically, integrating over those experiences would seem to require some mechanism for associating and simultaneously revisiting them while also casting some part of our thoughts forward to the future.
+
+While the studies presented in the \textit{Introduction and overview} on temporal receptive windows, the temporal window of integration, time cells, and temporal multiplexing present strong empirical evidence that our brains maintain parallel representations at multiple timescales, questions remain about how or whether we experience those representations at a conscious level.
+For example, are we consciously aware of spreading our mental state over many moments from our past?
+Or is our subjective experience of remembering more like mentally time traveling back to a single ``hybrid'' moment that blends information from several of our (actual) prior experiences?
 
@@
 A rich recall density function analogous to the one displayed in Figure~\ref{fig:reinstatement}B could, in principle, be characteristic of memory for \textit{any} past event or experience.
-For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline, are functionally associated in memory.
+For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline~\citep{ArzyEtal09}, are functionally associated in memory.
 Nonetheless, a potential challenge to testing this hypothesis in the laboratory is that in many memory studies there is no particular reason for participants to functionally associate the stimuli.
@@
 
-Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{DeerEtal90, LandDuma97} attempted to compute these embeddings using word co-occurance statistics.
+Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{DeerEtal90, LandDuma97} attempted to compute these embeddings using word co-occurence statistics.
 The logic driving these approaches posited that documents that contained similar mixtures of words would be likely to be conceptually similar, and words that appeared in similar documents would also likely be conceptually similar.
@@
 \textit{Topic models}, such as \textit{Latent Dirichlet Allocation}~\citep[LDA;][]{BleiEtal03} represent an attempt to characterize the latent \textit{topics} or themes that documents comprise (Fig.~\ref{fig:embedding-models}B).
-Whereas matrix factorization approaches represent semantic meanings using low-dimensional approximations of word co-occurances or free association frequencies, topic models leverage these properties of documents (or behaviors) to explain \textit{why} words co-occur.
+Whereas matrix factorization approaches represent semantic meanings using low-dimensional approximations of word co-occurrences or free association frequencies, topic models leverage these properties of documents (or behaviors) to explain \textit{why} words co-occur.
 LDA defines a topic as a distribution of weights over words in the vocabulary.
@@
 At different ponts in the passage, ``bat'' refers to an animal, a piece of sports equipment, and a verb.
-Because each of these uses of the word appear in the same document, models driven solely by word co-occurance statistics will blend these different meanings such that the embedding of ``bat'' is somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
+Because each of these uses of the word appear in the same document, models driven solely by word co-occurrence statistics will blend these different meanings such that the embedding of ``bat'' is somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
 By contrast, treating each document as reflecting a blend of topics allows these uses to be disambiguated.
@@
 
-Many of these differences between physical and mental navigation also apply to other mental domains, including how we \textit{conceptualize} space (e.g., when we trace out a potential route in our mind).
+Many of these differences between physical and mental navigation also apply to other mental domains, including how we \textit{conceptualize} space~\citep[e.g., when we trace out a potential route in our mind;][]{GautvanW16, GautEtal18, ArzyScha19}.
 Just as we mentally spread ourselves across a distribution of times when we remember, we can perform similar mental operations when we conceptualize and retrieve other types of information.

### 2020-09-26 paper/main.tex
inputted hand-written notes, updated word embeddings figure colors

@@
 One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
-An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
+An intuition for why contextual representations might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
 Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.
-This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12a, AlyEtal18}.
+This notion, that our brains maintain parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12a, AlyEtal18}.
 Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.
-This body of work suggests that different brain regions reflect \textit{temporal receptive windows} that characterize the timescale(s) to which that region is maximally sensitive.
+This body of work suggests that different brain regions reflect \textit{temporal receptive windows} that characterize the timescale(s) to which each region is maximally sensitive.
 This research complements work on the \textit{temporal window of integration} in the speech and perception literatures~\citep[e.g.,][]{ViemWake91, GautEtal12, vanWEtal07}.
@@
 Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
-In addition to this work suggesting that individual system or neurons maintain representations at their preferred timescales, a number of studies also report neurons and small populations exhibiting \textit{temporal multiplexing}, whereby information at different timescales is reflected in different features of an individual neuron's (or population's) activity patterns, such as spike timing versus firing rate and local field potential oscillations at different frequencies~\citep{CaruEtal18, KnigEich13, WatrEtal13, PanzEtal10}.
+In addition to this work suggesting that individual systems or neurons maintain representations at their preferred timescales, a number of studies also report neurons and small populations exhibiting \textit{temporal multiplexing}, whereby information at different timescales is reflected in different features of an individual neuron's (or population's) activity patterns, such as spike timing versus firing rate and local field potential oscillations at different frequencies~\citep{CaruEtal18, KnigEich13, WatrEtal13, PanzEtal10}.
 
@@
 Other recent work indicates that different aspects of the reinstatement process itself might also occur at multiple timescales.
-For example, \cite{LindEtal19} found that high-level semantic information was reinstated prior to low-level (e.g.
-perceptual) details in an associative memory task.
+For example, \cite{LindEtal19} found that high-level semantic information was reinstated prior to low-level (e.g., perceptual) details in an associative memory task.
 Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.
@@
 In turn, this may allow one to use the wave function to predict behaviors or constraints of the system(s) in which the particle is behaving and interacting.
-Within the domain of episodic memory, for the purposes of describing the \textit{quantum memory wave function} metaphor, we can conceptualize the moments along our \textit{autobiographical timeline} as the spatial coordinates ($x$) in Schr\"{o}dinger's equation.
+Within the domain of episodic memory, for the purposes of describing the \textit{quantum memory wave function} metaphor, we can conceptualize the moments along our \textit{autobiographical timeline}~\citep{ArzyEtal09} as the spatial coordinates ($x$) in Schr\"{o}dinger's equation.
 Then our mental state at time $t$ may be conceptualized as spreading its probability mass over our autobiographical timeline.
-In other words, like Schr\"{o}dinger's equation, we could in principle define a function describing how our mental state at time $t$ reflects the blends of thoughts we experienced at other moments of our lives.
+In other words, like Schr\"{o}dinger's equation, many episodic memory models~\citep[e.g.,][]{HowaKaha02a, PolyEtal09, ShanHowa12} define a function describing how our mental state at time $t$ reflects the blends of thoughts we experienced at other moments of our lives.
 Under this metaphor, the effect of mental operations we perform with respect to our autobiographical timeline, such as remembering past experiences or predicting what will happen in the future, is to shift the probability mass of our quantum memory wave function (Fig.~\ref{fig:qwave}).
@@
 
-The \textit{mental time travel} metaphor is defined by Tulving as ``the type of awareness experienced when one thinks back to a \textit{specific moment} in one's personal past and consciously recollects some prior episode or state as it was previously experienced''~\citep[Fig.~\ref{fig:qwave}B][emphasis added]{WheeEtal97}.
+The \textit{mental time travel} metaphor is defined by Tulving as ``the type of awareness experienced when one thinks back to a \textit{specific moment} in one's personal past and consciously recollects some prior episode or state as it was previously experienced''~\citep[Fig.~\ref{fig:qwave}B;][emphasis added]{WheeEtal97}.
 Whereas mental time travel refers to mentally revisiting to one specific moment in addition to now (potentially with some temporal blurring), the quantum memory wave framework implies that episodic memory entails spreading our awareness over \textit{many} moments (Fig.~\ref{fig:qwave}C).
@@
 
-The quantum memory wave and mental time travel metaphors also appear to diverge in explaining how we might predict what will happen in the future.
-The quantum memory wave framing suggests that our predictions about the future draw from information pertaining to a blend of our prior experiences, weighted by their expected relevance to the future event in question (Fig.~\ref{fig:qwave}D.
-To the extent that the mental time travel metaphor assumes that we can mentally visit just one other time (or time range) in addition to now~\citep[e.g.,][]{SzpuTulv11}, it is not clear how this integration over temporally separated prior experiences might occur.
+The quantum memory wave and mental time travel metaphors also diverge in explaining how we might predict what will happen in the future.
+The quantum memory wave framing suggests that our predictions about the future draw from information pertaining to a blend of our prior experiences, weighted by their expected relevance to the future event in question (Fig.~\ref{fig:qwave}D).
+To the extent that the mental time travel metaphor assumes that we can mentally visit just one other time (or time range) in addition to now~\citep[e.g.,][]{SzpuTulv11}, it is not clear how this sort of integration over temporally separated prior experiences might occur.
 Specifically, integrating over those experiences would seem to require some mechanism for associating and simultaneously revisiting them while also casting some part of our thoughts forward to the future.
@@
 One insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \cite{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
-The authors used topic models~\citep{BleiEtal03} applied to human-generated annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.
+The authors used topic models~\citep{BleiEtal03} applied to human-generated annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors (see \textit{Defining the geometries of ongoing thoughts and memories}).
 Each shot's annotation contained the following information: a brief narrative description of what was happening, the location where the scene took place, whether that location was indoors or outdoors, the names of all on-screen characters, the names of the character(s) who were in focus in the camera shot, the name(s) of any character(s) who spoke during the scene, the camera angle of the shot, a transcription of any on-screen text, and whether or not there was any music present in the auditory background~\citep[full details may be found in][]{ChenEtal17}.
-The resulting \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode.
+The resulting \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode (Fig.~\ref{fig:embedding-models}B).
 When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change.
@@
 Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were characterized in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
-The example event explored in the figure comprises a scene where Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another victim has been murdered.
+The example event explored in the figure comprises a scene where Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another victim has committed suicide.
+Later it is revealed that the suicides are actually murders.
 The language used by human annotators describing the moment-by-moment contents of the scene used a similar mix of themes only for moments in the episode that were temporally proximal to the event in question (gray line).
 However, the way participants described what happened when they \textit{recounted} the episode later revealed a much different pattern.
-Specifically, their language mirrored that used when they were recalling other murders, and other moments when Sherlock and John were bantering with or about each other (Fig.~\ref{fig:reinstatement}C).
+Specifically, their language mirrored that used when they were recalling other related murders, and other moments when Sherlock and John were bantering with or about each other (Fig.~\ref{fig:reinstatement}C).
 
@@
 Later in the episode, it becomes clear that the apparent suicides are actually murders.
-The annotations refer to early deaths as suicides, whereas participants often refer to those early deaths as ``murders'' even though they could not have known that they were murders early on in the episode.
+The annotations refer to early deaths (including that shown in Fig.~\ref{fig:reinstatement}B) as suicides, whereas participants often refer to those early deaths as ``murders'' even though they could not have known that they were murders early on in the episode.
 In this way, participants' experiences \textit{subsequent to} a target remembered event may distort its memory by incorporating information learned during other experiences.
@@
 A rich recall density function analogous to the one displayed in Figure~\ref{fig:reinstatement}B could, in principle, be characteristic of memory for \textit{any} past event or experience.
-For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline~\citep{ArzyEtal09}, are functionally associated in memory.
+For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline, are functionally associated in memory.
 Nonetheless, a potential challenge to testing this hypothesis in the laboratory is that in many memory studies there is no particular reason for participants to functionally associate the stimuli.
@@
 In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
-\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figure~\ref{fig:trajectories}, revealing another key property of how we remember.
+\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats} and~\ref{fig:reinstatement}, revealing another key property of how we remember.
 Although there are substantial individual differences in the specific words people use to describe their shared experience of watching the same television episode, and how much detail they recollected, the overall shapes of those recalled experience trajectories are highly similar across people.
@@
 The matching problem refers to discovering an alignment in the labels between experienced and remembered events.
-Traditional memory paradigms most commonly employed over the past century~\citep{Kaha12} ask participants to study lists of \textit{items} (often randomly selected words) and then engage in a memory task, such as verbally recalling any words the participant remembers studying, in any order.
-The matching problem in such experiments is trivial: if we treat each studied word on a list as an ``event,'' then we can assume that when the participant verbalizes that word during their memory test, they are remembering that word's event.
+Traditional memory paradigms most commonly studied over the past century~\citep{Kaha12} ask participants to study lists of \textit{items} (often randomly selected words) and then engage in a memory task, such as verbally recalling any words the participant remembers studying, in any order.
+Solving the matching problem in such experiments is trivial: if we treat each studied word on a list as an ``event,'' then we can assume that the participant is remembering a word's event when they verbalize that word during their memory test.
 In other words, matching entails finding the exact match between a word the participant is verbalizing and that same word on the list that they studied.
@@
 For real-world memory phenomena, and in many so-called \textit{naturalistic} memory paradigms, solving the matching problem is more complex.
-People often experience real-world events (and scenes from movies or stories) as rich sensory experiences that derive deep meaning from surrounding moments, other prior experiences, emotions, goals, high-order understandings of the current situation, and so on.
-Many of these aspects of our real-world experiences are not immediately verbalizable-- and even when we are able to describe some specific aspects of what we are experiencing, there are typically many possibly ways we could have done so (e.g., varying in level of detail, specific wording, objectivity versus subjectivity, contextual details, etc.).
-Further, because the way we experience each moment depends in part on our unique prior experiences, even ostensibly shared experiences can seem different to each individual.
-Similarly, the way we recount a given experience is often a reflection of our own idiosyncratic experience of the original event, colored and distorted in an idiosyncratic way by our memory systems.
-This distorted memory must then be verbalized or communicated, which is itself an idiosyncratic (and potentially lossy in the information theoretic sense) process.
-Solving the matching problem under these constraints requires building explicit models of the dynamic conceptual content of our experiences and memories, and then estimating when and how that experienced and remembered content aligns.
+People often experience real-world events (and scenes from movies or stories) as rich sensory experiences that derive deep meaning from surrounding moments, other prior experiences, emotions, goals, high-order understanding of the current situation, and so on.
+Many of these aspects of our real-world experiences are not immediately verbalizable-- and even when we are able to describe some specific aspects of what we are experiencing, there are typically many possible ways we could have done so (e.g., varying in level of detail, specific wording, objectivity versus subjectivity, contextual details, etc.).
+Further, because the way we experience each moment depends in part on our unique prior experiences, even ostensibly shared experiences can \textit{seem} different to each individual.
+The way we recount a given experience is often a reflection of our own idiosyncratic experience of the original event, and is colored and distorted in an idiosyncratic way by our memory systems.
+This distorted memory must then be verbalized or communicated, which is itself an idiosyncratic (and potentially lossy, in the information theoretic sense) process.
+Solving the matching problem under these constraints poses a substantial theoretical and methodological challenge.
+Specifically, matching requires building explicit models of the dynamic conceptual content of our experiences and memories, and then estimating when and how that experienced and remembered content aligns.
 The approach featured in Figure~\ref{fig:trajectories} uses a word embedding space that is defined by a topic model~\citep{BleiEtal03}.
-Each coordinate in this space reflects the conceptual content underlying a set of words, such that two different text passages that describe the same conceptual content (but using different words, levels of detail, etc.) are mapped on to nearby points in the word embedding space.
-When these models are sufficiently expressive and robust to model our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal18c}.
+Each coordinate in this space reflects the conceptual content underlying a set of words, such that two different text passages that describe the same conceptual content (but using different words, levels of detail, etc.) are mapped onto nearby points in the word embedding space.
+When these models are sufficiently expressive and robust to solve the matching problem by modeling our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal18c}.
 
@@
 The fundamental goal of these models is to take as input a sequence of words (text) and produce as output one or more feature vectors whose coordinates reflect their conceptual (semantic) content.
-For example, when two similar documents are passed through these models, the goal is to produce two feature vectors that are closer (e.g., in Euclidean distance) than the feature vectors for two dissimilar documents.
+For example, when two conceptually similar documents are passed through these models, the goal is to produce two feature vectors that are closer (e.g., in Euclidean distance) than the feature vectors for two dissimilar documents.
 
-Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{DeerEtal90, LandDuma97} attempted to compute these embeddings using word co-occurence statistics.
-The logic driving these approaches posited that documents that contained similar mixtures of words would be likely to be conceptually similar, and words that appeared in similar documents would also likely be conceptually similar.
-To quantify these intuitions, \cite{LandDuma97} first constructed \textit{word count matrices} (Fig.~\ref{fig:embedding-models}A) for several text corpora (each comprising many thousands of documents).
+Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{DeerEtal90, LandDuma97} derived text embeddings using word co-occurence statistics.
+The logic driving this approach posits that documents that contain similar mixtures of words are likely to be conceptually similar, and words that appear in similar documents are also likely to be conceptually similar.
+To quantify these intuitions, \cite{DeerEtal90} first constructed \textit{word count matrices} (Fig.~\ref{fig:embedding-models}A) for several text corpora (each comprising many thousands of documents).
 Each entry in these matrices denotes the number of time a given document (row) contained a given word (column).
@@
 In the experiment, each participant was asked to free-associate (i.e., say the first word that came to mind) in response to a series of cue words.
-The authors then constructed a number-of-cue-words by number-of-unique-response-words matrix whose entries reflected the proportion of times each cue word (row) produced a given response word (column) as an associate.
+The authors then constructed a number-of-cue-words by number-of-unique-response-words matrix whose entries reflected the proportion of times each cue word (row) produced a given response word (column) across participants.
 After using SVD to decompose this \textit{word frequency matrix}, they obtained low-dimensional embeddings of each cue word and response word.
-In one sense, WAS provides a closer reflection of how people think about the meanings of words than LSA, since nearby WAS vectors reflect either cue words that produced similar responses, or response words that were evoked by similar cues.
+In one sense, WAS might provide a closer reflection of how people think about the meanings of words than LSA, since nearby WAS vectors reflect either cue words that produced similar responses, or response words that were evoked by similar cues.
 In other words, WAS derives embeddings directly from people's semantic judgement behaviors.
@@
 As the target application or population diverges from the experimental population (e.g., over time, across cultures, demographics, etc.) the WAS embeddings may become less accurate or relevant.
-By contrast, LSA may be re-fit relatively quickly to a new corpus, enabling the resulting embeddings to stay current.
+By contrast, LSA may be re-fit relatively quickly to any new corpus, enabling the resulting embeddings to stay current.
 Further, because LSA may be applied to any corpus, word embeddings may be easily computed and/or compared across languages, times, subject areas, etc.
@@
 \textbf{B.~Topic models.}  Fitting a topic model to a word count matrix (e.g., Panel A) reveals a set of latent \textit{topics}, or themes that pervade each document.
-The topics are specified in a $K~\times$ number-of-words \textit{topics matrix}, and the per-document themes are specified in the number-of-documents $\times~K$ \textit{topic proportions matrix}.
+The topics are specified in a $K~\times$ number-of-words \textit{topics matrix}, and the per-document themes are specified in a number-of-documents $\times~K$ \textit{topic proportions matrix}.
 A given document's topic proportions reflect the specific mix of themes that are reflected in that document (bar graph on the left).
@@
 LDA defines a topic as a distribution of weights over words in the vocabulary.
-For example, if one takes the \textit{vocabulary} to be the top 10,000 most commonly used words in English (excluding stop words that are devoid of explicit semantic meaning such as ``and,'' ``or,'' ``the,'' ``but,'' etc.), then each topic would comprise a vector of 10,000 numbers (weights) between zero and one, inclusive, where the full set of weights summed to one.
-As shown in the cartoon on the left of the Panel (where the four top-weighted words from each of five topics are displayed), a fruit-related topic (shown in red) might weight words like ``apple,'' ``banana,'' ``pear,'' and ``orange'' heavily, whereas words like ``cat,'' ``dog,'' ``dolphin,'' and ``lobster'' might receive relatively small weights.
+For example, if one takes the \textit{vocabulary} to be the top 10,000 most commonly used words in English (excluding stop words that are devoid of explicit semantic meaning such as ``and,'' ``or,'' ``the,'' ``but,'' etc.), then each topic comprises a vector of 10,000 non-negative weights that sums to one.
+As shown in the cartoon on the left of the Panel (where the four top-weighted words from each of five hypothetical topics are displayed), a fruit-related topic (shown in red) might weight heavily on words like ``apple,'' ``banana,'' ``pear,'' and ``orange,'' whereas words like ``cat,'' ``dog,'' ``dolphin,'' and ``lobster'' might receive relatively small weights.
 Given a set of $K$ topic vectors, topic models compute the blend of topics most probably reflected in a given document-- i.e., its \textit{topic proportions}.
 These topic proportion estimates are made by examining how different topics weight each of the words in the document, and by assuming that documents tend to reflect few topics (rather than many topics).
-Given only a word count matrix as input (just as in LSA), topic models automatically discover a $K$ by number-of-words \textit{topics matrix}, which describes how each topic (row) weights each word in the vocabulary (column), and a number-of-documents by $K$ \textit{topic proportions matrix}, which describes each document's topic proportions.
+Given only a word count matrix as input (just as in LSA), topic models automatically discover a $K~\times~$ number-of-words \textit{topics matrix}, which describes how each topic (row) weights each word in the vocabulary (column), and a number-of-documents $~\times~K$ \textit{topic proportions matrix}, which describes each document's topic proportions.
 The rows of the topic proportions matrix may be treated as low-dimensional embeddings of the set of documents.
 
-A key advantage that topic models have over matrix factorization models is the ability to explain potentially ambiguous word usage.
+A key advantage of topic models over matrix factorization models is their ability to explain potentially ambiguous word usage.
 For example, consider the use of the word ``bat'' in the following illustrative passage:
@@
   \textit{``As the slavering mutant vampire bat swooped low to attack, its eyes ablaze, the explorer grunted and took a mighty swing of her trusty aluminum baseball bat.
-The powerful swing connected with a echoing metallic \textit{clink}, stunning the creature before it scrambled about on the cave floor and flew away.
+The powerful strike connected with a echoing metallic \textit{clink}, stunning the wretched creature before it scrambled about on the cave floor and flew away.
 She didn't even bat an eye-- as the Dark Wizard Chiroptera continued to draw strength from the mystic crystals, such things had become increasingly common of late, nearly to the point of mundanity...''}
@@
 At different ponts in the passage, ``bat'' refers to an animal, a piece of sports equipment, and a verb.
-Because each of these uses of the word appear in the same document, models driven solely by word co-occurrence statistics will blend these different meanings such that the embedding of ``bat'' is somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
+Because each of these uses of the word appear in the same (hypothetical) document, models driven solely by word co-occurrence statistics would blend these different meanings such that the embedding of ``bat'' was somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
 By contrast, treating each document as reflecting a blend of topics allows these uses to be disambiguated.
-If other documents that mention the word ``bat'' describe vampires, swooping, or other mammals, but \textit{not} sports equipment or eye-related verbs, then a topic model could explain this observation by inferring that multiple topics place substantial weight on the word ``bat.''  Another strength of topic models is their interpretability.
+If other documents that mention the word ``bat'' describe vampires, swooping, or other mammals, but \textit{not} sports equipment or eye-related verbs, then a topic model could explain this pattern by inferring that multiple topics place substantial weight on the word ``bat.''  Another strength of topic models is their interpretability.
 Because the models directly output each document's topic proportions along with each topic's weights on all words in the vocabulary, the general themes that each topic reflects are often intuitive and the per-document topic proportions may be directly related to those themes.
@@
 
-The Neural Probabilistic Language Model~\citep[NPLM;][]{BengEtal03} was one of the earliest deep learning-based sequence models.
+The \textit{Neural Probabilistic Language Model}~\citep[NPLM;][]{BengEtal03} was one of the earliest deep learning-based sequence models.
 NLPM derives word embeddings for each word in the vocabulary (such that nearby embeddings reflect semantically related words), and then models word sequences according to their meanings.
@@
 Effectively, NLPM attempts to learn arbitrarily long sequences.
-In a related approach, \cite{MnihHint09} proposed the Scalable Hierarchical Distributed Language Model, which learns sequences of $n$-grams to achieve greater explanatory power.
+In a related approach, \cite{MnihHint09} proposed the \textit{Scalable Hierarchical Distributed Language Model}, which learns sequences of $n$-grams to achieve greater explanatory power.
 
@@
 For example, if each word's semantic meaning is captured by a feature vector, then in principle one could perform vector algebra on different words' embeddings to produce new concepts that combine their meanings.
-For example, subtracting the vector for ``man'' from that of ``king'' and then adding the vector for ``woman'' would result in a new feature vector similar to that of ``queen''~\citep{MikoEtal13b}.
+For example, subtracting the vector for ``man'' from that of ``king'' and then adding the vector for ``woman'' might result in a new feature vector similar to that of ``queen''~\citep{MikoEtal13b}.
 \cite{MikoEtal13a} developed the deep learning model \textit{word2vec} to explicitly optimize the accuracy of these vector operations.
-Whereas word2vec is a bag of words model, Global Vectors for Word Representation~\citep[GloVe;][]{PennEtal14} extends this approach to model each word's local (temporal) context~\citep[also see][]{PeteEtal18}.
+Whereas word2vec is a bag of words model, \textit{Global Vectors for Word Representation}~\citep[GloVe;][]{PennEtal14} extends this approach to model each word's local (temporal) context~\citep[also see][]{PeteEtal18}.
 
 Several recent models have begun to move beyond single word-level embeddings.
-For example, SkipThought~\citep{KiroEtal15}, Universal Sentence Encoder~\citep{CerEtal18}, InferSent~\citep{ConnEtal18}, and RandSent~\citep{WietKiel19} each compute sentence-level embeddings.
+For example, \textit{SkipThought}~\citep{KiroEtal15}, \textit{Universal Sentence Encoder}~\citep{CerEtal18}, \textit{InferSent}~\citep{ConnEtal18}, and \textit{RandSent}~\citep{WietKiel19} each compute sentence-level embeddings.
 The central insight of these approaches is that capturing the nuanced meaning of a particular concept often requires \textit{several} words that have been organized into a complete sentence or sentence-like grammatical construct.
@@
 One of the first of this new class of model, Bidirectional Encoder Representations from Transformers~\citep[BERT;][]{DevlEtal18}, uses a pre-trained deep neural network, whose weights in a single output layer may be fine-tuned to suit a wide range of new tasks.
-The latest additions to this class of models are called Generative Pre-Trained (GPT) transformers~\citep{RadfEtal18, RadfEtal19, BrowEtal20}.
+The latest additions to this class of models are called \textit{Generative Pre-trained Transformers}~\citep[GPT;][]{RadfEtal18, RadfEtal19, BrowEtal20}.
 
@@
 Second, when two people visit the same fixed physical landmark, they necessarily visit the same physical location.
-When two people describe a shared experience, their idiosyncratic thoughts, goals, prior experiences, etc.
-can lead them to process and remember the experience differently.
+When two people describe a shared experience, their idiosyncratic thoughts, goals, prior experiences, etc.\ can lead them to process and remember the experience differently.
 Third, when we revisit a fixed physical landmark, we (by definition) revisit the same spatial location.
@@
 However, we cannot revisit our prior experiences in this way.
-We only get to experience each moment once; subsequent ``visits'' to a prior experience must occur through a different medium (e.g., our memory, another person’s memory, a physical recording, a written account, etc.).
+Rather, we only get to experience each moment once; subsequent ``visits'' to a prior experience must occur through a different medium (e.g., our memory, another person’s memory, a physical recording, a written account, etc.).
 Fourth, we can revisit different aspects of the same prior experience on different ``visits''-- e.g., when we remember a specific event we can focus on the physical occurrence in one remembering, the emotional occurence in another remembering, the social implications in another remembering, etc.
 We can also revisit the same experience in different levels of detail or depth.
-There are no analogs of these phenomena in spatial navigation.
+There are no deep analogs of these phenomena in spatial navigation.
 Finally, \textit{physically} (at a macro scale), we are always located at a single moment and in a single location.
@@
 In other words, although place cells encode an animal's location within an environment, even for a fixed cell and environment there is some variability with respect to where precisely within that environment the cell modulates its firing rate.
-Further, a number of studies have suggested that some neurons respond to distinct (non-contiguous) places~\citep{FentEtal08, RichEtal14b, LeeEtal19, DerdEtal09, GrieEtal20} and times~\citep{PastEtal08}.
+Further, a number of studies have suggested that some neurons respond to distinct non-contiguous places~\citep{FentEtal08, RichEtal14b, LeeEtal19, DerdEtal09, GrieEtal20} and times~\citep{PastEtal08}.
 Taken together, these studies suggest that variability in the set of locations and/or times to which a neuron responds is not merely a reflection of noise or imprecision.
@@
 \section*{Concluding remarks}
-When we remember an experience, we bring back thoughts related to a range of other semantically, conceptually, and functionally related experiences.
+When we remember an experience, we bring back thoughts related to a range of other conceptually and functionally related experiences.
 This can serve an important functional role: it allows us to integrate information across temporally separated experiences.

### 2020-09-29 paper/main.tex
another pass through the paper

@@
 \usepackage{geometry}
-\usepackage{natbib}
+\usepackage[sort]{natbib}
 \usepackage{pxfonts}
@@
 This view has fundamental implications for how we study and build theories about memory: it suggests that we can match up specific moments of recollection with specific moments from the past.
-These implications extend to how we design episodic memory experiments and how we analyze the behavioral and neural data from those experiments in order to understand the neural and/or cognitive mechanisms that underly memory retrieval.
+These implications extend from how we design episodic memory experiments to how we analyze and interpret the behavioral and neural data from those experiments in order to understand the neural and/or cognitive basis of memory retrieval.
 At its core, characterizing episodic memory as mental time travel makes studying and understanding memory about matching.
@@
 
-One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
+One practical way to distinguish content from context representations is to separate out mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
 An intuition for why contextual representations might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
 Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.
-This notion, that our brains maintain parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12a, AlyEtal18}.
+This notion, that our brains maintain parallel mental representations drifting at different timescales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12a, AlyEtal18}.
 Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.
@@
 This research complements work on the \textit{temporal window of integration} in the speech and perception literatures~\citep[e.g.,][]{ViemWake91, GautEtal12, vanWEtal07}.
-The temporal window of integration refers to the duration over which physically distinct stimuli are treated (by some brain system or network) as a single percept or an atomic unit.
+The temporal window of integration refers to the duration over which physically distinct stimuli are treated (by some brain system or network) as a single percept or atomic unit.
 This systems-level work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
 Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
-In addition to this work suggesting that individual systems or neurons maintain representations at their preferred timescales, a number of studies also report neurons and small populations exhibiting \textit{temporal multiplexing}, whereby information at different timescales is reflected in different features of an individual neuron's (or population's) activity patterns, such as spike timing versus firing rate and local field potential oscillations at different frequencies~\citep{CaruEtal18, KnigEich13, WatrEtal13, PanzEtal10}.
+In addition to this work suggesting that individual systems or neurons maintain representations at their preferred timescales, a number of studies also report neurons and small populations exhibiting \textit{temporal multiplexing}, whereby information at different timescales is reflected in different features of an individual neuron's (or population's) activity patterns, such as spike timing versus firing rate or local field potential oscillations at different frequencies~\citep{CaruEtal18, KnigEich13, WatrEtal13, PanzEtal10}.
 
@@
 First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).
-Second, the studies suggest that representations that reflect different timescales become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.
+Second, the studies suggest that representations that reflect different timescales become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of timescales.
 These reactivations explain how our brains reactivate prior contexts when we remember the past.
@@
 This is not simply a matter of imprecision in mental time travel (i.e., a reflection of the low effective resolution with which we can revisit specific moments from our past).
-Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences, analogous to a quantum wave function spreading its probability mass over space.
+Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences, analogous to a quantum wave function spreading its probability mass over space (see \textit{Defining the {\normalfont quantum memory wave function} metaphor and comparing it to the {\normalfont mental time travel} metaphor}).
 The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each drifting at a different timescale.
@@
 For example, having an important conversation with a friend might require bringing in information acquired during other interactions with that friend, other conversations about related or similarly important topics, general knowledge about communication styles, etc.
-\citep[This notion of bringing a range of prior experiences to bear on guiding our ongoing behaviors is also related to a literature on \textit{situtation models}; for review see][]{RangRitc12}.
+This notion of bringing a range of prior experiences to bear on guiding our ongoing behaviors is also related to a literature on \textit{situtation models}~\citep[for review see][]{RangRitc12}.
 That we can integrate information across experiences suggests that the integration of incoming information at different rates need not be a purely mechanical process, but rather may reflect a deeper functional role~\citep[also see][]{BrigEtal18}.
@@
 
-\section*{Defining the ``quantum memory wave function'' metaphor and comparing it to the ``mental time travel'' metaphor}
+\section*{Defining the \textit{quantum memory wave function} metaphor and comparing it to the \textit{mental time travel} metaphor}
 Unlike in classical physics, where matter dutifully obeys Newton's laws, quantum physics postulates a different (though potentially complementary) set of rules governing how matter behaves and interacts.
-Perhaps the most famous equation in quantum physics is \cite{Schr26}'s equation describing the quantum \textit{wave function}~\citep[this finding was also inspired by][]{deBr24}.
+Perhaps the most famous equation in quantum physics is \cite{Schr26}'s equation describing the quantum \textit{wave function}~\citep[this work was also inspired by][]{deBr24}.
 Schr\"{o}dinger's equation describes, in probabilistic terms, the location of a particle in space, at a given time.
 In other words, it defines a function, $\Psi(x, t)$, such that the probability that a particle is located at position $x$ at time $t$ is equal to $|\Psi(x, t)|^2$.
-The central idea is that, although it may not be possible to know or predict the particle's precise position at a given moment, the wave function tells us how it's location's probability mass extends over space.
-In turn, this may allow one to use the wave function to predict behaviors or constraints of the system(s) in which the particle is behaving and interacting.
+The central idea is that, although it may not be possible to know or predict the particle's precise position at a given moment, the wave function tells us how its location's probability mass extends over space.
+In turn, this allows us to use the wave function to predict behaviors or constraints of the system(s) in which the particle is behaving and interacting.
 Within the domain of episodic memory, for the purposes of describing the \textit{quantum memory wave function} metaphor, we can conceptualize the moments along our \textit{autobiographical timeline}~\citep{ArzyEtal09} as the spatial coordinates ($x$) in Schr\"{o}dinger's equation.
@@
   \caption{\textbf{Memory wave functions.
-All panels.} Each panel displays a quantum memory wave function for a hypothetical individual at a particular moment (time $t=0$ indicated by $\blacktriangle$)).
-The $x$-axes correspond to the position ($t$) along an autobiographical timeline.
-The $y$-axes display how much the individual's thoughts at time $t = 0$ include information about the individual's past (or future) perceived or predicted experiences at other times.
+All panels.} Each panel displays a quantum memory wave function for a hypothetical individual at a particular moment (time $t=0$ indicated by $\blacktriangle$).
+The $x$-axes correspond to the position ($t$) along the individual's autobiographical timeline.
+The $y$-axes display how much the individual's thoughts at time $t = 0$ include information about the individual's past (or future) perceived or predicted experiences (i.e., at other times).
+Different colors denote different amounts of temporal blurring.
 All curves have been normalized to have a maximum value of 1 and a minimum value of 0.
@@
 Mental time travel.} This constitutes a special case of the quantum memory wave function framework, whereby the individual's thoughts reflect a recency-weighted blend of prior thoughts, along with thoughts specific to one remembered prior experience (occurring at the timepoint indicated by the $\bigstar$ symbol).
-Each curve's coloring denotes the amount of temporal blurring.
-In this example, greater amounts of blurring denotes that the individual is bringing to mind information that is farther from the remembered moment or from $t = 0$.
 \textbf{C.
@@
 In addition, they bring to mind information pertaining to other conceptually related (but temporally non-contiguous) moments, indicated by $\star$ symbols.
-As in Panel B, different curves are colored according to their degree of temporal blurring.
 \textbf{D.
@@
 
-While the studies presented in the \textit{Introduction and overview} on temporal receptive windows, the temporal window of integration, time cells, and temporal multiplexing present strong empirical evidence that our brains maintain parallel representations at multiple timescales, questions remain about how or whether we experience those representations at a conscious level.
-For example, are we consciously aware of spreading our mental state over many moments from our past?
+While the studies presented in the \textit{Introduction and overview} (on temporal receptive windows, the temporal window of integration, time cells, and temporal multiplexing) present strong empirical evidence that our brains maintain parallel representations at multiple timescales, questions remain about how or whether we experience those representations at a conscious level.
+For example, are we consciously aware of spreading our mental states over many moments from our past?
 Or is our subjective experience of remembering more like mentally time traveling back to a single ``hybrid'' moment that blends information from several of our (actual) prior experiences?
@@
 If we examine the temporal autocorrelation matrices of participants' recalls of the television episode (characterized using the same topic model used to capture the content of the episode; Figs.~\ref{fig:corrmats}B, C), we see a different structure than that of the episode itself.
-Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off-diagonal correlations between the thematic content of temporally distant events are substantially stronger in participants' recalls than in the original episode.
-Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode and participants' recalls (Fig.~\ref{fig:reinstatement}A).
+Although individual participants' recalls also exhibit periods of content stability punctuated by rapid change (visible in the block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off-diagonal correlations between the thematic content of temporally distant events are substantially stronger in participants' recalls than in the original episode.
+Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode versus participants' recalls (Fig.~\ref{fig:reinstatement}A).
 Whereas the autocorrelations in the original episode's content rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around an average of 0.2.
@@
 To understand which specific events are described (during recall) as more similar than what their original experiences (as characterized by their topic vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.
-Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were characterized in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
-The example event explored in the figure comprises a scene where Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another victim has committed suicide.
-Later it is revealed that the suicides are actually murders.
-The language used by human annotators describing the moment-by-moment contents of the scene used a similar mix of themes only for moments in the episode that were temporally proximal to the event in question (gray line).
+Figure~\ref{fig:reinstatement}B displays, for a representative example recalled event, which other parts of the episode were characterized in a similar way by the participants when they recounted the episode later (black and colored lines), and which other parts of the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
+In the example event explored in the figure, Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another person has committed suicide.
+Later the viewer learns that the suicides were actually murders.
+The language used by human annotators describing the moment-by-moment contents of the event used a similar mix of themes only for temporally nearby parts of the episode (gray line).
 However, the way participants described what happened when they \textit{recounted} the episode later revealed a much different pattern.
@@
 
-One concrete example of this phenomenon is that, early in the episode, Sherlock and John appear to be investigating a series of suicides.
+One concrete example of this phenomenon is that early in the episode, including during the target event, Sherlock and John appear to be investigating a series of suicides.
 Later in the episode, it becomes clear that the apparent suicides are actually murders.
@@
 \textbf{B.
-Recall density functions for one recalled moment.} Each curve reflects the correlations between the topic vector at each (relative) moment of the episode (or participants' recalls of the episode), and the topic vector of an event beginning approximately $\frac{1}{3}$ through the episode (relative time: 0.328).
+Recall density functions for one recalled event.} Each curve reflects the correlations between the topic vector at each (relative) moment of the episode (or participants' recalls of the episode), and the topic vector of an event beginning approximately $\frac{1}{3}$ through the episode (relative time: 0.328).
 The gray curve denotes these correlations for the topic vectors derived from annotations of the episode itself.
@@
 
-A rich recall density function analogous to the one displayed in Figure~\ref{fig:reinstatement}B could, in principle, be characteristic of memory for \textit{any} past event or experience.
+A recall density function analogous to the one displayed in Figure~\ref{fig:reinstatement}B could, in principle, be characteristic of memory for \textit{any} past event or experience.
 For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline, are functionally associated in memory.
@@
 Other work suggests that our memory systems may also leverage these regularities to parse our ongoing experiences into discrete events~\citep{SchaEtal13, Shap19}.
-Are these examples in the literature on random list learning studies, indicating that our memory systems leverage ``accidental'' structure in random sequences, a reflection of the same fundamental principle driving the sorts of memory reactivation function displayed in Figure~\ref{fig:reinstatement}B?
+Are these examples in the literature on random list learning studies, indicating that our memory systems leverage ``accidental'' structure in random sequences, a reflection of the same fundamental principle driving the memory reactivation function displayed in Figure~\ref{fig:reinstatement}B?
 
 It is difficult to explicitly test whether (apparent) semantic reinstatement in word list learning, as reflected in people's tendencies to semantically cluster recalls of random word lists, reflects integration of information across the experiences of studying those words.
-However, remembering word lists does not \textit{require} integrating information across words in order to specifically gain a deep understanding of the list.
+However, remembering word lists does not specifically \textit{require} integrating information across words in order to gain a deep understanding of the list.
 By design, there \textit{is} no deeper meaning of random word lists.
@@
 The trajectory through topic space taken by participants' recalls of the television episode.} The average two-dimensional trajectory captured by 17 participants' recalls of the episode, with the same format and coloring as the trajectory in Panel A.
-The arrows reflect the directions of the average transitions taken by participants whose trajectories crossed that part of topic space; blue denotes reliable agreement across participants.
+The arrows reflect the directions of the average transitions taken by participants whose trajectories crossed that part of topic space; blue arrows denote directions of reliable agreement across participants.
 Figure adapted from \cite{HeusEtal18c}, where additional details may be found.}
@@
 Although there are substantial individual differences in the specific words people use to describe their shared experience of watching the same television episode, and how much detail they recollected, the overall shapes of those recalled experience trajectories are highly similar across people.
-This has potentially exciting implications for what it means to communicate memories to other people: we need to convey the ``essence'' (shape) of our experience.
+This has potentially exciting implications for what it means to communicate memories to other people: to effectively communicate, we often need to convey the ``essence'' (gross overall shape) of our experience.
 It also has implications for how we learn by analogy: perhaps events that share schemas~\citep[e.g., ][]{BaldEtal18}  unfold in similar ways (e.g., have similar shapes).
@@
 The high-level (essential) details of the episode are reflected in the shape defined by the sequence of coordinates of its events.
-This may be characterized by examining the low spatial frequency details of the trajectory's shape.
-Low-level details are reflected by (typically smaller-scale) within-event geometric patterns, which may be characterized by examining the high spatial frequency details of the trajectory's shape.
+These are captured by low spatial frequency properties of the trajectory's shape.
+Low-level details are characterized by (typically smaller-scale) within-event geometric patterns, which are captured by the high spatial frequency properties of the trajectory's shape.
 A key finding of that work was that people who watched the episode recalled high-level details reliably, and in a similar way across people, whereas people varied substantially in how they remembered low-level details.
-This suggests that our memory systems may prioritize certain types of information (e.g., high-level details) about our experience relative to others (e.g., low-level details) as we encode our experiences into memories or when we recount our prior experiences.
+This suggests that our memory systems may prioritize certain types of information (e.g., high-level details) about our experience over others (e.g., low-level details) as we encode our experiences into memories or when we recount our prior experiences.
 
@@
 
-For real-world memory phenomena, and in many so-called \textit{naturalistic} memory paradigms, solving the matching problem is more complex.
+For real-world memory phenomena, and in many naturalistic memory paradigms, solving the matching problem is more complex.
 People often experience real-world events (and scenes from movies or stories) as rich sensory experiences that derive deep meaning from surrounding moments, other prior experiences, emotions, goals, high-order understanding of the current situation, and so on.
@@
 \subsection*{Defining the geometries of ongoing thoughts and memories}
-Defining geometric spaces that can appropriately capture the conceptual contents of our thoughts and experiences has become a major focus of computational science.
+Defining geometric spaces that can reliably capture the conceptual contents of our thoughts and experiences has become a major focus of computational science.
 One family of models that has shown particular promise over the past several decades is broadly referred to as \textit{word embedding models} (Fig.~\ref{fig:embedding-models}).
@@
 Each entry in these matrices denotes the number of time a given document (row) contained a given word (column).
-They then used the Singular Value Decomposition (SVD) to decompose the number-of-documents by number-of-words word count matrix into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K~\times~K$ \textit{singular values} matrix, and a $K~\times$ number-of-unique-words \textit{word embeddings matrix}.
-The rows of the document embeddings matrix reflected the similarities between documents, and the columns of the word embeddings matrix reflected similarities between words.
+They then used the Singular Value Decomposition (SVD) to decompose the number-of-documents $~\times~$ number-of-words word count matrix into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K~\times~K$ \textit{singular values} matrix, and a $K~\times$ number-of-unique-words \textit{word embeddings matrix}.
+The rows of the document embeddings matrix reflect the similarities between documents, and the columns of the word embeddings matrix reflect similarities between words.
 
@@
 In the experiment, each participant was asked to free-associate (i.e., say the first word that came to mind) in response to a series of cue words.
-The authors then constructed a number-of-cue-words by number-of-unique-response-words matrix whose entries reflected the proportion of times each cue word (row) produced a given response word (column) across participants.
+The authors then constructed a number-of-cue-words $~\times~$ number-of-unique-response-words matrix whose entries reflected the proportion of times each cue word (row) produced a given response word (column) across participants.
 After using SVD to decompose this \textit{word frequency matrix}, they obtained low-dimensional embeddings of each cue word and response word.
-In one sense, WAS might provide a closer reflection of how people think about the meanings of words than LSA, since nearby WAS vectors reflect either cue words that produced similar responses, or response words that were evoked by similar cues.
+In one sense, WAS might provide a more accurate and behaviorally relevant reflection of how people think about the meanings of words than LSA, since nearby WAS vectors reflect either cue words that produced similar responses, or response words that were evoked by similar cues.
 In other words, WAS derives embeddings directly from people's semantic judgement behaviors.
@@
 \includegraphics[width=0.8\textwidth]{figs/word_embedding_models}
-\caption{\textbf{Three approaches to text embedding.}  \textbf{A.~Matrix factorization models.} Per-document (or per-participant) word counts and/or frequencies are collected into a word count matrix (left).
+\caption{\textbf{Three approaches to text embedding.}  \textbf{A.~Matrix factorization models.} Per-document (or per-stimulus) word counts and/or frequencies are collected into a word count matrix (left).
 To compute $K$-dimensional embeddings of the documents and words, the word count matrix is factorized (typically by computing its Singular Value Decomposition) into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values} matrix, and a $K~\times $ number-of-unique-words \textit{word embeddings matrix}.
@@
 For example, if one takes the \textit{vocabulary} to be the top 10,000 most commonly used words in English (excluding stop words that are devoid of explicit semantic meaning such as ``and,'' ``or,'' ``the,'' ``but,'' etc.), then each topic comprises a vector of 10,000 non-negative weights that sums to one.
-As shown in the cartoon on the left of the Panel (where the four top-weighted words from each of five hypothetical topics are displayed), a fruit-related topic (shown in red) might weight heavily on words like ``apple,'' ``banana,'' ``pear,'' and ``orange,'' whereas words like ``cat,'' ``dog,'' ``dolphin,'' and ``lobster'' might receive relatively small weights.
+As shown in the cartoon on the left of the Panel (where the four top-weighted words from each of five hypothetical topics are displayed), a fruit-related topic (shown in purple) might weight heavily on words like ``apple,'' ``banana,'' ``pear,'' and ``orange,'' whereas words like ``cat,'' ``dog,'' ``dolphin,'' and ``lobster'' might receive relatively small weights.
 Given a set of $K$ topic vectors, topic models compute the blend of topics most probably reflected in a given document-- i.e., its \textit{topic proportions}.
@@
 \begin{quote}
-  \textit{``As the slavering mutant vampire bat swooped low to attack, its eyes ablaze, the explorer grunted and took a mighty swing of her trusty aluminum baseball bat.
+  \textit{``As the slavering mutant vampire \textbf{bat} swooped low to attack, its eyes ablaze, the explorer grunted and took a mighty swing of her trusty aluminum baseball \textbf{bat}.
 The powerful strike connected with a echoing metallic \textit{clink}, stunning the wretched creature before it scrambled about on the cave floor and flew away.
-She didn't even bat an eye-- as the Dark Wizard Chiroptera continued to draw strength from the mystic crystals, such things had become increasingly common of late, nearly to the point of mundanity...''}
+She didn't even \textbf{bat} an eye-- as the Dark Wizard Chiroptera continued to draw strength from the mystic crystals, such things had become increasingly common of late, nearly to the point of mundanity...''}
   \end{quote}
-At different ponts in the passage, ``bat'' refers to an animal, a piece of sports equipment, and a verb.
+At different ponts in the passage, ``bat'' refers to an animal, a piece of sports equipment, and an eye-related verb.
 Because each of these uses of the word appear in the same (hypothetical) document, models driven solely by word co-occurrence statistics would blend these different meanings such that the embedding of ``bat'' was somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
@@
 Effectively, NLPM attempts to learn arbitrarily long sequences.
-In a related approach, \cite{MnihHint09} proposed the \textit{Scalable Hierarchical Distributed Language Model}, which learns sequences of $n$-grams to achieve greater explanatory power.
+In a related approach, \cite{MnihHint09} proposed the \textit{Scalable Hierarchical Distributed Language Model}, which learns sequences of variable length $n$-grams to achieve greater explanatory power.
 
-The next major advance in word embeddings came from attempts to more precisely capture the geometric relations between words.
+The next major advances in word embeddings came from attempts to more precisely capture the geometric relations between words.
 For example, if each word's semantic meaning is captured by a feature vector, then in principle one could perform vector algebra on different words' embeddings to produce new concepts that combine their meanings.
@@
 Because these models are very computationally expensive to fit, an important discovery has been that many layers in these models may be pre-trained without substantially affecting performance, even on new tasks.
-One of the first of this new class of model, Bidirectional Encoder Representations from Transformers~\citep[BERT;][]{DevlEtal18}, uses a pre-trained deep neural network, whose weights in a single output layer may be fine-tuned to suit a wide range of new tasks.
+One of the first of this new class of model, \textit{Bidirectional Encoder Representations from Transformers}~\citep[BERT;][]{DevlEtal18}, uses a pre-trained deep neural network, whose weights in a single output layer may be fine-tuned to suit a wide range of new tasks.
 The latest additions to this class of models are called \textit{Generative Pre-trained Transformers}~\citep[GPT;][]{RadfEtal18, RadfEtal19, BrowEtal20}.
@@
 Second, when two people visit the same fixed physical landmark, they necessarily visit the same physical location.
-When two people describe a shared experience, their idiosyncratic thoughts, goals, prior experiences, etc.\ can lead them to process and remember the experience differently.
-Third, when we revisit a fixed physical landmark, we (by definition) revisit the same spatial location.
-In other words, it is possible to visit the same spatial location twice.
+When two people describe a shared experience, their idiosyncratic thoughts, goals, prior experiences, etc.\ can lead them to process and remember the experience differently, thereby visiting different locations within thought space.
+Third, each time we revisit a fixed physical landmark, we (by definition) revisit the same spatial location.
+In other words, it is possible to visit the same spatial location multiple times.
 However, we cannot revisit our prior experiences in this way.
-Rather, we only get to experience each moment once; subsequent ``visits'' to a prior experience must occur through a different medium (e.g., our memory, another person’s memory, a physical recording, a written account, etc.).
+Rather, we only get to experience each moment once; subsequent ``visits'' to a prior experience must occur through a different medium than the original experience (e.g., our memory, another person’s memory, a physical recording, a written account, etc.).
 Fourth, we can revisit different aspects of the same prior experience on different ``visits''-- e.g., when we remember a specific event we can focus on the physical occurrence in one remembering, the emotional occurence in another remembering, the social implications in another remembering, etc.
@@
 A prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} concept or mental state.
-Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states, including non-contiguous mental states.
+Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading the probability masses of their predictions over multiple possible mental states, including non-contiguous mental states.
 Traditional approaches then ``clean up'' this distribution over possible mental states by taking its average (expected), maximum, or modal value.
@@
 Rather, the set of stimuli likely to be encountered live within subspaces or on manifolds in the relevant representational space.
-The geometry of these manifolds provides insights into the statistical structure of the stimuli.
-It also provides potential insights into how the brain represents those stimuli, for example to the extent that the geometries of these manifolds may be leveraged to improve brain decoding.
+The geometries of these manifolds provides insights into the statistical structure of the stimuli.
+They also provide potential insights into how the brain represents those stimuli, for example to the extent that the geometries of these manifolds may be leveraged to improve brain decoding.
 The shapes of these manifolds can potentially provide insights into how our thoughts are distributed over representational spaces.

### 2021-01-11 paper/main.tex
first attempt at addressing R1 concerns

@@
 Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences, analogous to a quantum wave function spreading its probability mass over space (see \textit{Defining the {\normalfont quantum memory wave function} metaphor and comparing it to the {\normalfont mental time travel} metaphor}).
-The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each drifting at a different timescale.
+The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each reflecting information at a different timescale.
 The notion of thinking of any one particular moment from the past is incompatible with this multiple timescales view, in that our thoughts and brain activity patterns reflect information from a range of moments, even when we are not specifically engaged in remembering.
@@
 The \textit{mental time travel} metaphor is defined by Tulving as ``the type of awareness experienced when one thinks back to a \textit{specific moment} in one's personal past and consciously recollects some prior episode or state as it was previously experienced''~\citep[Fig.~\ref{fig:qwave}B;][emphasis added]{WheeEtal97}.
-Whereas mental time travel refers to mentally revisiting to one specific moment in addition to now (potentially with some temporal blurring), the quantum memory wave framework implies that episodic memory entails spreading our awareness over \textit{many} moments (Fig.~\ref{fig:qwave}C).
+Whereas mental time travel refers to mentally revisiting one specific moment in addition to now (potentially with some temporal blurring), the quantum memory wave framework implies that episodic memory entails spreading our awareness over \textit{many} moments (Fig.~\ref{fig:qwave}C).
 In this way, the key distinction between these framings is whether remembering entails simultaneously revisiting one single moment versus many moments.
@@
 If we examine the temporal autocorrelation matrices of participants' recalls of the television episode (characterized using the same topic model used to capture the content of the episode; Figs.~\ref{fig:corrmats}B, C), we see a different structure than that of the episode itself.
-Although individual participants' recalls also exhibit periods of content stability punctuated by rapid change (visible in the block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off-diagonal correlations between the thematic content of temporally distant events are substantially stronger in participants' recalls than in the original episode.
+Although individual participants' recalls also exhibit periods of content stability punctuated by rapid change (visible in the blocky structure of the matrices in Fig.~\ref{fig:corrmats}C), the off-diagonal correlations between the thematic content of temporally distant events are substantially stronger in participants' recalls than in the original episode.
 Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode versus participants' recalls (Fig.~\ref{fig:reinstatement}A).
@@
 
-In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
+In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
 \cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats} and~\ref{fig:reinstatement}, revealing another key property of how we remember.
@@
 A key finding of that work was that people who watched the episode recalled high-level details reliably, and in a similar way across people, whereas people varied substantially in how they remembered low-level details.
-This suggests that our memory systems may prioritize certain types of information (e.g., high-level details) about our experience over others (e.g., low-level details) as we encode our experiences into memories or when we recount our prior experiences.
+This suggests that our memory systems may prioritize certain types of information (e.g., high-level details) about our experiences over others (e.g., low-level details) as we encode our experiences into memories or when we recount our prior experiences.
 
@@
 Each coordinate in this space reflects the conceptual content underlying a set of words, such that two different text passages that describe the same conceptual content (but using different words, levels of detail, etc.) are mapped onto nearby points in the word embedding space.
-When these models are sufficiently expressive and robust to solve the matching problem by modeling our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal18c}.
+When these models are sufficiently expressive and robust to solve the matching problem, by modeling our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal18c}.
 
@@
 Defining geometric spaces that can reliably capture the conceptual contents of our thoughts and experiences has become a major focus of computational science.
-One family of models that has shown particular promise over the past several decades is broadly referred to as \textit{word embedding models} (Fig.~\ref{fig:embedding-models}).
+One family of models that has shown particular promise over the past several decades is broadly referred to as \textit{text embedding models} (Fig.~\ref{fig:embedding-models}).
 The fundamental goal of these models is to take as input a sequence of words (text) and produce as output one or more feature vectors whose coordinates reflect their conceptual (semantic) content.
@@
 
-Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{DeerEtal90, LandDuma97} derived text embeddings using word co-occurence statistics.
+When text embedding models succeed at capturing semantic overlap between different words or text passages, they can provide an elegant means of solving the matching problem.
+For example, when two or more individuals describe a common experience, embedding the text of those descriptions in a high dimensional feature space abstracts away the specific word choices and grammatical choices in favor of the descriptions' underlying conceptual content.
+Further, geometric differences in those embeddings (e.g., the Euclidean distances between the corresponding vectors) can capture subtle differences in conceptual content that might not have been readily apparent in the original text.
+Of course, there are also limits on how accurately specific word embedding models characterize different conceptual aspects of a given text passage, and on how informative or meaningful those differences are from a psychological perspective.
+The field of text embedding has advanced rapidly, particularly in recent years, enabling more accurate and informative approaches to characterizing the conceptual content of arbitrary text passages.
+
+Early embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{DeerEtal90, LandDuma97} derived text embeddings using word co-occurence statistics.
 The logic driving this approach posits that documents that contain similar mixtures of words are likely to be conceptually similar, and words that appear in similar documents are also likely to be conceptually similar.
 To quantify these intuitions, \cite{DeerEtal90} first constructed \textit{word count matrices} (Fig.~\ref{fig:embedding-models}A) for several text corpora (each comprising many thousands of documents).
-Each entry in these matrices denotes the number of time a given document (row) contained a given word (column).
-They then used the Singular Value Decomposition (SVD) to decompose the number-of-documents $~\times~$ number-of-words word count matrix into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K~\times~K$ \textit{singular values} matrix, and a $K~\times$ number-of-unique-words \textit{word embeddings matrix}.
+Each entry in these matrices denotes the number of times a given document (row) contained a given word (column).
+They then used the Singular Value Decomposition (SVD) to decompose the number-of-documents $~\times~$ number-of-words word count matrix into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K~\times~K$ \textit{singular values matrix}, and a $K~\times$ number-of-unique-words \textit{word embeddings matrix}.
 The rows of the document embeddings matrix reflect the similarities between documents, and the columns of the word embeddings matrix reflect similarities between words.
@@
 \caption{\textbf{Three approaches to text embedding.}  \textbf{A.~Matrix factorization models.} Per-document (or per-stimulus) word counts and/or frequencies are collected into a word count matrix (left).
-To compute $K$-dimensional embeddings of the documents and words, the word count matrix is factorized (typically by computing its Singular Value Decomposition) into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values} matrix, and a $K~\times $ number-of-unique-words \textit{word embeddings matrix}.
+To compute $K$-dimensional embeddings of the documents and words, the word count matrix is factorized (typically by computing its Singular Value Decomposition) into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values matrix}, and a $K~\times $ number-of-unique-words \textit{word embeddings matrix}.
 \textbf{B.~Topic models.}  Fitting a topic model to a word count matrix (e.g., Panel A) reveals a set of latent \textit{topics}, or themes that pervade each document.
@@
 LDA defines a topic as a distribution of weights over words in the vocabulary.
-For example, if one takes the \textit{vocabulary} to be the top 10,000 most commonly used words in English (excluding stop words that are devoid of explicit semantic meaning such as ``and,'' ``or,'' ``the,'' ``but,'' etc.), then each topic comprises a vector of 10,000 non-negative weights that sums to one.
+For example, if one takes the \textit{vocabulary} to be the top 10,000 most commonly used words in English (excluding stop words that are devoid of explicit semantic meaning such as ``and,'' ``or,'' ``the,'' ``but,'' etc.), then each topic comprises a vector of 10,000 non-negative weights that sum to one.
 As shown in the cartoon on the left of the Panel (where the four top-weighted words from each of five hypothetical topics are displayed), a fruit-related topic (shown in purple) might weight heavily on words like ``apple,'' ``banana,'' ``pear,'' and ``orange,'' whereas words like ``cat,'' ``dog,'' ``dolphin,'' and ``lobster'' might receive relatively small weights.
@@
   \textit{``As the slavering mutant vampire \textbf{bat} swooped low to attack, its eyes ablaze, the explorer grunted and took a mighty swing of her trusty aluminum baseball \textbf{bat}.
-The powerful strike connected with a echoing metallic \textit{clink}, stunning the wretched creature before it scrambled about on the cave floor and flew away.
+The powerful strike connected with an echoing metallic \textit{clink}, stunning the wretched creature before it scrambled about on the cave floor and flew away.
 She didn't even \textbf{bat} an eye-- as the Dark Wizard Chiroptera continued to draw strength from the mystic crystals, such things had become increasingly common of late, nearly to the point of mundanity...''}
@@
 At different ponts in the passage, ``bat'' refers to an animal, a piece of sports equipment, and an eye-related verb.
-Because each of these uses of the word appear in the same (hypothetical) document, models driven solely by word co-occurrence statistics would blend these different meanings such that the embedding of ``bat'' was somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
+Because each of these uses of the word appears in the same (hypothetical) document, models driven solely by word co-occurrence statistics would blend these different meanings such that the embedding of ``bat'' was somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
 By contrast, treating each document as reflecting a blend of topics allows these uses to be disambiguated.
@@
 
+When the ``surface features'' of our experiences repeat at non-contiguous timepoints, or when one moment overlaps conceptually with others, this overlap can be characterized by text embedding models like those described above.
+In turn, this can help explain non-contiguous ``bumps'' in the quantum memory wave function (e.g., Figs.~\ref{fig:qwave}C, \ref{fig:reinstatement}B) and off-diagonal peaks in semantic temporal correlation matrices (e.g., Fig.~\ref{fig:corrmats}).
+
 \subsection*{Extensions to spatial, semantic, affective, and social representations}
@@
 Traditional approaches then ``clean up'' this distribution over possible mental states by taking its average (expected), maximum, or modal value.
-However, following the ``quantum wave function'' logic described in this manuscript, these multi-peaked receptive fields may provide a mechanism for associating non-contiguous representations.
+However, following the ``quantum wave function'' logic described in this manuscript, these multi-peaked response patterns may provide a mechanism for associating non-contiguous representations.
 What are some implications of this view for non-memory domains?
@@
 Rather, the set of stimuli likely to be encountered live within subspaces or on manifolds in the relevant representational space.
-The geometries of these manifolds provides insights into the statistical structure of the stimuli.
+The geometries of these manifolds provide insights into the statistical structure of the stimuli.
 They also provide potential insights into how the brain represents those stimuli, for example to the extent that the geometries of these manifolds may be leveraged to improve brain decoding.

### 2021-01-11 paper/old.tex
first attempt at addressing R1 concerns

--- 2019-01-16 main.tex
@@
 \usepackage{geometry}
-\usepackage{natbib}
+\usepackage[sort]{natbib}
 \usepackage{pxfonts}
 \usepackage{graphicx}
+\usepackage{setspace}
 \usepackage{hyperref}
+\usepackage{lineno}
+
+\doublespacing
+\linenumbers
 
@@
 How do our brains organize, retrieve, and act upon the ceaseless flow of incoming information and experiences?
-\cite{Tulv83} contends that episodic memory is like a sort of ``mental time travel,'' whereby we send some part of our mental state back along our autobiographical timeline to a specific moment in our past.
+\cite{Tulv83} contends that episodic memory is like a sort of ``mental time travel,'' whereby we send some part of our mental state back along our autobiographical timeline to a specific moment in our past~\citep[also see][]{Tulv72, HassEtal07b, SchaEtal07, MannEtal11}.
+This process has also been described as mentally ``jumping back in time''~\citep[e.g.,][]{HowaEtal12, FolkEtal18} to when we experienced what we are now remembering.
 This view has fundamental implications for how we study and build theories about memory: it suggests that we can match up specific moments of recollection with specific moments from the past.
-These implications extend to how we design episodic memory experiments and how we analyze the behavioral and neural data from those experiments in order to understand the neural and/or cognitive mechanisms that underly memory retrieval.
+These implications extend from how we design episodic memory experiments to how we analyze and interpret the behavioral and neural data from those experiments in order to understand the neural and/or cognitive basis of memory retrieval.
 At its core, characterizing episodic memory as mental time travel makes studying and understanding memory about matching.
 Specifically, the primary implied questions are about how and when we match up the thoughts (and/or brain activity patterns) we are experiencing at one moment with the thoughts (and/or brain activity patterns) we experienced at another previous moment.
-The challenge, then, is to understand the neural (structural) and cognitive (functional) mechanisms that characterize when, how, and why our memory systems allow us to relive those prior experiences.
+The implied challenge, then, is to understand the neural (structural) and cognitive (functional) mechanisms that characterize when, how, and why our memory systems allow us to relive those prior experiences.
 One way of illustrating how this framing misses out on key aspects of memory is to examine how experimentalists and theorists typically conceive of and model a mental construct called \textit{context}.
@@
 Defining context precisely (e.g., as distinct from content) is an ongoing challenge for our field, but theorists generally describe context as reflecting a mental combination of external cues (where you are, who you are with, background sensory information, etc.) and internal cues (thoughts, feelings, emotions, etc.) that uniquely define each moment~\citep[e.g., see review by][]{MannEtal15}.
-Reactivating the mental representation of the context associated with a given episode is what can give us the feeling of some ``part of ourselves'' re-experiencing the past.
+Reactivating the mental representation of the context associated with a given episode is what can give us the feeling of some ``part of ourselves'' re-experiencing the past~\citep[Tulving refers to this feeling as \textit{autonoetic awareness};][]{Tulv02a}.
 
-One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
-An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
+One practical way to distinguish content from context representations is to separate out mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
+An intuition for why contextual representations might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
 Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.
-This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12, AlyEtal18}.
+This notion, that our brains maintain parallel mental representations drifting at different timescales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12a, AlyEtal18}.
 Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}.
-This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
+This body of work suggests that different brain regions reflect \textit{temporal receptive windows} that characterize the timescale(s) to which each region is maximally sensitive.
+This research complements work on the \textit{temporal window of integration} in the speech and perception literatures~\citep[e.g.,][]{ViemWake91, GautEtal12, vanWEtal07}.
+The temporal window of integration refers to the duration over which physically distinct stimuli are treated (by some brain system or network) as a single percept or atomic unit.
+This systems-level work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
 Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
-These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
+In addition to this work suggesting that individual systems or neurons maintain representations at their preferred timescales, a number of studies also report neurons and small populations exhibiting \textit{temporal multiplexing}, whereby information at different timescales is reflected in different features of an individual neuron's (or population's) activity patterns, such as spike timing versus firing rate or local field potential oscillations at different frequencies~\citep{CaruEtal18, KnigEich13, WatrEtal13, PanzEtal10}.
+
+The above ideas about parallel neural representations that reflect information at multiple timescales have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02a, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
-Other recent work
-indicates that different aspects of the reinstatement process itself might also occur at multiple timescales.
-For example, \cite{LindEtal19} found that high-level semantic information was reinstated prior to low-level (e.g.
-perceptual) details in an associative memory task.
+Other recent work indicates that different aspects of the reinstatement process itself might also occur at multiple timescales.
+For example, \cite{LindEtal19} found that high-level semantic information was reinstated prior to low-level (e.g., perceptual) details in an associative memory task.
 Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.
 First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).
-Second, the studies suggest that members of this family of representations drifting at different rates become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.
+Second, the studies suggest that representations that reflect different timescales become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of timescales.
 These reactivations explain how our brains reactivate prior contexts when we remember the past.
@@
 This is not simply a matter of imprecision in mental time travel (i.e., a reflection of the low effective resolution with which we can revisit specific moments from our past).
-Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences.
-The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each drifting at a different timescale.
+Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences, analogous to a quantum wave function spreading its probability mass over space (see \textit{Defining the {\normalfont quantum memory wave function} metaphor and comparing it to the {\normalfont mental time travel} metaphor}).
+The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each reflecting information at a different timescale.
 The notion of thinking of any one particular moment from the past is incompatible with this multiple timescales view, in that our thoughts and brain activity patterns reflect information from a range of moments, even when we are not specifically engaged in remembering.
@@
 This is how we can integrate information across the \textit{many} discrete experiences relevant to ``now.''
+
+\section*{Defining the \textit{quantum memory wave function} metaphor and comparing it to the \textit{mental time travel} metaphor}
+Unlike in classical physics, where matter dutifully obeys Newton's laws, quantum physics postulates a different (though potentially complementary) set of rules governing how matter behaves and interacts.
+Perhaps the most famous equation in quantum physics is \cite{Schr26}'s equation describing the quantum \textit{wave function}~\citep[this work was also inspired by][]{deBr24}.
+Schr\"{o}dinger's equation describes, in probabilistic terms, the location of a particle in space, at a given time.
+In other words, it defines a function, $\Psi(x, t)$, such that the probability that a particle is located at position $x$ at time $t$ is equal to $|\Psi(x, t)|^2$.
+The central idea is that, although it may not be possible to know or predict the particle's precise position at a given moment, the wave function tells us how its location's probability mass extends over space.
+In turn, this allows us to use the wave function to predict behaviors or constraints of the system(s) in which the particle is behaving and interacting.
+Within the domain of episodic memory, for the purposes of describing the \textit{quantum memory wave function} metaphor, we can conceptualize the moments along our \textit{autobiographical timeline}~\citep{ArzyEtal09} as the spatial coordinates ($x$) in Schr\"{o}dinger's equation.
+Then our mental state at time $t$ may be conceptualized as spreading its probability mass over our autobiographical timeline.
+In other words, like Schr\"{o}dinger's equation, many episodic memory models~\citep[e.g.,][]{HowaKaha02a, PolyEtal09, ShanHowa12} define a function describing how our mental state at time $t$ reflects the blends of thoughts we experienced at other moments of our lives.
+Under this metaphor, the effect of mental operations we perform with respect to our autobiographical timeline, such as remembering past experiences or predicting what will happen in the future, is to shift the probability mass of our quantum memory wave function (Fig.~\ref{fig:qwave}).
+This metaphor is also related to work on quantum probability decision models~\citep[e.g.,][]{KhreEtal18} in that it extends similar concepts to the domain of episodic memory.
+
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/wave_functions}
+  \caption{\textbf{Memory wave functions.
+All panels.} Each panel displays a quantum memory wave function for a hypothetical individual at a particular moment (time $t=0$ indicated by $\blacktriangle$).
+The $x$-axes correspond to the position ($t$) along the individual's autobiographical timeline.
+The $y$-axes display how much the individual's thoughts at time $t = 0$ include information about the individual's past (or future) perceived or predicted experiences (i.e., at other times).
+Different colors denote different amounts of temporal blurring.
+All curves have been normalized to have a maximum value of 1 and a minimum value of 0.
+\textbf{A.
+Contextual drift (no explicit reinstatement).}  The blend of thoughts leading up to $t = 0$ reflects a recency-weighted average of information from the past.
+\textbf{B.
+Mental time travel.} This constitutes a special case of the quantum memory wave function framework, whereby the individual's thoughts reflect a recency-weighted blend of prior thoughts, along with thoughts specific to one remembered prior experience (occurring at the timepoint indicated by the $\bigstar$ symbol).
+\textbf{C.
+Quantum memory waves.}  In this example, the individual remembering the moment indicated by the $\bigstar$ symbol brings back a blend of thoughts similar to those shown in Panel B.
+In addition, they bring to mind information pertaining to other conceptually related (but temporally non-contiguous) moments, indicated by $\star$ symbols.
+\textbf{D.
+Generating predictions about the future.} To predict what will happen during a series of future events ($\diamond$ symbols) the individual spreads their thoughts over their prior experiences according to the expected relevance of those experiences to the future events.
+Local peaks in the curve denote prior experiences with high expected relevance.
+The left part of the panel (prior to $\blacktriangle$) illustrates how the individual's thoughts are spread out over their past, and the right part of the panel illustrates how the individual's thoughts are spread out over their (predicted) future.}
+\label{fig:qwave}
+\end{figure}
+
+The \textit{mental time travel} metaphor is defined by Tulving as ``the type of awareness experienced when one thinks back to a \textit{specific moment} in one's personal past and consciously recollects some prior episode or state as it was previously experienced''~\citep[Fig.~\ref{fig:qwave}B;][emphasis added]{WheeEtal97}.
+Whereas mental time travel refers to mentally revisiting one specific moment in addition to now (potentially with some temporal blurring), the quantum memory wave framework implies that episodic memory entails spreading our awareness over \textit{many} moments (Fig.~\ref{fig:qwave}C).
+In this way, the key distinction between these framings is whether remembering entails simultaneously revisiting one single moment versus many moments.
+The two frameworks converge when we engage in \textit{chronesthesia}~\citep{Tulv02b}-- i.e., when we are consciously aware of subjective time.
+Analogous to how measuring the physical location of a particle collapses its quantum wave function at the moment of measurement, engaging in chronesthesia can collapse our quantum memory wave function such that our thoughts temporarily converge on the moment we are revisiting.
+This wave function expands again when our mental state changes in the next moment.
+
+The quantum memory wave and mental time travel metaphors also diverge in explaining how we might predict what will happen in the future.
+The quantum memory wave framing suggests that our predictions about the future draw from information pertaining to a blend of our prior experiences, weighted by their expected relevance to the future event in question (Fig.~\ref{fig:qwave}D).
+To the extent that the mental time travel metaphor assumes that we can mentally visit just one other time (or time range) in addition to now~\citep[e.g.,][]{SzpuTulv11}, it is not clear how this sort of integration over temporally separated prior experiences might occur.
+Specifically, integrating over those experiences would seem to require some mechanism for associating and simultaneously revisiting them while also casting some part of our thoughts forward to the future.
+
+While the studies presented in the \textit{Introduction and overview} (on temporal receptive windows, the temporal window of integration, time cells, and temporal multiplexing) present strong empirical evidence that our brains maintain parallel representations at multiple timescales, questions remain about how or whether we experience those representations at a conscious level.
+For example, are we consciously aware of spreading our mental states over many moments from our past?
+Or is our subjective experience of remembering more like mentally time traveling back to a single ``hybrid'' moment that blends information from several of our (actual) prior experiences?
 
@@
 One insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \cite{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
-The authors used topic models~\citep{BleiEtal03} applied to human-generated annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.
-These \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode.
+The authors used topic models~\citep{BleiEtal03} applied to human-generated annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors (see \textit{Defining the geometries of ongoing thoughts and memories}).
+Each shot's annotation contained the following information: a brief narrative description of what was happening, the location where the scene took place, whether that location was indoors or outdoors, the names of all on-screen characters, the names of the character(s) who were in focus in the camera shot, the name(s) of any character(s) who spoke during the scene, the camera angle of the shot, a transcription of any on-screen text, and whether or not there was any music present in the auditory background~\citep[full details may be found in][]{ChenEtal17}.
+The resulting \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode (Fig.~\ref{fig:embedding-models}B).
 When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change.
@@
 If we examine the temporal autocorrelation matrices of participants' recalls of the television episode (characterized using the same topic model used to capture the content of the episode; Figs.~\ref{fig:corrmats}B, C), we see a different structure than that of the episode itself.
-Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off-diagonal correlations between the thematic content of temporally distant events are substantially stronger in participants' recalls than in the original episode.
-Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode and participants' recalls (Fig.~\ref{fig:reinstatement}A).
+Although individual participants' recalls also exhibit periods of content stability punctuated by rapid change (visible in the blocky structure of the matrices in Fig.~\ref{fig:corrmats}C), the off-diagonal correlations between the thematic content of temporally distant events are substantially stronger in participants' recalls than in the original episode.
+Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode versus participants' recalls (Fig.~\ref{fig:reinstatement}A).
 Whereas the autocorrelations in the original episode's content rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around an average of 0.2.
@@
 To understand which specific events are described (during recall) as more similar than what their original experiences (as characterized by their topic vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.
-Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were characterized in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
-The example event explored in the figure comprises a scene where Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another victim has been murdered.
-The language used by human annotators describing the moment-by-moment contents of the scene used a similar mix of themes only for moments in the episode that were temporally proximal to the event in question (gray line).
+Figure~\ref{fig:reinstatement}B displays, for a representative example recalled event, which other parts of the episode were characterized in a similar way by the participants when they recounted the episode later (black and colored lines), and which other parts of the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).
+In the example event explored in the figure, Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another person has committed suicide.
+Later the viewer learns that the suicides were actually murders.
+The language used by human annotators describing the moment-by-moment contents of the event used a similar mix of themes only for temporally nearby parts of the episode (gray line).
 However, the way participants described what happened when they \textit{recounted} the episode later revealed a much different pattern.
-Specifically, their language mirrored that used when they were recalling other murders, and other moments when Sherlock and John were bantering with or about each other (Fig.~\ref{fig:reinstatement}C).
-The way they described the example event did \textit{not} mirror the way they described other semantically (but not conceptually) related events, such as events involving unrelated violence, or Sherlock and John interacting with other characters about other subjects.
+Specifically, their language mirrored that used when they were recalling other related murders, and other moments when Sherlock and John were bantering with or about each other (Fig.~\ref{fig:reinstatement}C).
+
+The way they described the example target event did \textit{not} mirror the way they described other events that were not directly narratively linked with the target event, even when those other events overlapped semantically.
+For example, thematic elements of events involving unrelated violence, or Sherlock and John interacting with other characters about other subjects, tended not to feature in participants' recountings of the target event.
 Taken together, this suggests that these specific events were associated in participants' memories, perhaps contributing to (or reflecting) their understanding that those events were linked in the episode's narrative.
 
-\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement.png} \caption{\textbf{Content drift in a television episode and participants' recalls of that episode.
+One concrete example of this phenomenon is that early in the episode, including during the target event, Sherlock and John appear to be investigating a series of suicides.
+Later in the episode, it becomes clear that the apparent suicides are actually murders.
+The annotations refer to early deaths (including that shown in Fig.~\ref{fig:reinstatement}B) as suicides, whereas participants often refer to those early deaths as ``murders'' even though they could not have known that they were murders early on in the episode.
+In this way, participants' experiences \textit{subsequent to} a target remembered event may distort its memory by incorporating information learned during other experiences.
+
+\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Content drift in a television episode and participants' recalls of that episode.
 A.
@@
 
-In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
-\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figure~\ref{fig:trajectories}, revealing another key property of how we remember.
+In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
+\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats} and~\ref{fig:reinstatement}, revealing another key property of how we remember.
 Although there are substantial individual differences in the specific words people use to describe their shared experience of watching the same television episode, and how much detail they recollected, the overall shapes of those recalled experience trajectories are highly similar across people.
-This has potentially exciting implications for what it means to communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.
+This has potentially exciting implications for what it means to communicate memories to other people: to effectively communicate, we often need to convey the ``essence'' (gross overall shape) of our experience.
 It also has implications for how we learn by analogy: perhaps events that share schemas~\citep[e.g., ][]{BaldEtal18}  unfold in similar ways (e.g., have similar shapes).
@@
 
+In their paper, \cite{HeusEtal18c} use the geometric shapes of experience trajectories (through word embedding space) to functionally distinguish high-level essential details of the experience from low-level (non-essential) details.
+When the trajectory exhibits a sharp change in direction, this indicates a rapid shift in the thematic content of the episode.
+\cite{HeusEtal18c} used hidden Markov models to detect these rapid shifts and segment the trajectories into discrete \textit{events}.
+The high-level (essential) details of the episode are reflected in the shape defined by the sequence of coordinates of its events.
+These are captured by low spatial frequency properties of the trajectory's shape.
+Low-level details are characterized by (typically smaller-scale) within-event geometric patterns, which are captured by the high spatial frequency properties of the trajectory's shape.
+A key finding of that work was that people who watched the episode recalled high-level details reliably, and in a similar way across people, whereas people varied substantially in how they remembered low-level details.
+This suggests that our memory systems may prioritize certain types of information (e.g., high-level details) about our experiences over others (e.g., low-level details) as we encode our experiences into memories or when we recount our prior experiences.
+
+\subsection*{The \textit{matching problem}}
+The trajectories through word embedding space displayed in Figure~\ref{fig:trajectories} demonstrate that topic models are sufficiently expressive and robust to provide one solution to the \textit{matching problem} in memory analysis.
+The matching problem refers to discovering an alignment in the labels between experienced and remembered events.
+Traditional memory paradigms most commonly studied over the past century~\citep{Kaha12} ask participants to study lists of \textit{items} (often randomly selected words) and then engage in a memory task, such as verbally recalling any words the participant remembers studying, in any order.
+Solving the matching problem in such experiments is trivial: if we treat each studied word on a list as an ``event,'' then we can assume that the participant is remembering a word's event when they verbalize that word during their memory test.
+In other words, matching entails finding the exact match between a word the participant is verbalizing and that same word on the list that they studied.
+
+For real-world memory phenomena, and in many naturalistic memory paradigms, solving the matching problem is more complex.
+People often experience real-world events (and scenes from movies or stories) as rich sensory experiences that derive deep meaning from surrounding moments, other prior experiences, emotions, goals, high-order understanding of the current situation, and so on.
+Many of these aspects of our real-world experiences are not immediately verbalizable-- and even when we are able to describe some specific aspects of what we are experiencing, there are typically many possible ways we could have done so (e.g., varying in level of detail, specific wording, objectivity versus subjectivity, contextual details, etc.).
+Further, because the way we experience each moment depends in part on our unique prior experiences, even ostensibly shared experiences can \textit{seem} different to each individual.
+The way we recount a given experience is often a reflection of our own idiosyncratic experience of the original event, and is colored and distorted in an idiosyncratic way by our memory systems.
+This distorted memory must then be verbalized or communicated, which is itself an idiosyncratic (and potentially lossy, in the information theoretic sense) process.
+Solving the matching problem under these constraints poses a substantial theoretical and methodological challenge.
+Specifically, matching requires building explicit models of the dynamic conceptual content of our experiences and memories, and then estimating when and how that experienced and remembered content aligns.
+The approach featured in Figure~\ref{fig:trajectories} uses a word embedding space that is defined by a topic model~\citep{BleiEtal03}.
+Each coordinate in this space reflects the conceptual content underlying a set of words, such that two different text passages that describe the same conceptual content (but using different words, levels of detail, etc.) are mapped onto nearby points in the word embedding space.
+When these models are sufficiently expressive and robust to solve the matching problem, by modeling our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal18c}.
+
+\subsection*{Defining the geometries of ongoing thoughts and memories}
+Defining geometric spaces that can reliably capture the conceptual contents of our thoughts and experiences has become a major focus of computational science.
+One family of models that has shown particular promise over the past several decades is broadly referred to as \textit{word embedding models} (Fig.~\ref{fig:embedding-models}).
+The fundamental goal of these models is to take as input a sequence of words (text) and produce as output one or more feature vectors whose coordinates reflect their conceptual (semantic) content.
+For example, when two conceptually similar documents are passed through these models, the goal is to produce two feature vectors that are closer (e.g., in Euclidean distance) than the feature vectors for two dissimilar documents.
+
+Early word embedding models such as \textit{Latent Semantic Analysis}~\citep[LSA;][]{DeerEtal90, LandDuma97} derived text embeddings using word co-occurence statistics.
+The logic driving this approach posits that documents that contain similar mixtures of words are likely to be conceptually similar, and words that appear in similar documents are also likely to be conceptually similar.
+To quantify these intuitions, \cite{DeerEtal90} first constructed \textit{word count matrices} (Fig.~\ref{fig:embedding-models}A) for several text corpora (each comprising many thousands of documents).
+Each entry in these matrices denotes the number of times a given document (row) contained a given word (column).
+They then used the Singular Value Decomposition (SVD) to decompose the number-of-documents $~\times~$ number-of-words word count matrix into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K~\times~K$ \textit{singular values matrix}, and a $K~\times$ number-of-unique-words \textit{word embeddings matrix}.
+The rows of the document embeddings matrix reflect the similarities between documents, and the columns of the word embeddings matrix reflect similarities between words.
+
+A related approach, termed \textit{Word Association Spaces}~\citep[WAS;][]{NelsEtal04, SteyEtal04} replaces the word count matrix with data from a large free-association experiment.
+In the experiment, each participant was asked to free-associate (i.e., say the first word that came to mind) in response to a series of cue words.
+The authors then constructed a number-of-cue-words $~\times~$ number-of-unique-response-words matrix whose entries reflected the proportion of times each cue word (row) produced a given response word (column) across participants.
+After using SVD to decompose this \textit{word frequency matrix}, they obtained low-dimensional embeddings of each cue word and response word.
+In one sense, WAS might provide a more accurate and behaviorally relevant reflection of how people think about the meanings of words than LSA, since nearby WAS vectors reflect either cue words that produced similar responses, or response words that were evoked by similar cues.
+In other words, WAS derives embeddings directly from people's semantic judgement behaviors.
+However, because WAS is based on a large amount of experimental data, updating the embeddings requires collecting new data, which can be slow and costly.
+As the target application or population diverges from the experimental population (e.g., over time, across cultures, demographics, etc.) the WAS embeddings may become less accurate or relevant.
+By contrast, LSA may be re-fit relatively quickly to any new corpus, enabling the resulting embeddings to stay current.
+Further, because LSA may be applied to any corpus, word embeddings may be easily computed and/or compared across languages, times, subject areas, etc.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.8\textwidth]{figs/word_embedding_models}
+\caption{\textbf{Three approaches to text embedding.}  \textbf{A.~Matrix factorization models.} Per-document (or per-stimulus) word counts and/or frequencies are collected into a word count matrix (left).
+To compute $K$-dimensional embeddings of the documents and words, the word count matrix is factorized (typically by computing its Singular Value Decomposition) into the product of a number-of-documents $\times~K$ \textit{document embeddings matrix}, a $K \times K$ \textit{singular values matrix}, and a $K~\times $ number-of-unique-words \textit{word embeddings matrix}.
+\textbf{B.~Topic models.}  Fitting a topic model to a word count matrix (e.g., Panel A) reveals a set of latent \textit{topics}, or themes that pervade each document.
+The topics are specified in a $K~\times$ number-of-words \textit{topics matrix}, and the per-document themes are specified in a number-of-documents $\times~K$ \textit{topic proportions matrix}.
+A given document's topic proportions reflect the specific mix of themes that are reflected in that document (bar graph on the left).
+These proportions are discovered automatically by estimating the topic from which each word in the document is most probably drawn (highlighted words in \textit{Document text}).
+\textbf{C.~Deep learning models.}  A wide variety of deep learning approaches to text embedding have been proposed.
+Broadly, these approaches often combine a common set of elements, including \textit{transformers}~\citep[diagram adapted from][]{ViswEtal17}, \textit{deep averaging networks}~\citep[diagram adapted from][]{IyyeEtal15}, \textit{recurrent neural networks}~\citep[diagram adapted from][]{IyyeEtal15}, and \textit{deep autoencoders}~\citep[diagram adapted from][]{YousHame17}.}
+\label{fig:embedding-models}
+\end{figure}
+
+\textit{Topic models}, such as \textit{Latent Dirichlet Allocation}~\citep[LDA;][]{BleiEtal03} represent an attempt to characterize the latent \textit{topics} or themes that documents comprise (Fig.~\ref{fig:embedding-models}B).
+Whereas matrix factorization approaches represent semantic meanings using low-dimensional approximations of word co-occurrences or free association frequencies, topic models leverage these properties of documents (or behaviors) to explain \textit{why} words co-occur.
+LDA defines a topic as a distribution of weights over words in the vocabulary.
+For example, if one takes the \textit{vocabulary} to be the top 10,000 most commonly used words in English (excluding stop words that are devoid of explicit semantic meaning such as ``and,'' ``or,'' ``the,'' ``but,'' etc.), then each topic comprises a vector of 10,000 non-negative weights that sum to one.
+As shown in the cartoon on the left of the Panel (where the four top-weighted words from each of five hypothetical topics are displayed), a fruit-related topic (shown in purple) might weight heavily on words like ``apple,'' ``banana,'' ``pear,'' and ``orange,'' whereas words like ``cat,'' ``dog,'' ``dolphin,'' and ``lobster'' might receive relatively small weights.
+Given a set of $K$ topic vectors, topic models compute the blend of topics most probably reflected in a given document-- i.e., its \textit{topic proportions}.
+These topic proportion estimates are made by examining how different topics weight each of the words in the document, and by assuming that documents tend to reflect few topics (rather than many topics).
+Given only a word count matrix as input (just as in LSA), topic models automatically discover a $K~\times~$ number-of-words \textit{topics matrix}, which describes how each topic (row) weights each word in the vocabulary (column), and a number-of-documents $~\times~K$ \textit{topic proportions matrix}, which describes each document's topic proportions.
+The rows of the topic proportions matrix may be treated as low-dimensional embeddings of the set of documents.
+
+A key advantage of topic models over matrix factorization models is their ability to explain potentially ambiguous word usage.
+For example, consider the use of the word ``bat'' in the following illustrative passage:
+\begin{quote}
+  \textit{``As the slavering mutant vampire \textbf{bat} swooped low to attack, its eyes ablaze, the explorer grunted and took a mighty swing of her trusty aluminum baseball \textbf{bat}.
+The powerful strike connected with an echoing metallic \textit{clink}, stunning the wretched creature before it scrambled about on the cave floor and flew away.
+She didn't even \textbf{bat} an eye-- as the Dark Wizard Chiroptera continued to draw strength from the mystic crystals, such things had become increasingly common of late, nearly to the point of mundanity...''}
+  \end{quote}
+At different ponts in the passage, ``bat'' refers to an animal, a piece of sports equipment, and an eye-related verb.
+Because each of these uses of the word appears in the same (hypothetical) document, models driven solely by word co-occurrence statistics would blend these different meanings such that the embedding of ``bat'' was somewhat similar to other animals \textit{and} other pieces of sports equipment \textit{and} other eye-related verbs.
+By contrast, treating each document as reflecting a blend of topics allows these uses to be disambiguated.
+If other documents that mention the word ``bat'' describe vampires, swooping, or other mammals, but \textit{not} sports equipment or eye-related verbs, then a topic model could explain this pattern by inferring that multiple topics place substantial weight on the word ``bat.''  Another strength of topic models is their interpretability.
+Because the models directly output each document's topic proportions along with each topic's weights on all words in the vocabulary, the general themes that each topic reflects are often intuitive and the per-document topic proportions may be directly related to those themes.
+
+Both the matrix factorization models and topic model described above are each examples of \textit{bag of words} models.
+Given a document, these models will generate the same embeddings for any random re-ordering of the same words.
+This over-simplifies the intended conceptual content of documents, where word order often conveys important meaning.
+An alternative framing of the text embedding problem is to use \textit{sequence} models that explicitly account for word order effects.
+Most modern sequence-based models are implemented using deep neural networks (Fig.~\ref{fig:embedding-models}C).
+
+The \textit{Neural Probabilistic Language Model}~\citep[NPLM;][]{BengEtal03} was one of the earliest deep learning-based sequence models.
+NLPM derives word embeddings for each word in the vocabulary (such that nearby embeddings reflect semantically related words), and then models word sequences according to their meanings.
+This model was inspired in part by $n$-gram models~\citep{BrowEtal92}, which estimate the probability of the next word in a sequence given the $n$ words that came before it.
+Effectively, NLPM attempts to learn arbitrarily long sequences.
+In a related approach, \cite{MnihHint09} proposed the \textit{Scalable Hierarchical Distributed Language Model}, which learns sequences of variable length $n$-grams to achieve greater explanatory power.
+
+The next major advances in word embeddings came from attempts to more precisely capture the geometric relations between words.
+For example, if each word's semantic meaning is captured by a feature vector, then in principle one could perform vector algebra on different words' embeddings to produce new concepts that combine their meanings.
+For example, subtracting the vector for ``man'' from that of ``king'' and then adding the vector for ``woman'' might result in a new feature vector similar to that of ``queen''~\citep{MikoEtal13b}.
+\cite{MikoEtal13a} developed the deep learning model \textit{word2vec} to explicitly optimize the accuracy of these vector operations.
+Whereas word2vec is a bag of words model, \textit{Global Vectors for Word Representation}~\citep[GloVe;][]{PennEtal14} extends this approach to model each word's local (temporal) context~\citep[also see][]{PeteEtal18}.
+
+Several recent models have begun to move beyond single word-level embeddings.
+For example, \textit{SkipThought}~\citep{KiroEtal15}, \textit{Universal Sentence Encoder}~\citep{CerEtal18}, \textit{InferSent}~\citep{ConnEtal18}, and \textit{RandSent}~\citep{WietKiel19} each compute sentence-level embeddings.
+The central insight of these approaches is that capturing the nuanced meaning of a particular concept often requires \textit{several} words that have been organized into a complete sentence or sentence-like grammatical construct.
+
+Cutting-edge approaches to word embeddings that have achieved the best-to-date performance on modern natural language processing benchmarks are built using enormous deep neural networks often with many billions of parameters.
+Because these models are very computationally expensive to fit, an important discovery has been that many layers in these models may be pre-trained without substantially affecting performance, even on new tasks.
+One of the first of this new class of model, \textit{Bidirectional Encoder Representations from Transformers}~\citep[BERT;][]{DevlEtal18}, uses a pre-trained deep neural network, whose weights in a single output layer may be fine-tuned to suit a wide range of new tasks.
+The latest additions to this class of models are called \textit{Generative Pre-trained Transformers}~\citep[GPT;][]{RadfEtal18, RadfEtal19, BrowEtal20}.
+
 \subsection*{Extensions to spatial, semantic, affective, and social representations}
+The way we navigate through \textit{thought space} (i.e., word embedding spaces; Figs.~\ref{fig:trajectories}, \ref{fig:embedding-models}) differs from how we navigate in physical spatial environments.
+First, we cannot teleport in real space, whereas our thoughts can exhibit rapid jumps between different non-contiguous regions of word embedding space.
+Second, when two people visit the same fixed physical landmark, they necessarily visit the same physical location.
+When two people describe a shared experience, their idiosyncratic thoughts, goals, prior experiences, etc.\ can lead them to process and remember the experience differently, thereby visiting different locations within thought space.
+Third, each time we revisit a fixed physical landmark, we (by definition) revisit the same spatial location.
+In other words, it is possible to visit the same spatial location multiple times.
+However, we cannot revisit our prior experiences in this way.
+Rather, we only get to experience each moment once; subsequent ``visits'' to a prior experience must occur through a different medium than the original experience (e.g., our memory, another person’s memory, a physical recording, a written account, etc.).
+Fourth, we can revisit different aspects of the same prior experience on different ``visits''-- e.g., when we remember a specific event we can focus on the physical occurrence in one remembering, the emotional occurence in another remembering, the social implications in another remembering, etc.
+We can also revisit the same experience in different levels of detail or depth.
+There are no deep analogs of these phenomena in spatial navigation.
+Finally, \textit{physically} (at a macro scale), we are always located at a single moment and in a single location.
+The main argument of the quantum memory wave function framework I propose here is that we may be \textit{mentally} spread over many times or locations simultaneously.
+For example, this provides a way for us to integrate information from multiple prior experiences into a single decision or conceptualization, even if those experiences were not temporally contiguous.
+
+Many of these differences between physical and mental navigation also apply to other mental domains, including how we \textit{conceptualize} space~\citep[e.g., when we trace out a potential route in our mind;][]{GautvanW16, GautEtal18, ArzyScha19}.
 Just as we mentally spread ourselves across a distribution of times when we remember, we can perform similar mental operations when we conceptualize and retrieve other types of information.
 Further, these distributions may be estimated using existing methods.
-For example, consider how we use pattern classifiers~\citep{NormEtal06b} to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, affective~\citep{ChanEtal18}, and social~\citep{MeyeEtal18} content being mentally manipulated.
+For example, consider how we use pattern classifiers~\citep{NormEtal06b} to connect neural activity patterns to specific semantic~\citep{PolyEtal05a, MitcEtal08a, MannEtal12}, spatial~\citep{MillEtal13}, affective~\citep{ChanEtal18}, and social~\citep{MeyeEtal18} content being mentally manipulated.
 A prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} concept or mental state.
-Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states.
+Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading the probability masses of their predictions over multiple possible mental states, including non-contiguous mental states.
 Traditional approaches then ``clean up'' this distribution over possible mental states by taking its average (expected), maximum, or modal value.
-However, following the ``quantum wave function'' logic described in this manuscript, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.
+However, following the ``quantum wave function'' logic described in this manuscript, these multi-peaked response patterns may provide a mechanism for associating non-contiguous representations.
 What are some implications of this view for non-memory domains?
@@
 In other words, although place cells encode an animal's location within an environment, even for a fixed cell and environment there is some variability with respect to where precisely within that environment the cell modulates its firing rate.
-These studies suggest that variability in the set of locations to which a place cell responds is not merely a reflection of noise or imprecision.
+Further, a number of studies have suggested that some neurons respond to distinct non-contiguous places~\citep{FentEtal08, RichEtal14b, LeeEtal19, DerdEtal09, GrieEtal20} and times~\citep{PastEtal08}.
+Taken together, these studies suggest that variability in the set of locations and/or times to which a neuron responds is not merely a reflection of noise or imprecision.
 Rather, this variability might have some functional relevance in terms of facilitating the learning of paths around obstacles and towards goals.
@@
 Rather, the set of stimuli likely to be encountered live within subspaces or on manifolds in the relevant representational space.
-The geometry of these manifolds provides insights into the statistical structure of the stimuli.
-It also provides potential insights into how the brain represents those stimuli, for example to the extent that the geometries of these manifolds may be leveraged to improve brain decoding.
+The geometries of these manifolds provide insights into the statistical structure of the stimuli.
+They also provide potential insights into how the brain represents those stimuli, for example to the extent that the geometries of these manifolds may be leveraged to improve brain decoding.
 The shapes of these manifolds can potentially provide insights into how our thoughts are distributed over representational spaces.

### 2021-01-12 paper/main.tex
updated bibliography, draft of edits to address R1 comments, updated diff files

@@
 
-One practical way to distinguish content from context representations is to separate out mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.
+One practical way to distinguish content from context representations is to separate out mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, FolkEtal18}.
 An intuition for why contextual representations might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
@@
 The temporal window of integration refers to the duration over which physically distinct stimuli are treated (by some brain system or network) as a single percept or atomic unit.
-This systems-level work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
+This systems-level work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference (e.g., time elapsed since starting to run on an exercise wheel), whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08b, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}.
 Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.
-In addition to this work suggesting that individual systems or neurons maintain representations at their preferred timescales, a number of studies also report neurons and small populations exhibiting \textit{temporal multiplexing}, whereby information at different timescales is reflected in different features of an individual neuron's (or population's) activity patterns, such as spike timing versus firing rate or local field potential oscillations at different frequencies~\citep{CaruEtal18, KnigEich13, WatrEtal13, PanzEtal10}.
+In addition to this work suggesting that individual systems or neurons maintain representations at their preferred timescales, a number of studies also report neurons and small populations exhibiting \textit{temporal multiplexing}, whereby information at different timescales is reflected in different features of an individual neuron's (or population's) activity patterns, such as spike timing versus firing rate or local field potential oscillations at different frequencies~\citep{CaruEtal18, KnigEich13, WatrEtal13b, PanzEtal10}.
 
-The above ideas about parallel neural representations that reflect information at multiple timescales have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02a, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.
+The above ideas about parallel neural representations that reflect information at multiple timescales have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02a, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang19}.
 In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.
@@
 
-One insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \cite{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
+One insight into this distinction comes from a recent study by \cite{HeusEtal21}, which analyzed data collected by \cite{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.
 The authors used topic models~\citep{BleiEtal03} applied to human-generated annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors (see \textit{Defining the geometries of ongoing thoughts and memories}).
@@
 When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change.
-This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \cite{HeusEtal18c} performed detailed analyses of these putative \textit{events} and how they are remembered.
+This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \cite{HeusEtal21} performed detailed analyses of these putative \textit{events} and how they are remembered.
 Notably, the off-diagonal entries of the temporal correlation matrix (which reflect content similarity between temporally distant events in the episode) are nearly all close to 0.
@@
 Nonetheless, when considering solely the semantic content of each moment of the episode, those long timescale associations are nearly entirely absent.
-To the extent that this finding extends to real-world experiences, it comports with Heraclitus' notion that one cannot step in the same river twice~\citep{Hera}.
+To the extent that this finding extends to real-world experiences, it comports with Heraclitus' notion that one cannot step in the same river twice~\citep{HeraCE}.
 
-\begin{figure}[tp] \centering \includegraphics[width=0.8\textwidth]{figs/pres_rec_corrmats_sherlock} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal18c}, where full details may be found.
+\begin{figure}[tp] \centering \includegraphics[width=0.8\textwidth]{figs/pres_rec_corrmats_sherlock} \caption{\textbf{Temporal correlation matrices of the dynamic content of a television episode and people's recalls of that episode.}  The panels of this figure are adapted from \cite{HeusEtal21}, where full details may be found.
 \textbf{A.
@@
 \section*{The shapes of remembered experiences}
-The study by \cite{HeusEtal18c} described above suggests one way of characterizing the dynamic content of complex naturalistic stimuli like television episodes.
+The study by \cite{HeusEtal21} described above suggests one way of characterizing the dynamic content of complex naturalistic stimuli like television episodes.
 The sequence of topic vectors over the course of the episode traces out a \textit{topic trajectory} through a high-dimensional feature space.
@@
 The arrows reflect the directions of the average transitions taken by participants whose trajectories crossed that part of topic space; blue arrows denote directions of reliable agreement across participants.
-Figure adapted from \cite{HeusEtal18c}, where additional details may be found.}
+Figure adapted from \cite{HeusEtal21}, where additional details may be found.}
 \label{fig:trajectories}
@@
 In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
-\cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats} and~\ref{fig:reinstatement}, revealing another key property of how we remember.
+\cite{HeusEtal21} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats} and~\ref{fig:reinstatement}, revealing another key property of how we remember.
 Although there are substantial individual differences in the specific words people use to describe their shared experience of watching the same television episode, and how much detail they recollected, the overall shapes of those recalled experience trajectories are highly similar across people.
@@
 
-In their paper, \cite{HeusEtal18c} use the geometric shapes of experience trajectories (through word embedding space) to functionally distinguish high-level essential details of the experience from low-level (non-essential) details.
+In their paper, \cite{HeusEtal21} use the geometric shapes of experience trajectories (through word embedding space) to functionally distinguish high-level essential details of the experience from low-level (non-essential) details.
 When the trajectory exhibits a sharp change in direction, this indicates a rapid shift in the thematic content of the episode.
-\cite{HeusEtal18c} used hidden Markov models to detect these rapid shifts and segment the trajectories into discrete \textit{events}.
+\cite{HeusEtal21} used hidden Markov models to detect these rapid shifts and segment the trajectories into discrete \textit{events}.
 The high-level (essential) details of the episode are reflected in the shape defined by the sequence of coordinates of its events.
@@
 Low-level details are characterized by (typically smaller-scale) within-event geometric patterns, which are captured by the high spatial frequency properties of the trajectory's shape.
-A key finding of that work was that people who watched the episode recalled high-level details reliably, and in a similar way across people, whereas people varied substantially in how they remembered low-level details.
+A key finding of that work was that people who watched the episode recalled high-level details reliably, and in a similar way across people~\citep[also see][]{MortEtal20}, whereas people varied substantially in how they remembered low-level details.
 This suggests that our memory systems may prioritize certain types of information (e.g., high-level details) about our experiences over others (e.g., low-level details) as we encode our experiences into memories or when we recount our prior experiences.
@@
 Each coordinate in this space reflects the conceptual content underlying a set of words, such that two different text passages that describe the same conceptual content (but using different words, levels of detail, etc.) are mapped onto nearby points in the word embedding space.
-When these models are sufficiently expressive and robust to solve the matching problem, by modeling our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal18c}.
+When these models are sufficiently expressive and robust to solve the matching problem, by modeling our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal21}.
 
@@
 One of the first of this new class of model, \textit{Bidirectional Encoder Representations from Transformers}~\citep[BERT;][]{DevlEtal18}, uses a pre-trained deep neural network, whose weights in a single output layer may be fine-tuned to suit a wide range of new tasks.
-The latest additions to this class of models are called \textit{Generative Pre-trained Transformers}~\citep[GPT;][]{RadfEtal18, RadfEtal19, BrowEtal20}.
+The latest additions to this class of models are called \textit{Generative Pre-trained Transformers}~\citep[GPT;][]{RadfEtal19, BrowEtal20}.
 
@@
 In other words, although place cells encode an animal's location within an environment, even for a fixed cell and environment there is some variability with respect to where precisely within that environment the cell modulates its firing rate.
-Further, a number of studies have suggested that some neurons respond to distinct non-contiguous places~\citep{FentEtal08, RichEtal14b, LeeEtal19, DerdEtal09, GrieEtal20} and times~\citep{PastEtal08}.
+Further, a number of studies have suggested that some neurons respond to distinct non-contiguous places~\citep{FentEtal08, RichEtal14b, LeeEtal19, DerdEtal09, GrieEtal20} and times~\citep{PastEtal08b}.
 Taken together, these studies suggest that variability in the set of locations and/or times to which a neuron responds is not merely a reflection of noise or imprecision.
@@
 I thank Luke Chang, Janice Chen, Paxton Fitzpatrick, Andrew Heusser, Christopher Honey, Caroline Lee, Talia Manning, Lucy Owen, Matthijs van der Meer, and Kirsten Ziman for feedback and scientific discussions.
-I also thank Janice Chen, Yuan Chang Leong, Christopher Honey, Chung Yong, Kenneth Norman, and Uri Hasson for sharing the data re-analyzed here~\citep{ChenEtal17}, and Paxton Fitzpatrick and Andrew Heusser for contributing analysis code that I adapted from a prior publication~\citep{HeusEtal18c}.
+I also thank Janice Chen, Yuan Chang Leong, Christopher Honey, Chung Yong, Kenneth Norman, and Uri Hasson for sharing the data re-analyzed here~\citep{ChenEtal17}, and Paxton Fitzpatrick and Andrew Heusser for contributing analysis code that I adapted from a prior publication~\citep{HeusEtal21}, from which Figures~\ref{fig:corrmats} and~\ref{fig:trajectories} are also adapted.
+I also note that I presented some of the work in this manuscript in a talk at the 2019 Context and Episodic Symposium.
 This work was supported in part by NSF EPSCoR Award Number 1632738.
@@
 \bibliographystyle{apa}
-\bibliography{CDL-bibliography/memlab}
+\bibliography{CDL-bibliography/cdl}
 \end{document}

### 2021-01-12 paper/old.tex
updated bibliography, draft of edits to address R1 comments, updated diff files

@@
 I thank Luke Chang, Janice Chen, Paxton Fitzpatrick, Andrew Heusser, Christopher Honey, Caroline Lee, Talia Manning, Lucy Owen, Matthijs van der Meer, and Kirsten Ziman for feedback and scientific discussions.
-I also thank Janice Chen, Yuan Chang Leong, Christopher Honey, Chung Yong, Kenneth Norman, and Uri Hasson for sharing the data re-analyzed here~\citep{ChenEtal17}, and Paxton Fitzpatrick and Andrew Heusser for contributing analysis code that I adapted from a prior publication~\citep{HeusEtal18c}.
+I also thank Janice Chen, Yuan Chang Leong, Christopher Honey, Chung Yong, Kenneth Norman, and Uri Hasson for sharing the data re-analyzed here~\citep{ChenEtal17}, and Paxton Fitzpatrick and Andrew Heusser for contributing analysis code that I adapted from a prior publication~\citep{HeusEtal21}.
 This work was supported in part by NSF EPSCoR Award Number 1632738.

### 2021-01-12 paper/main.tex
finished pass through manuscript, updated diffs

@@
 
-One practical way to distinguish content from context representations is to separate out mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} or gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, FolkEtal18}.
+One practical way to distinguish content from context representations is to separate out mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05a, MannEtal12} versus gradually \citep[context;][]{PolyEtal05a, MannEtal11, HowaEtal12, LohnEtal18, FolkEtal18}.
 An intuition for why contextual representations might be expected to evolve gradually is laid out by \cite{PolyKaha08}.
@@
 
-In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
+In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized by projecting text embedding representations of the events onto lower-dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.
 \cite{HeusEtal21} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats} and~\ref{fig:reinstatement}, revealing another key property of how we remember.
@@
 
-In their paper, \cite{HeusEtal21} use the geometric shapes of experience trajectories (through word embedding space) to functionally distinguish high-level essential details of the experience from low-level (non-essential) details.
+In their paper, \cite{HeusEtal21} use the geometric shapes of experience trajectories (through text embedding space) to functionally distinguish high-level essential details of the experience from low-level (non-essential) details.
 When the trajectory exhibits a sharp change in direction, this indicates a rapid shift in the thematic content of the episode.
@@
 \subsection*{The \textit{matching problem}}
-The trajectories through word embedding space displayed in Figure~\ref{fig:trajectories} demonstrate that topic models are sufficiently expressive and robust to provide one solution to the \textit{matching problem} in memory analysis.
+The trajectories through text embedding space displayed in Figure~\ref{fig:trajectories} demonstrate that topic models are sufficiently expressive and robust to provide one solution to the \textit{matching problem} in memory analysis.
 The matching problem refers to discovering an alignment in the labels between experienced and remembered events.
@@
 Specifically, matching requires building explicit models of the dynamic conceptual content of our experiences and memories, and then estimating when and how that experienced and remembered content aligns.
-The approach featured in Figure~\ref{fig:trajectories} uses a word embedding space that is defined by a topic model~\citep{BleiEtal03}.
-Each coordinate in this space reflects the conceptual content underlying a set of words, such that two different text passages that describe the same conceptual content (but using different words, levels of detail, etc.) are mapped onto nearby points in the word embedding space.
-When these models are sufficiently expressive and robust to solve the matching problem, by modeling our experiences and memories within the same word embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through word embedding space~\citep[also see][]{HeusEtal21}.
+The approach featured in Figure~\ref{fig:trajectories} uses a text embedding space that is defined by a topic model~\citep{BleiEtal03}.
+Each coordinate in this space reflects the conceptual content underlying a set of words, such that two different text passages that describe the same conceptual content (but using different words, levels of detail, etc.) are mapped onto nearby points in the text embedding space.
+When these models are sufficiently expressive and robust to solve the matching problem, by modeling our experiences and memories within the same text embedding space, we can see a clear geometric alignment between the trajectories that our experiences (Fig.~\ref{fig:trajectories}A) and memories (Fig.~\ref{fig:trajectories}B) take through text embedding space~\citep[also see][]{HeusEtal21}.
 
@@
 Further, geometric differences in those embeddings (e.g., the Euclidean distances between the corresponding vectors) can capture subtle differences in conceptual content that might not have been readily apparent in the original text.
-Of course, there are also limits on how accurately specific word embedding models characterize different conceptual aspects of a given text passage, and on how informative or meaningful those differences are from a psychological perspective.
+Of course, there are also limits on how accurately specific text embedding models characterize different conceptual aspects of a given text passage, and on how informative or meaningful those differences are from a psychological perspective~\citep[also see][]{MannKaha12}.
 The field of text embedding has advanced rapidly, particularly in recent years, enabling more accurate and informative approaches to characterizing the conceptual content of arbitrary text passages.
@@
 The \textit{Neural Probabilistic Language Model}~\citep[NPLM;][]{BengEtal03} was one of the earliest deep learning-based sequence models.
-NLPM derives word embeddings for each word in the vocabulary (such that nearby embeddings reflect semantically related words), and then models word sequences according to their meanings.
+NLPM derives embeddings for each word in the vocabulary (such that nearby embeddings reflect semantically related words), and then models word sequences according to their meanings.
 This model was inspired in part by $n$-gram models~\citep{BrowEtal92}, which estimate the probability of the next word in a sequence given the $n$ words that came before it.
@@
 
-The next major advances in word embeddings came from attempts to more precisely capture the geometric relations between words.
+The next major advances in text embeddings came from attempts to more precisely capture the geometric relations between words.
 For example, if each word's semantic meaning is captured by a feature vector, then in principle one could perform vector algebra on different words' embeddings to produce new concepts that combine their meanings.
@@
 
-Cutting-edge approaches to word embeddings that have achieved the best-to-date performance on modern natural language processing benchmarks are built using enormous deep neural networks often with many billions of parameters.
+Cutting-edge approaches to text embeddings that have achieved the best-to-date performance on modern natural language processing benchmarks are built using enormous deep neural networks often with many billions of parameters.
 Because these models are very computationally expensive to fit, an important discovery has been that many layers in these models may be pre-trained without substantially affecting performance, even on new tasks.
@@
 In turn, this can help explain non-contiguous ``bumps'' in the quantum memory wave function (e.g., Figs.~\ref{fig:qwave}C, \ref{fig:reinstatement}B) and off-diagonal peaks in semantic temporal correlation matrices (e.g., Fig.~\ref{fig:corrmats}).
+The precision and accuracy with which text embedding models capture the conceptual nuances within a text passage also influence the accuracy and interpretability of the resulting trajectory of an experience or memory through embedding space (e.g., Fig.~\ref{fig:trajectories}).
+For example, an accurate but imprecise model might capture the low spatial frequency features of a trajectory's shape (i.e., the high-level details of the experience), whereas an accurate and precise model could also capture the high spatial frequency features (i.e., low-level details).
 
 \subsection*{Extensions to spatial, semantic, affective, and social representations}
-The way we navigate through \textit{thought space} (i.e., word embedding spaces; Figs.~\ref{fig:trajectories}, \ref{fig:embedding-models}) differs from how we navigate in physical spatial environments.
-First, we cannot teleport in real space, whereas our thoughts can exhibit rapid jumps between different non-contiguous regions of word embedding space.
+The way we navigate through \textit{thought space} (i.e., text embedding spaces; Figs.~\ref{fig:trajectories}, \ref{fig:embedding-models}) differs from how we navigate in physical spatial environments.
+First, we cannot teleport in real space, whereas our thoughts can exhibit rapid jumps between different non-contiguous regions of text embedding space.
 Second, when two people visit the same fixed physical landmark, they necessarily visit the same physical location.
